### 2001-04-16 fr/about/Constitution-fr.fr.tex
Add about section
svn path=/trunk/; revision=77

%% $Id: Constitution-fr.fr.tex,v 1.1 2001-04-16 18:10:09 loic Exp $
%% 
%% Constitution of the Free Software Foundation Chapter France
%%
%% Language: German
%%

%{{{ Header

\documentclass[a4wide,12pt]{scrartcl}
\usepackage{german,umlaut}
\usepackage{fancyheadings}
\usepackage{alltt}
\usepackage[german]{babel}

\selectlanguage{german}
\pagestyle{fancyplain}

\lhead[\fancyplain{}{\bfseries\thepage}]
        {\fancyplain{}{\bfseries\rightmark}}
\rhead[\fancyplain{}{\bfseries\leftmark}]
        {\fancyplain{}{\bfseries\thepage}}
\cfoot{}
\renewcommand{\sectionmark}[1]{\markright{\thesection\ #1}}
\renewcommand{\thesection}{\S~\arabic{section}}

%
% alternative ways of doing it...
%

% this makes the references work correctly:
%\newcommand{\A}{\subsection{}}
%\renewcommand{\subsectionmark}[1]{\markright{\thesection\ }}

% this looks better:
\newcounter{absatz}[section]
\newcommand{\A}{\par\vspace{1ex}
                \stepcounter{absatz}\noindent(\arabic{absatz})~~}

%}}}

\begin{document}

%{{{ 0 Préambule

\thispagestyle{empty}
\begin{center}
\Huge\bf STATUTS
\end{center}

\section*{Préambule}

L'entrée dans l'ère numérique, où l'espace réel et l'espace virtuel
vont coexister à titre d'égalité et marquer ensemble l'évolution
sociale, culturelle et scientifique de l'humanité, impose des
questions dont la formulation et le traitement sera le but à long
terme de la Free Software Foundation Europe.\\

A cet égard, ses tâches immédiates seront d'avancer de manière
désintéressée les Logiciels libres (en anglais, «Free
Software»), ainsi que de formuler et de disséminer les questions
philosophiques et sociales qui s'y attachent.\\

Le terme de Logiciels libres au sens employé par la FSF France ne fait
pas référence au prix, mais se rapporte aux quatre libertés suivantes:\\

\begin{enumerate}

\item 1ère liberté : La liberté d'utiliser un logiciel pour tout objectif ;

\item 2ème liberté : La liberté d'examiner le fonctionnement d'un logiciel
et de l'adapter à ses propres besoins ;

\item 3ème liberté : La liberté de faire des copies pour d'autres personnes ;

\item 4ème liberté : La liberté d'améliorer le logiciel et de rendre ces améliorations largement disponibles pour le bien public.
\end{enumerate}

Cette définition de Logiciels libres remonte à l'idée fondamentale du
libre échange de savoir et d'idées, comme on le trouve
traditionnellement dans les sciences. Les logiciels, comme les
pensées, sont immatériels et peuvent être copiés sans pertes. Leur
transmission sert un procès évolutif de développement continu d'idées
et de logiciels.\\

Seuls les Logiciels libres permettent la traçabilité et le
développement de résultats scientifiques. Dans le discours
scientifique, ils sont donc le seul type de logiciels correspondant
aux idéaux d'une science libre. L'avancement de Logiciels libres est
donc en même temps l'avancement de la recherche scientifique.\\

La dissémination d'informations et la formation d'opinions est de plus
en plus fortement marquée par les médias numériques, et il existe des
projets pour impliquer les citoyens plus directement dans la
démocratie avec l'aide de technologies numériques. L'une des tâches
centrales de la FSF France sera donc de former des citoyens émancipés
à ces égards et de faire progresser l'Etat démocratique.\\

L'espace numérique, dont les logiciels sont le médium et le langage,
dispose d'un formidable potentiel de progrès dans tous les domaines
d'intérêt spirituel et culturel pour l'humanité.  En rendant ce médium
communément disponible et ouvert, les Logiciels libres garantissent
l'égalité des possibilités et la protection du domaine privé.\\

Développer, parmi toutes les sections de la population, la
connaissance des questions liées à l'ère numérique, c'est là le but à
long terme et l'un des noyaux de l'activité de la FSF France.\\

La FSF France milite donc pour le renforcement de l'usage de Logiciels
libres dans les écoles et les universités, pour engendrer pendant la
période formative, parallèlement à la connaissance des questions de
l'espace réel, une conscience et une compréhension de l'espace
virtuel.\\

Les Logiciels libres garantissent des résultats et des processus de
prise de décision plus traçables dans les sciences et dans la vie
publique, ainsi que le droit de l'individu à l'épanouissement de sa
personnalité et de sa liberté d'opinion. La FSF France tâchera donc de
porter les Logiciels libres dans tous les domaines qui touchent à la
vie publique ou aux « droits de l'homme informationnels » des
citoyens.\\

Dans la suite de ce document «FSF Europe» désigne l'association à but non
lucratif dont le nom est «Free Software Foundation Europe» et dont le siège
social est localisé en Allemagne.

%}}}

%{{{ 1

\section{Nom}

Il est fondé entre les adhérents aux présents status une association
régie par la loi du 1er Juillet 1901 et le décret du 16 Août 1901,
ayant pour titre:\\

«Free Software Foundation Chapter France» \\

dont l'acronyme est\\

«FSF France»\\

L'association se nommera «Free Software Foundation Chapter
France» aussi longtemps qu'elle sera membre de la FSF Europe
au titre de personne morale et s'oblige à modifier son nom
si elle était exclue à titre définitif de la FSF Europe.

%}}}

%{{{ 2

\section{Objet}

L'objectif de la FSF France est l'avancement des buts cités dans le préambule. \\

Spécifiquement, l'objectif de l'association sera servi par :\\

le soutien à tous les organismes d'Etat et privés dans toutes les
questions de Logiciels libres, la collaboration coordonnée avec les
associations nationales poursuivant les mêmes objectifs et avec la FSF
Europe, le soutien de projets développant des Logiciels libres, la
dissémination des idéaux philosophiques des Logiciels libres.\\

La FSF France se dédie aux seules et immédiates finalités
scientifiques et d'utilité publique. La FSF France agit de manière
désintéressée et ne poursuit aucun objectif lucratif.\\

L'association se réserve le droit d'exercer toute action en justice
pour assurer la défense des intérêts individuels de la totalité ou d'une
partie de ses membres.\\

Les moyens de la FSF France ne doivent être utilisés que pour des
objectifs conformes à son statut. Nul ne doit être favorisé par des
dépenses qui seraient étrangères aux objectifs de la FSF France, ou
par une compensation hors proportion.\\

%}}}

%{{{ 3

\section{Durée}

La durée de l'association est illimitée.

%}}}

%{{{ 4

\section{Siège social}

Le siège social est fixé au 8, rue Riant 93200 Saint Denis. Il pourra
être transféré par simple décision du Conseil d'Administration.\\

%}}}

%{{{ 5

\section{Adhérents}

L'association se compose de membres actifs ou adhérents.

%}}}

%{{{ 6

\section{Admission}

Pour faire partie de l'association, il faut être agrée par le bureau
qui statue lors de chacune de ses réunions, sur les demandes
d'admission présentées.\\

Seuls les membres de la FSF Europe peuvent devenir membre de la FSF
France.

%}}}

%{{{ 7

\section{Les membres}

Les membres actifs se cotisent par une activité bénévole ou par
l'exercice d'une fonction dans l'association.\\

La qualité de membre se perd par:\\

\begin{enumerate}
\item La démission
\item Le décès
\item La radiation prononcée par le conseil d'Administration pour motif
grave, l'intéressé ayant été invité par lettre recommandée à se présenter 
devant le bureau pour fournir des explications.
\end{enumerate}

%}}}

%{{{ 8

\section{Ressources}

Les ressources de l'association comprennent:\\

\begin{enumerate}
\item les subventions qui pourront lui être accordées par les
 structures de l'Union Européenne, de l'Etat, des collectivités
 locales, des collectivités publiques ou des établissements publics,
 ainsi que d'associations ou toute autre personne morale dans les
 conditions légales

 \item toutes autres ressources ou subventions qui lui seraient
 accordées et qui ne seraient pas contraires aux lois en vigueur

\end{enumerate}

%}}}

%{{{ 9

\section{Conseil d'Administration}

L'association est dirigée par un Conseil d'Administration, élu pour 2 ans,
par l'assemblée générale. Les membres sont rééligibles.\\

Le Conseil d'Administration choisit parmis ses membres, au scrutin secret,
un bureau composé de:\\

\begin{enumerate}
\item Un président.
\item Un secrétaire et, s'il y a lieu, un secrétaire adjoint.
\item Un trésorier et, s'il y a lieu, un trésorier adjoint.
\end{enumerate}

Le président du bureau est le président de l'association.\\

Le conseil est investi des pouvoirs les plus étendus pour décider et
autoriser toutes opérations ayant pour but la réalisation de l'objet
de l'association ainsi que pour gérer sa gestion, son développement et
son fonctionnement.\\

Le conseil étant renouvelé chaque année par moitié, la première année les
membres sortants sont désignés par le sort. \\

En cas de vacances, le conseil d'Administration pourvoit provisoirement au
remplacement de ses membres. Il est procédé à leur remplacement 
définitif par la plus prochaine assemblée générale. Les pouvoirs des membres
ainsi élus prennent fin à l'époque où devrait normalement expirer le mandat
des membres remplacés.\\

Afin de faciliter les communications avec la FSF Europe il est convenu
que le président pourra porter le titre de Chancellor et que le secrétaire
pourra porter le titre de Vice Chancellor.

%}}}

%{{{ 10

\section{Réunion du Conseil d'Administration}

Le Conseil d'Administration se réunit une fois au moins tous les six mois
sur convocation du président, ou sur la demande du quart de ses membres.\\

Les décisions sont prises à la majorité des voix, en cas de partage, la
voix du président est prépondérante. \\

Tout membre du conseil qui, sans excuse, n'aura pas assisté à trois 
réunion consécutives, pourra être considéré comme démissionaire.\\

Nul ne peut faire partie du conseil s'il n'est pas majeur. \\
%}}}

%{{{ 11

\section{Rôle des membres du bureau}

Le président convoque les assemblées générales et les réunions du
conseil d'administration. Il représente l'association dans tous les
actes de la vie civile et est investi des pouvoirs à cet effet. Il
représente l'association en justice. En cas d'absence ou de maladie,
il est remplacé par le secrétaire qui dispose alors des mêmes
pouvoirs.\\

Le secrétaire est chargé de tout ce qui concerne la correspondance et
les archives. Il rédige les procès-verbaux des délibérations du
conseil et en assure la transcription sur les registres. En cas
d'absence ou de maladie, il est remplacé par un membre du conseil
désigné par le président qui dispose alors des mêmes pouvoirs.\\

Le trésorier est chargé de tout ce qui concerne la gestion du
patrimoine de l'association. Il effectue tous paiements et perçoit
toutes recettes sous la surveillance du président. Il tient une
comptabilité régulière de toutes les opérations. En cas d'absence ou
de maladie, il est remplacé par un membre du conseil désigné par le
président qui dispose alors des mêmes pouvoirs.\\

%}}}

%{{{ 12

\section{Assemblée Générale Ordinaire}

L'Assemblée Générale Ordinaire comprend tous les membres de
l'association à quelque titre qu'ils y soient affiliés. Les membres de
l'association sont convoqués au moins une fois par année civile en
assemblée générale.\\

Quinze jours au moins avant la date fixée, les membres de l'association
sont convoqués par les soins du secrétaire. L'ordre du jour est indiqué
sur les convocations. \\

L'ordre du jour doit inclure en
priorité l'examen des guidelines les plus récentes éditées par la FSF
Europe et les modalités de leur mise en oeuvre. \\

Pour délibérer valablement, l'assemblée générale devra être composée
d'au moins 50% des membres actifs, participants ou représentés. Dans
le cas où le quorum ne serait pas atteint, une seconde assemblée
générale sera convoquée et statuera valablement sans règle de quorum,
après un délai de 15 jours.\\

Le président, assisté des membres du conseil, préside l'assemblée et
expose la situation morale de l'association. \\

Le trésorier rend compte de sa gestion et soumet le bilan à l'approbation
de l'assemblée.\\

Les réunions de l'Assemblée Générale ne sont pas publiques.\\

%}}}

%{{{ 13

\section{Assemblée Générale Extraordinaire}

Si besoin est, ou sur la demande de la moitié plus un des membres inscrits,
le Président peut convoquer une Assemblée Générale Extraordinaire, suivant
les formalités prévues par l'Article 10.\\

L'assemblée générale, convoquée de façon extraordinaire, délibère
exclusivement sur les questions portées à son ordre du jour.\\

Elle peut, en particulier, modifier les statuts de l'association ou
proposer la dissolution de l'association, mais seulement sur
proposition du conseil d'administration. Ces décisions ne peuvent
alors être votées que si les 3/4 des membres de l'association sont
présents ou représentés, et à la majorité qualifiée des 3/4 des
votants. Si une première assemblée ne réunit pas le quorum des 3/4,
une seconde assemblée doit être convoquée dans un délai d'un mois et
peut valablement délibérer.\\

%}}}

%{{{ 14

\section{Règlement Intérieur}

Un règlement intérieur peut être établi par le Conseil d'Administration
qui le fait alors approuver par l'Assemblée Générale. \\

Ce règlement éventuel est destiné à fixer les divers points non prévus
par les statuts, notament ceux qui ont trait à l'administration interne
de l'association.

%}}}

%{{{ 15

\section{Dissolution}

En cas de dissolution prononcée par les deux tiers au moins des membres
présents à l'Assemblée Générale, un ou plusieurs liquidateurs sont nommés 
par celle-ci et l'actif, s'il y a lieu, est dévolu conformément à l'article 
9 de la loi du 1er Juillet 1901 et au décret du 16 Août 1901.

%}}}

%{{{ 16

\section{Forme écrite}

\begin{enumerate}
\item Un document papier signé manuellement.
\item Un courrier électronique encodé par un code correspondant au
standards actuels d'encodage.
\end{enumerate}

%}}}

\thispagestyle{empty}
\begin{center}
\bf Fait à Paris, le 28 Avril 2001
\end{center}

\end{document}

### 2001-04-17 fr/about/Constitution-fr.fr.tex
Format statute
svn path=/trunk/; revision=79
Fix spelling and add a section
svn path=/trunk/; revision=89

@@
-%% $Id: Constitution-fr.fr.tex,v 1.1 2001-04-16 18:10:09 loic Exp $
+%% $Id: Constitution-fr.fr.tex,v 1.3 2001-04-17 11:36:48 mad Exp $
 %% 
@@
 A cet égard, ses tâches immédiates seront d'avancer de manière
-désintéressée les Logiciels libres (en anglais, «Free
+désintéressée les Logiciels Libres (en anglais, «Free
 Software»), ainsi que de formuler et de disséminer les questions
@@
 
-Le terme de Logiciels libres au sens employé par la FSF France ne fait
+Le terme de Logiciels Libres au sens employé par la FSF France ne fait
 pas référence au prix, mais se rapporte aux quatre libertés suivantes:\\
@@
 
-Cette définition de Logiciels libres remonte à l'idée fondamentale du
+Cette définition de Logiciels Libres remonte à l'idée fondamentale du
 libre échange de savoir et d'idées, comme on le trouve
@@
 
-Seuls les Logiciels libres permettent la traçabilité et le
+Seuls les Logiciels Libres permettent la traçabilité et le
 développement de résultats scientifiques. Dans le discours
 scientifique, ils sont donc le seul type de logiciels correspondant
-aux idéaux d'une science libre. L'avancement de Logiciels libres est
+aux idéaux d'une science libre. L'avancement de Logiciels Libres est
 donc en même temps l'avancement de la recherche scientifique.\\
@@
 d'intérêt spirituel et culturel pour l'humanité.  En rendant ce médium
-communément disponible et ouvert, les Logiciels libres garantissent
+communément disponible et ouvert, les Logiciels Libres garantissent
 l'égalité des possibilités et la protection du domaine privé.\\
@@
 La FSF France milite donc pour le renforcement de l'usage de Logiciels
-libres dans les écoles et les universités, pour engendrer pendant la
+Libres dans les écoles et les universités, pour engendrer pendant la
 période formative, parallèlement à la connaissance des questions de
@@
 
-Les Logiciels libres garantissent des résultats et des processus de
+Les Logiciels Libres garantissent des résultats et des processus de
 prise de décision plus traçables dans les sciences et dans la vie
@@
 personnalité et de sa liberté d'opinion. La FSF France tâchera donc de
-porter les Logiciels libres dans tous les domaines qui touchent à la
+porter les Logiciels Libres dans tous les domaines qui touchent à la
 vie publique ou aux « droits de l'homme informationnels » des
@@
 
-Il est fondé entre les adhérents aux présents status une association
+Il est fondé entre les adhérents aux présents statuts une association
 régie par la loi du 1er Juillet 1901 et le décret du 16 Août 1901,
@@
 le soutien à tous les organismes d'Etat et privés dans toutes les
-questions de Logiciels libres, la collaboration coordonnée avec les
+questions de Logiciels Libres, la collaboration coordonnée avec les
 associations nationales poursuivant les mêmes objectifs et avec la FSF
-Europe, le soutien de projets développant des Logiciels libres, la
-dissémination des idéaux philosophiques des Logiciels libres.\\
+Europe, le soutien de projets développant des Logiciels Libres, la
+dissémination des idéaux philosophiques des Logiciels Libres.\\
 
@@
 
-Le siège social est fixé au 8, rue Riant 93200 Saint Denis. Il pourra
-être transféré par simple décision du Conseil d'Administration.\\
+Le siège social est fixé au 8, rue Riant 93200 Saint Denis.\\
+
+Il pourra être transféré par simple décision du Conseil
+d'Administration.\\
 
@@
 
-Pour faire partie de l'association, il faut être agrée par le bureau
+Pour faire partie de l'association, il faut être agréé par le bureau
 qui statue lors de chacune de ses réunions, sur les demandes
@@
 \item Le décès
-\item La radiation prononcée par le conseil d'Administration pour motif
-grave, l'intéressé ayant été invité par lettre recommandée à se présenter 
-devant le bureau pour fournir des explications.
+\item La radiation prononcée par le conseil d'Administration pour
+infraction aux présents statuts, pour non respect du règlement
+intérieur ou pour motif grave portant préjudice moral ou matériel à
+l'association. L'intéressé ayant été invité par lettre recommandée à
+se présenter devant le bureau pour fournir des explications. 
 \end{enumerate}
@@
 
-En cas de vacances, le conseil d'Administration pourvoit provisoirement au
+En cas de vacances, le Conseil d'Administration pourvoit provisoirement au
 remplacement de ses membres. Il est procédé à leur remplacement 
@@
 %{{{ 12
+\section{Gratuité du mandat}
+
+Les membres du conseil d'administration de l'association ne peuvent
+recevoir aucune rétribution à raison des fonctions qui leur sont
+conférées. Ils pourront toutefois obtenir le remboursement des
+dépenses engagées pour les besoins de l'association sur justification
+et après accord du bureau.
+
+%}}}
+
+%{{{ 13
 
@@
 
-%{{{ 13
+%{{{ 14
 
@@
 
-%{{{ 14
+%{{{ 15
 
@@
 
-%{{{ 15
+%{{{ 16
 
@@
 
-%{{{ 16
+%{{{ 17
 
@@
 \begin{center}
-\bf Fait à Paris, le 28 Avril 2001
+\bf Fait à Paris, le 19 Avril 2001
 \end{center}

### 2001-04-19 fr/about/Constitution-fr.fr.tex
* fr/about/Constitution-fr.fr.tex: Specify that France is the   realm of action, a member excluded from FSF Europe is excluded   from FSF France.
svn path=/trunk/; revision=90
Fix latex bug
svn path=/trunk/; revision=100
s/Chapter/Europe - Chapter/
svn path=/trunk/; revision=112

@@
-%% $Id: Constitution-fr.fr.tex,v 1.3 2001-04-17 11:36:48 mad Exp $
+%% $Id: Constitution-fr.fr.tex,v 1.6 2001-04-19 06:46:07 loic Exp $
 %% 
-%% Constitution of the Free Software Foundation Chapter France
+%% Constitution of the Free Software Foundation Europe - Chapter France
 %%
@@
 
-«Free Software Foundation Chapter France» \\
+«Free Software Foundation Europe - Chapter France» \\
 
@@
 
-L'association se nommera «Free Software Foundation Chapter
+L'association se nommera «Free Software Foundation Europe - Chapter
 France» aussi longtemps qu'elle sera membre de la FSF Europe
@@
 par une compensation hors proportion.\\
+
+La FSF France focalise son action sur le territoire Français et
+coopère avec la FSF Europe pour les sujet impliquant d'autre pays.\\
 
@@
 \end{enumerate}
+
+L'exclusion de la FSF Europe est réputé être un motif grave suite auquel
+le conseil d'Administration doit statuer sur la radiation du ou des membres
+concernés.\\
 

### 2001-05-29 fr/about/article-fsfeurope.tex
Add temporary file
svn path=/trunk/; revision=628

% $Id: article-fsfeurope.tex,v 1.1 2001-05-29 12:07:42 mad Exp $
% $Revision: 1.1 $

\documentclass[a4paper]{article}
\usepackage{isolatin1}
\usepackage[T1]{fontenc}
\usepackage[french]{babel}
\usepackage{url}
\usepackage{boxedminipage}

\renewcommand{\rmdefault}{phv}
\renewcommand{\sfdefault}{phv}
\renewcommand{\ttdefault}{pcr}

\setlength{\parskip}{1.2ex}
\setlength{\parindent}{0cm}
\setlength{\oddsidemargin}{-0.5cm}
\setlength{\evensidemargin}{-0.5cm}
\setlength{\marginparwidth}{0cm}
\setlength{\marginparsep}{0cm}

\setlength{\topmargin}{-1cm}
\setlength{\textheight}{25cm}
\setlength{\textwidth}{17cm}

\usepackage{epsf}
\newcommand{\screendump}[2]{%   
  \begingroup                % keep redefinition of \epsfsize local
    \def\epsfsize##1##2{#1##1}% scale to #1% of natural size
    \epsfbox{#2}%            % read image
  \endgroup
}

\begin{document}

  \title{\Huge\bf FSF Europe\\Liberté et Coopération}
  \author{Fred Couchet}
  \date{28 mai 2001}
  \maketitle

\vskip 1cm

\begin{abstract}
  Ce document est un article de présentation de la FSF Europe et plus
  particulièrement de la FSF France.
\end{abstract}

\section*{Présentation}

La Free Software Foundation Europe est l'organisation européenne soeur
de la Free Software Foundation fondée en 1985 par Richard M. Stallman
(l'organisation américaine historique dédiée à l'abolition des
restrictions sur la copie, la redistribution, la compréhension et la
modification des programmes informatiques). La FSF promeut le
développement et l'utilisation du Logiciel Libre, et est l'initiatrice
du projet GNU, qui a notamment donné naissance au système
d'exploitation GNU/Linux.

Comme la FSF originelle, la FSF Europe fournit une organisation
politique et lobbyiste pour le Logiciel Libre et le Projet GNU. Elle
travaille pour devenir l'épine dorsale pour les parties du Projet GNU
basées en Europe et entretient des contacts proches avec la presse et
les représentants politiques pour accroître la couverture du Logiciel
Libre et du Projet GNU. Elle aide en outre à lever et à distribuer des
fonds pour le Logiciel Libre partout en Europe. Pour cette raison,
elle est exempte de taxes partout où c'est possible.

\section*{Organisation}

Comme il n'y a pas de législation européenne commune pour les
institutions, la FSF Europe est organisée de façon modulaire. Elle
consiste en une organisation pivot fondée à Hambourg (Allemagne) avec
des sous-organisations dans chaque pays d'Europe. Ces chapitres locaux
auront un chancelier (président) ainsi qu'un vice-chancelier
(secrétaire) parlant la langue du pays, disponibles pour la presse
locale et les représentants politiques, et en contact permanent avec
la communauté du Logiciel Libre du pays. Des organisations, auxquelles
tout le monde peut prendre part, pourront devenir organisations
associées et constituer les muscles du squelette de la FSF
Europe. Enfin, les volontaires pourront participer notamment via
Savannah. De cette façon, chacun peut être impliqué dans le Mouvement
du Logiciel Libre et la FSF Europe.

La FSF Europe travaille de concert avec ses organisations
associées. Deux de ces organisations sont enregistrées en France :
l'Association pour la Promotion et la Recherche en Informatique Libre
(APRIL, fondée en 1996) qui fut la première organisation associée à la
FSF Europe, et l'Organization for Free Software in Education and
Teaching (OFSET), très impliquée dans le domaine de l'éducation.

\section*{Objectifs}

La FSF Europe a deux objectifs principaux :

\subsubsection*{Compléter l'offre Logiciel Libre}

Aujourd'hui il n'existe pas de Logiciels Libres pour tous les
besoins. La FSF, tout comme la FSF Europe, oeuvre à compléter l'offre
Logiciel Libre afin qu'il soit possible d'utiliser un Logiciel Libre
quel que soit le domaine d'activité. Par exemple, il est aujourd'hui
difficile voire impossible de transmettre efficacement du son et de
l'image sur les réseaux en utilisant du Logiciel Libre.

Le rôle de la FSF et de la FSF Europe est d'encourager les entreprises
et les individus à identifier les friches du Logiciel Libre et à les
compléter. Il suffit souvent d'un simple dialogue pour qu'un créateur
réalise qu'il peut combler un manque, compléter une brèche dans
l'univers Logiciel Libre. Ainsi, a l'occasion des FSF Awards au début
de l'année 2001, après un dialogue intense et constructif, l'école
Centrale Paris (ECP) a édité sous licence GNU GPL les développements
de VideoLAN, ajoutant ainsi une pierre importante à l'édifice.

Par ailleurs, la FSF étant à l'origine de l'univers Logiciel Libre qui
existe aujourd'hui, les entreprises pourront s'y intégrer facilement
par son intermédiaire. Un aspect que réalisent difficilement les
individus et les entrepreneurs est la facilité d'établir des relations
bilatérales dans la communauté du Logiciel Libre. Une entreprise
développant son activité sur MySQL, GNU/Linux et Perl peut facilement
entrer en contact avec les groupes d'individus et entreprises qui
développent et maintiennent ces éléments. Ils n'ont besoin de personne
pour établir ce contact mais l'expérience nous a montré qu'une
réticence naturelle empêche qu'il s'établisse. Le rôle de la FSF
Europe est alors simplement de montrer comment établir ce contact et
maintenir un dialogue. L'éloignement, la communication par e-mail, la
sociologie des groupes de développement rendent cet exercice
suffisamment inhabituel pour qu'une aide extérieure soient la
bienvenue.

En bref, établir des relations avec la communauté du Logiciel Libre
est un aspect indispensable qui doit compléter l'activité de
rapprochement avec les entreprises au niveau national. Les entreprises
doivent pouvoir s'intégrer au mouvement du Logiciel Libre et non
seulement se reposer dessus. Elles consolideront ainsi leurs
fondations.  Aussi impressionnant qu'il y paraisse, le mouvement du
Logiciel Libre est assez petit pour qu'un groupe de quelques
entreprises puissent y jouer un rôle considérable. C'est une chance à
saisir.

\subsubsection*{Combattre les dangers qui menacent les fondements du
Logiciel Libre}

Les lois autour du logiciel sont très mouvantes car il s'agit d'une
science jeune. Par manque d'information nos législateurs peuvent être
amenés à prendre des mesures qui interdisent l'existence du Logiciel
Libre. Les quatre libertés fondamentales du Logiciel Libre sont encore
fragiles~:

\begin{itemize}
\item Utilisation - la liberté d'utiliser le logiciel, pour quelque usage que ce soit
\item Étude - la liberté d'étudier le fonctionnement du programme, et de
  l'adapter à vos propres besoins. L'accès au code source est une
  condition pour tout ceci
\item Redistribution - la liberté de redistribuer des copies de façon à pouvoir aider
  votre voisin
\item Modification - la liberté d'améliorer le programme, et de diffuser vos
  améliorations au public, de façon à ce que l'ensemble de la
  communauté en tire avantage. L'accès au code source est
  une condition pour tout ceci
\end{itemize}

Par exemple la loi SIAE en Italie impose que chaque logiciel faisant
l'objet d'une facturation fasse l'objet d'un enregistrement officiel.
Cet enregistrement a bien entendu un coût. Ainsi, vendre une
distribution GNU/Linux (Debian par exemple) impose d'enregistrer
individuellement les milliers de logiciels qu'elle contient. Cette
nouvelle loi handicape sérieusement la liberté de distribution et
menace donc directement l'existence du Logiciel Libre.

Une autre menace est constituée par les brevets logiciels.  Chaque
logiciel libre est potentiellement à la merci d'un brevet logiciel. En
effet il est pratiquement impossible de déterminer avec certitude si
un logiciel tombe sous les coup d'un brevet ou non.  Si les brevets
logiciels devenaient légaux en Europe il serait donc potentiellement
illégal d'utiliser, modifier ou redistribuer l'ensemble des Logiciels
Libres.

Nous souhaiterions que la situation soit moins périlleuse. Mais comme
ce n'est pas le cas, la FSF Europe doit lutter pour consolider les
fondements légaux de l'univers des Logiciels Libres. La tâche est
d'autant plus importante que la FSF est auteur de la licence GNU GPL
qui couvre près de 70\% des Logiciels Libres existant.

La FSF Europe peut fournir aux entreprises, administrations une
information précise sur certains problèmes légaux et elles pourront
agir et veiller avec nous pour consolider les bases juridiques du
Logiciel Libre. La FSF Europe s'engage actuellement à assister et
coordonner le processus d'adaptation de la licence GNU GPL au droit
français et allemand, et plus généralement met en place un réseau de
juristes compétents dans le domaine du logiciel libre.

La FSF Europe joue un rôle fédérateur dans chaque pays, rassemblant
les acteurs du Logiciel Libre afin de leur permettre de communiquer.
Grâce aux présences que nous avons établies en Suède, France,
Allemagne, Italie et bientôt en Suisse, au Portugal, en Belgique et en
Autriche nous pouvons agir concrètement au niveau Européen. Dans
chaque pays des relations de travail se mettent en place avec les
associations qui défendent depuis de nombreuses années le Logiciel
Libre afin d'apporter une dimension Européenne à leur action. L'APRIL
en France peut par exemple coordonner les efforts de traduction de la
licence GNU GPL avec des juristes français.

\section*{Actualité}

La FSF Europe a démarré officiellement ses activités le 10 mars
2001. La première assemblée générale de la FSF Europe a eu lieu le 6
mai 2001, à Essen (Allemagne). Les premiers chapitres locaux
(Allemagne et France) sont actuellement créés et actifs. Les chapitres
Suédois et Italien devraient être actifs très bientôt.

\subsubsection*{Chronologie des actions du chapitre Français}

La FSF Europe a de fait commencé ses activités en France le 12
avril. Après la création des aspect administratifs (par exemple
l'hébergement du siège social par Lolix) et d'infrastructure
(notamment l'installation de la machine fr.fsf.org, première machine
intégré au projet GNU présente hors des USA, chez Nevrax). L'activité
quotidienne du chapitre français utilise des ressources GNU (DNS,
Savannah). Une des premières taches importantes a été l'établissement
de relation de travail avec la communauté française du logiciel libre
(notamment à travers l'APRIL et OFSET).

\paragraph{Traduction de la GNU GPL en français} Mélanie Clément
Fontaine, juriste chez Alcôve et doctorante en propriété
intellectuelle, a démarré la traduction et l'adaptation au droit
français de la GNU GPL. Elle reçoit maintenant le support de la FSF
Europe, qui démarre également un effort similaire en Allemagne. Plus
généralement, la FSF Europe va regrouper des juristes compétents et
intéressés par le logiciel libre pour constituer un réseau d'expert
qui pourra l'aider dans le futur.

\paragraph{Rencontres}  La FSF France a rencontré de nombreuses
personnes à l'occasion de diverses manifestations (comme <<~Libre et
Vie Locale~>> par exemple) et a commencé à établir des contacts avec
différents projets et institutions. Le premier exemple concret fût la
rencontre avec les responsables de la future Technopole Logiciel Libre
à Soissons. Soutenue par le gouvernement et en liaison étroite avec
l'université de Soissons, la Technopole Logiciel Libre accompagnera
les entreprises développant une activité commerciale basée sur les
Logiciels Libres.  En dispensant des formations initiant les chefs
d'entreprise aux modèles économiques créés par le Logiciel Libre, la
technopole redonnera vie à un paysage économique appauvri par la
disparition des industries les plus importantes de la région. La
dernière en date concerne une réunion avec la MTIC (Mission
interministérielle de soutien technique pour le développement des
technologies de l'information et de la communication dans
l'administration) et les responsables du <<~bouquet du libre~>>. Ces
rencontres ont été l'occasion de présenter la FSF France et aussi de
mettre en relation différentes personnes intéressées par des sujets
connexes, mais qui ne se connaissaient pas.

\paragraph{Savannah et l'Europe}  Savannah est un clone de SourceForge
dédié au projet GNU et à tout projet Logiciel Libre. Grâce aux dons de
membres de l'APRIL et de la FSF Europe, une nouvelle machine va être
installée pour Savannah. Les développeurs les plus actifs de Savannah
sont en France et au Portugal. Nous avons également pris contacts avec
les responsables de deux projets similaires à Savannah (PeCoVall,
pure/source) et discuté de la manière de partager les efforts et les
ressources pour obtenir une plate-forme de développement coopérative
encore plus puissante.

\paragraph{Sensibilisation}  La FSF Europe a eu des entretiens avec des journalistes pour
présenter ses objectifs et ses activités. La présence lors de diverses
manifestations a permis également de sensibiliser les gens sur
l'importance de la FSF Europe.

\section*{Où puis-je obtenir d'autres informations~?}

Pour obtenir davantage d'informations sur la FSF Europe, vous devriez
vous abonner aux listes de diffusion. Pour ce faire et pour accéder
aux archives, rendez-vous sur la page d'accueil {\sc
http://www.fsfeurope.org}. Vous pouvez également entrer en contact
avec les personnes mentionnées ci-dessous ou envoyer un email à toute
l'équipe de la FSF Europe à {\tt team@fsfeurope.org}.

Vous pouvez également lire le channel de news de la FSF France à
l'adresse \url{http://france.fsfeurope.org/news/}.

\section*{Références}

\begin{itemize}
\item Free Software Foundation - \url{http://www.fsf.org/}

\item Projet GNU - \url{http://www.gnu.org/}

\item Free Software Foundation Europe -
\url{http://www.fsfeurope.org/}

\item Free Software Foundation Europe Chapter France aka FSF France -
\url{http://france.fsfeurope.org/}

\item Free Software Foundation Europe -
\url{http://www.fsfeurope.org/}

\item Savannah - \url{http://savannah.gnu.org/projects/fsfe/}

\item APRIL - \url{http://www.april.org/}

\item OFSET - \url{http://www.ofset.org/}

\item VideoLan - \url{http://www.videolan.org/}

\item Ecole Centrale de Paris - \url{http://www.ecp.fr/}

\item Loi SIAE - \url{http://www.fsfeurope.org/siae/siae.fr.html}

\item Lolix - \url{http://www.lolix.org/}

\item Nevrax - \url{http://www.nevrax.org/}

\item Alcôve - \url{http://www.alcove.com/}

\item Technopole Logiciel Libre -
\url{http://www.soissons-technopole.org/}

\item MTIC - \url{http://www.mtic.pm.gouv.fr/}

\item PeCoVall - \url{http://pecovall.eu.org/}

\end{itemize}

\section*{Contacts}

\subsection*{Coordonnées postales de la FSF France}

\textbf{Association FSF Europe - Chapitre France}

8, Rue de Valois\\
75001 Paris\\
e-mail: \url{contact@france.fsfeurope.org}\\

\subsection*{Contacts avec le bureau}

Frédéric Couchet, Président. E-mail : \url{FredericCouchet@fsfeurope.org}.

Raphaël Rousseau, Secrétaire. E-mail : \url{RaphaëlRousseau@fsfeurope.org}.

Loïc Dachary, Trésorier. E-mail : \url{LoicDachary@fsfeurope.org}.

\vskip 1cm

\begin{boxedminipage}{\textwidth}

{\sf
Copyright \copyright~2001 FSF Europe - Chapter France - \url{http://france.fsfeurope.org/}\\

Ce document peut être reproduit par n'importe quel moyen que
ce soit, pourvu qu'aucune modification ne soit effectuée et que cette
notice soit préservée.}

\end{boxedminipage}

\end{document}

### 2001-05-29 fr/about/article-fsfeurope.tex
ortho/typo
svn path=/trunk/; revision=630

@@
-% $Id: article-fsfeurope.tex,v 1.1 2001-05-29 12:07:42 mad Exp $
-% $Revision: 1.1 $
+% $Id: article-fsfeurope.tex,v 1.2 2001-05-29 13:17:27 r4f Exp $
+% $Revision: 1.2 $
 
@@
   \title{\Huge\bf FSF Europe\\Liberté et Coopération}
-  \author{Fred Couchet}
+  \author{Frédéric Couchet}
   \date{28 mai 2001}
@@
 (l'organisation américaine historique dédiée à l'abolition des
-restrictions sur la copie, la redistribution, la compréhension et la
+restrictions sur l'utilisation, la compréhension, la redistribution et la
 modification des programmes informatiques). La FSF promeut le
@@
 Comme la FSF originelle, la FSF Europe fournit une organisation
-politique et lobbyiste pour le Logiciel Libre et le Projet GNU. Elle
-travaille pour devenir l'épine dorsale pour les parties du Projet GNU
+politique et lobbyiste pour le Logiciel Libre et le projet GNU. Elle
+travaille pour devenir l'épine dorsale pour les parties du projet GNU
 basées en Europe et entretient des contacts proches avec la presse et
 les représentants politiques pour accroître la couverture du Logiciel
-Libre et du Projet GNU. Elle aide en outre à lever et à distribuer des
+Libre et du projet GNU. Elle aide en outre à lever et à distribuer des
 fonds pour le Logiciel Libre partout en Europe. Pour cette raison,
@@
 réalise qu'il peut combler un manque, compléter une brèche dans
-l'univers Logiciel Libre. Ainsi, a l'occasion des FSF Awards au début
-de l'année 2001, après un dialogue intense et constructif, l'école
+l'univers Logiciel Libre. Ainsi, à l'occasion des FSF Awards au début
+de l'année 2001, après un dialogue intense et constructif, l'École
 Centrale Paris (ECP) a édité sous licence GNU GPL les développements
@@
 doivent pouvoir s'intégrer au mouvement du Logiciel Libre et non
-seulement se reposer dessus. Elles consolideront ainsi leurs
+pas uniquement se reposer dessus. Elles consolideront ainsi leurs
 fondations.  Aussi impressionnant qu'il y paraisse, le mouvement du
 Logiciel Libre est assez petit pour qu'un groupe de quelques
-entreprises puissent y jouer un rôle considérable. C'est une chance à
+entreprises puissent y jouer un rôle considérable ; c'est une chance à
 saisir.
@@
 http://www.fsfeurope.org}. Vous pouvez également entrer en contact
-avec les personnes mentionnées ci-dessous ou envoyer un email à toute
+avec les personnes mentionnées ci-dessous ou envoyer un e-mail à toute
 l'équipe de la FSF Europe à {\tt team@fsfeurope.org}.
@@
 
-\item Free Software Foundation Europe Chapter France aka FSF France -
+\item Free Software Foundation Europe Chapter France AKA FSF France -
 \url{http://france.fsfeurope.org/}
@@
 
-\item Ecole Centrale de Paris - \url{http://www.ecp.fr/}
+\item École Centrale de Paris - \url{http://www.ecp.fr/}
 
@@
 75001 Paris\\
+FRANCE\\
 e-mail: \url{contact@france.fsfeurope.org}\\
@@
 
-Raphaël Rousseau, Secrétaire. E-mail : \url{RaphaëlRousseau@fsfeurope.org}.
+Raphaël Rousseau, Secrétaire. E-mail : \url{RaphaelRousseau@fsfeurope.org}.
 

### 2001-05-29 fr/about/article-fsfeurope.tex
ortho
svn path=/trunk/; revision=631

@@
-% $Id: article-fsfeurope.tex,v 1.2 2001-05-29 13:17:27 r4f Exp $
-% $Revision: 1.2 $
+% $Id: article-fsfeurope.tex,v 1.3 2001-05-29 13:54:21 r4f Exp $
+% $Revision: 1.3 $
 
@@
 Une autre menace est constituée par les brevets logiciels.  Chaque
-logiciel libre est potentiellement à la merci d'un brevet logiciel. En
+Logiciel Libre est potentiellement à la merci d'un brevet logiciel. En
 effet il est pratiquement impossible de déterminer avec certitude si
@@
 Logiciel Libre. La FSF Europe s'engage actuellement à assister et
-coordonner le processus d'adaptation de la licence GNU GPL au droit
+coordonner le processus d'adaptation de la licence GNU GPL aux droits
 français et allemand, et plus généralement met en place un réseau de
-juristes compétents dans le domaine du logiciel libre.
+juristes compétents dans le domaine du Logiciel Libre.
 
@@
 quotidienne du chapitre français utilise des ressources GNU (DNS,
-Savannah). Une des premières taches importantes a été l'établissement
-de relation de travail avec la communauté française du logiciel libre
+Savannah). Une des premières tâches importantes a été l'établissement
+de relations de travail avec la communauté française du Logiciel Libre
 (notamment à travers l'APRIL et OFSET).
@@
 généralement, la FSF Europe va regrouper des juristes compétents et
-intéressés par le logiciel libre pour constituer un réseau d'expert
+intéressés par le Logiciel Libre pour constituer un réseau d'experts
 qui pourra l'aider dans le futur.
@@
 Vie Locale~>> par exemple) et a commencé à établir des contacts avec
-différents projets et institutions. Le premier exemple concret fût la
+différents projets et institutions. Le premier exemple concret fut la
 rencontre avec les responsables de la future Technopole Logiciel Libre
-à Soissons. Soutenue par le gouvernement et en liaison étroite avec
+à Soissons. Soutenue par le gouvernement français et en liaison étroite avec
 l'université de Soissons, la Technopole Logiciel Libre accompagnera
@@
 interministérielle de soutien technique pour le développement des
-technologies de l'information et de la communication dans
+Technologies de l'Information et de la Communication dans
 l'administration) et les responsables du <<~bouquet du libre~>>. Ces
@@
 installée pour Savannah. Les développeurs les plus actifs de Savannah
-sont en France et au Portugal. Nous avons également pris contacts avec
+sont en France et au Portugal. Nous avons également pris contact avec
 les responsables de deux projets similaires à Savannah (PeCoVall,
-pure/source) et discuté de la manière de partager les efforts et les
+Pure/Source) et discuté de la manière de partager les efforts et les
 ressources pour obtenir une plate-forme de développement coopérative

### 2001-05-31 fr/about/article-fsfeurope.tex
Added SourceForge in the references
svn path=/trunk/; revision=632
Patch for Alain Forcioli
svn path=/trunk/; revision=677

@@
-% $Id: article-fsfeurope.tex,v 1.3 2001-05-29 13:54:21 r4f Exp $
-% $Revision: 1.3 $
+% $Id: article-fsfeurope.tex,v 1.5 2001-05-31 09:43:48 mad Exp $
+% $Revision: 1.5 $
 
@@
 modification des programmes informatiques). La FSF promeut le
-développement et l'utilisation du Logiciel Libre, et est l'initiatrice
+développement et l'utilisation du Logiciel Libre, et est à l'origine
 du projet GNU, qui a notamment donné naissance au système
@@
 des sous-organisations dans chaque pays d'Europe. Ces chapitres locaux
-auront un chancelier (président) ainsi qu'un vice-chancelier
-(secrétaire) parlant la langue du pays, disponibles pour la presse
-locale et les représentants politiques, et en contact permanent avec
+ont un chancelier (président) ainsi qu'un vice-chancelier
+(secrétaire) parlant la langue du pays. Ils sont disponibles pour la presse
+locale et les représentants politiques, et sont en contact permanent avec
 la communauté du Logiciel Libre du pays. Des organisations, auxquelles
@@
 
-Par exemple la loi SIAE en Italie impose que chaque logiciel faisant
-l'objet d'une facturation fasse l'objet d'un enregistrement officiel.
-Cet enregistrement a bien entendu un coût. Ainsi, vendre une
-distribution GNU/Linux (Debian par exemple) impose d'enregistrer
-individuellement les milliers de logiciels qu'elle contient. Cette
-nouvelle loi handicape sérieusement la liberté de distribution et
-menace donc directement l'existence du Logiciel Libre.
+Par exemple la loi SIAE en Italie impose que chaque logiciel destiné à
+être vendu fasse l'objet d'un enregistrement officiel. Cet
+enregistrement a bien entendu un coût. Ainsi, vendre une distribution
+GNU/Linux (Debian par exemple) impose d'enregistrer individuellement
+les milliers de logiciels qu'elle contient. Cette nouvelle loi
+handicape sérieusement la liberté de distribution et menace donc
+directement l'existence du Logiciel Libre.
 
@@
 
-La FSF Europe peut fournir aux entreprises, administrations une
-information précise sur certains problèmes légaux et elles pourront
-agir et veiller avec nous pour consolider les bases juridiques du
-Logiciel Libre. La FSF Europe s'engage actuellement à assister et
-coordonner le processus d'adaptation de la licence GNU GPL aux droits
-français et allemand, et plus généralement met en place un réseau de
+La FSF Europe peut fournir aux entreprises et aux administrations une
+information précise sur certains problèmes légaux. Elles pourront agir
+et veiller avec nous pour consolider les bases juridiques du Logiciel
+Libre. La FSF Europe s'engage actuellement à assister et coordonner le
+processus d'adaptation de la licence GNU GPL au droit français et
+allemand. De manière plus générale, elle met en place un réseau de
 juristes compétents dans le domaine du Logiciel Libre.
@@
 
-La FSF Europe a démarré officiellement ses activités le 10 mars
-2001. La première assemblée générale de la FSF Europe a eu lieu le 6
-mai 2001, à Essen (Allemagne). Les premiers chapitres locaux
-(Allemagne et France) sont actuellement créés et actifs. Les chapitres
-Suédois et Italien devraient être actifs très bientôt.
+La FSF Europe a démarré officiellement ses activités le 10 mars 2001.
+Sa première assemblée générale a eu lieu le 6 mai 2001, à Essen
+(Allemagne). Les premiers chapitres locaux (Allemagne et France) sont
+actuellement créés et actifs. Les chapitres Suédois et Italien
+devraient être actifs très bientôt.
 
@@
 
-La FSF Europe a de fait commencé ses activités en France le 12
-avril. Après la création des aspect administratifs (par exemple
-l'hébergement du siège social par Lolix) et d'infrastructure
-(notamment l'installation de la machine fr.fsf.org, première machine
-intégré au projet GNU présente hors des USA, chez Nevrax). L'activité
+La FSF Europe a de fait commencé ses activités en France le 12 avril.
+Après la création des aspect administratifs (par exemple l'hébergement
+du siège social par Lolix) et d'infrastructure (notamment
+l'installation de la machine fr.fsf.org, première machine intégrée au
+projet GNU, présente hors des USA, chez Nevrax). L'activité
 quotidienne du chapitre français utilise des ressources GNU (DNS,
@@
 
+\item SourceForge - \url{http://SourceForge.net/}
+
 \item APRIL - \url{http://www.april.org/}

### 2001-06-06 fr/about/article-fsfeurope.tex
Actualisation + typo
svn path=/trunk/; revision=740

@@
-% $Id: article-fsfeurope.tex,v 1.5 2001-05-31 09:43:48 mad Exp $
-% $Revision: 1.5 $
+% $Id: article-fsfeurope.tex,v 1.6 2001-06-06 19:16:21 r4f Exp $
+% $Revision: 1.6 $
 
@@
 Libre. La FSF Europe s'engage actuellement à assister et coordonner le
-processus d'adaptation de la licence GNU GPL au droit français et
+processus d'adaptation de la licence GNU GPL aux droits français et
 allemand. De manière plus générale, elle met en place un réseau de
@@
 Allemagne, Italie et bientôt en Suisse, au Portugal, en Belgique et en
-Autriche nous pouvons agir concrètement au niveau Européen. Dans
+Autriche nous pouvons agir concrètement au niveau européen. Dans
 chaque pays des relations de travail se mettent en place avec les
 associations qui défendent depuis de nombreuses années le Logiciel
-Libre afin d'apporter une dimension Européenne à leur action. L'APRIL
+Libre afin d'apporter une dimension européenne à leur action. L'APRIL
 en France peut par exemple coordonner les efforts de traduction de la
@@
 
-\subsubsection*{Chronologie des actions du chapitre Français}
+\subsubsection*{Chronologie des actions du chapitre français}
 
-La FSF Europe a de fait commencé ses activités en France le 12 avril.
+La FSF Europe a, de fait, commencé ses activités en France le 12 avril 2001.
 Après la création des aspect administratifs (par exemple l'hébergement
@@
 l'installation de la machine fr.fsf.org, première machine intégrée au
-projet GNU, présente hors des USA, chez Nevrax). L'activité
+projet GNU présente hors des USA, chez Nevrax). L'activité
 quotidienne du chapitre français utilise des ressources GNU (DNS,
@@
 dédié au projet GNU et à tout projet Logiciel Libre. Grâce aux dons de
-membres de l'APRIL et de la FSF Europe, une nouvelle machine va être
+membres de l'APRIL et de la FSF Europe, une nouvelle machine a été
 installée pour Savannah. Les développeurs les plus actifs de Savannah
@@
 http://www.fsfeurope.org}. Vous pouvez également entrer en contact
-avec les personnes mentionnées ci-dessous ou envoyer un e-mail à toute
+avec les personnes mentionnées ci-dessous ou envoyer un courrier électronique à toute
 l'équipe de la FSF Europe à {\tt team@fsfeurope.org}.
 
-Vous pouvez également lire le channel de news de la FSF France à
+Vous pouvez également prendre des nouvelles de la FSF France à
 l'adresse \url{http://france.fsfeurope.org/news/}.
@@
 
-\item Free Software Foundation Europe Chapter France AKA FSF France -
+\item Free Software Foundation Europe-Chapter France alias FSF France -
 \url{http://france.fsfeurope.org/}

### 2001-06-27 fr/about/article-fsfeurope.tex
Modifs from Anne Possoz
svn path=/trunk/; revision=889

@@
-% $Id: article-fsfeurope.tex,v 1.6 2001-06-06 19:16:21 r4f Exp $
-% $Revision: 1.6 $
+% $Id: article-fsfeurope.tex,v 1.7 2001-06-27 08:25:23 mad Exp $
+% $Revision: 1.7 $
 
@@
 ont un chancelier (président) ainsi qu'un vice-chancelier
-(secrétaire) parlant la langue du pays. Ils sont disponibles pour la presse
-locale et les représentants politiques, et sont en contact permanent avec
-la communauté du Logiciel Libre du pays. Des organisations, auxquelles
+(secrétaire). Ils sont disponibles pour la presse locale et les
+représentants politiques, et sont en contact permanent avec les
+communautés du Logiciel Libre du pays. Des organisations, auxquelles
 tout le monde peut prendre part, pourront devenir organisations
@@
 Europe. Enfin, les volontaires pourront participer notamment via
-Savannah. De cette façon, chacun peut être impliqué dans le Mouvement
-du Logiciel Libre et la FSF Europe.
+Savannah (\url{http://savannah.gnu.org/}). De cette façon, chacun peut
+être impliqué dans le Mouvement du Logiciel Libre et la FSF Europe.
 
@@
 l'université de Soissons, la Technopole Logiciel Libre accompagnera
-les entreprises développant une activité commerciale basée sur les
+les entreprises développant une activité de services basée sur les
 Logiciels Libres.  En dispensant des formations initiant les chefs
@@
 
-\paragraph{Savannah et l'Europe}  Savannah est un clone de SourceForge
-dédié au projet GNU et à tout projet Logiciel Libre. Grâce aux dons de
-membres de l'APRIL et de la FSF Europe, une nouvelle machine a été
-installée pour Savannah. Les développeurs les plus actifs de Savannah
-sont en France et au Portugal. Nous avons également pris contact avec
-les responsables de deux projets similaires à Savannah (PeCoVall,
-Pure/Source) et discuté de la manière de partager les efforts et les
-ressources pour obtenir une plate-forme de développement coopérative
-encore plus puissante.
+\paragraph{Savannah et l'Europe}
+
+Savannah (\url{http://savannah.gnu.org/}) est un point central pour le
+développement, la distribution, la maintenance de logiciels
+libres. Savannah met à disposition, sous la forme d'un site Web, un
+ensemble d'outils tels que un environnement pour archivage, un système
+de gestion de versions, un gestionnaire de tâches, des listes de
+diffusion, un espace FTP.
+
+Savannah a été mis en place par le projet GNU, mais est utilisable par
+tout projet Logiciel Libre. Grâce aux dons de membres de l'APRIL et de
+la FSF Europe, une nouvelle machine a été installée pour Savannah. Les
+développeurs les plus actifs de Savannah sont en France et au
+Portugal. Nous avons également pris contact avec les responsables de
+deux projets similaires à Savannah (PeCoVall, Pure/Source) et discuté
+de la manière de partager les efforts et les ressources pour obtenir
+une plate-forme de développement coopérative encore plus puissante.
 

### 2001-07-17 about/Constitution.de.tex
Stuff for constitution
svn path=/trunk/; revision=1001

%% $Id: Constitution.de.tex,v 1.1 2001-07-17 13:16:16 taz Exp $
%% 
%% Constitution of the Free Software Foundation Europe
%%
%% Language: German
%%

%{{{ Header

\documentclass[a4wide,12pt]{scrartcl}
\usepackage{german,umlaut}
\usepackage{fancyheadings}
\usepackage{alltt}
\usepackage[german]{babel}

\selectlanguage{german}
\pagestyle{fancyplain}

\lhead[\fancyplain{}{\bfseries\thepage}]
        {\fancyplain{}{\bfseries\rightmark}}
\rhead[\fancyplain{}{\bfseries\leftmark}]
        {\fancyplain{}{\bfseries\thepage}}
\cfoot{}
\renewcommand{\sectionmark}[1]{\markright{\thesection\ #1}}
\renewcommand{\thesection}{\S~\arabic{section}}

%
% alternative ways of doing it...
%

% this makes the references work correctly:
%\newcommand{\A}{\subsection{}}
%\renewcommand{\subsectionmark}[1]{\markright{\thesection\ }}

% this looks better:
\newcounter{absatz}[section]
\newcommand{\A}{\par\vspace{1ex}
                \stepcounter{absatz}\noindent(\arabic{absatz})~~}

%}}}

\begin{document}

%{{{ 0 Präambel

\thispagestyle{empty}
\begin{center}
\Huge\bf Satzung
\end{center}

\section*{Präambel}

Angesichts des Eintritts in das digitale Zeitalter, in dem realer und
virtueller Raum gleichberechtigt nebeneinander die soziale, kulturelle
und wissenschaftliche Entwicklung der Menschheit prägen werden,
ergeben sich Fragen, deren Stellung und Bearbeitung langfristiges Ziel
der Free Software Foundation Europe sein wird.\\

Unmittelbare Aufgabe ist im Hinblick darauf die selbstlose Tätigkeit
zur Förderung von Freier Software (engl. {\glqq}Free Software{\grqq}) sowie der
Prägung und Verbreitung des Bewusstseins der mit ihr zusammenhängenden
philosophischen und gesellschaftlichen Fragen.\\

Damit tritt die FSF Europe als offizielle Schwesterorganisation der
durch Richard M. Stallman in den Vereinigten Staaten von Amerika
gegründeten Free Software Foundation an deren Seite. Diese in den USA
als gemeinnützig anerkannte Organisation widmet sich seit 1984 der
Förderung und Verbreitung von Free Software und insbesondere dem
GNU-System, einem Unix-artigen Betriebssystem auf Basis Freier
Software, dessen bekannteste Variante, GNU/Linux, mit wachsendem
Erfolg seit etwa 1993 auf vielen Rechnern im Einsatz ist.\\

Der Begriff Freier Software im Sinne der FSF Europe bezieht sich nicht auf den Preis, sondern vielmehr auf die folgenden vier Freiheiten:
\begin{enumerate}
\item Freiheit: Die Freiheit, ein Programm für jeden Zweck einsetzen
zu dürfen; 
\item Freiheit: Die Freiheit, untersuchen zu dürfen, wie ein Programm
funktioniert, und es den eigenen Bedürfnissen anzupassen; 
\item Freiheit: Die Freiheit, Kopien für andere machen zu dürfen; 
\item Freiheit: Die Freiheit, das Programm verbessern zu dürfen und
diese Verbesserungen zum allgemeinen Wohl zugänglich zu machen. 
\end{enumerate}

Diese Definition Freier Software geht ursprünglich zurück auf den
Grundgedanken des freien Austauschs von Wissen und Ideen, wie er
traditionell im Feld der Wissenschaft zu finden ist. Software ist, wie
auch Gedanken, nichtstofflich und verlustfrei kopierbar. Das
Weiterreichen dient einem evolutionären Prozess, über den Ideen und
Software weiterentwickelt werden.\\

Nur Freie Software bewahrt die Nachvollziehbarkeit und Möglichkeit der
Fortentwicklung wissenschaftlicher Ergebnisse. Sie ist daher im
wissenschaftlichen Diskurs die einzige Art von Software, die den
Idealen einer freien Wissenschaft entspricht. Dementsprechend
resultiert aus der Förderung Freier Software auch eine Förderung der
Forschung.\\

Die Verbreitung von Informationen und die Meinungsbildung wird
zunehmend durch digitale Medien geprägt, und es existieren Pläne zur
direkteren Einbindung der Bürger mit Hilfe digitaler Technologien in
die Demokratie. Eine zentrale Aufgabe der FSF Europe besteht daher in
der Heranbildung eines in diesen Belangen mündigen Bürgers und der
Förderung eines demokratischen Staatswesens.\\

Der digitale Raum mit Software als seinem Medium und seiner Sprache
besitzt ein gewaltiges Potential zur Förderung aller
geistig-kulturellen Belange der Menschheit. Durch die allgemeine
Verfügbarmachung und Offenlegung dieses Mediums gewährleistet Freie
Software Chancengleichheit sowie den Schutz der Privatsphäre.\\

Die Ausprägung des Bewusstseins in allen Bevölkerungsschichten für die
mit dem digitalen Zeitalter verbundenen Probleme ist langfristiges
Ziel und einer der Kernpunkte der Tätigkeit der FSF Europe.\\

Daher wird mit der FSF Europe die Förderung des verstärkten Einsatzes
von Freier Software in Schulen und Universitäten angestrebt, um
parallel mit der Prägung des Problembewusstseins für den realen Raum
während der Ausbildung auch die Prägung des Problembewusstseins und
Verständnisses für den virtuellen Raum zu vollziehen.\\

Freie Software garantiert nachvollziehbare Ergebnisse und
Entscheidungsprozesse in Wissenschaft und öffentlichem Leben sowie die
Rechte des Einzelnen auf Entfaltung der Persönlichkeit und
Meinungsfreiheit. Daher ist es Aufgabe der FSF Europe, Freie Software
in alle Gebiete zu tragen, die das öffentliche Leben oder die
{\glqq}informationellen Menschenrechte{\grqq} der Bürger berühren.\\

%}}}

%{{{ 1 Name, Sitz, Geschäftsjahr

\section{Name, Sitz, Geschäftsjahr}
\A
Der Verein führt den Namen {\glqq}Free Software Foundation Europe{\grqq}. Er
soll in das Vereinsregister eingetragen werden; nach der Eintragung
führt er den Zusatz {\glqq}e.V.{\grqq}.
\A
Der Verein hat seinen Sitz in Hamburg.
\A
Das Geschäftsjahr ist das Kalenderjahr.

%}}}

%{{{ 2 Zweck, Aufgaben, Gemeinnützigkeit

\section{Zweck, Aufgaben, Gemeinnützigkeit}
\A
Zweck der FSF Europe ist die Förderung der in der Präambel
aufgeführten Ziele. 
\A
Dem Zweck der FSF Europe sollen namentlich dienen:
\begin{enumerate}
\item die Unterstützung von staatlichen Stellen und privaten
Organisationen in allen Fragen der Freien Software, 
\item die Zusammenarbeit und Koordination mit den Landesvereinen, die
die gleichen Ziele verfolgen, 
\item die Förderung von Projekten, die Freie Software entwickeln, 
\item die Verbreitung der philosophischen Ideale von Freier Software 
\end{enumerate}
\A
Die FSF Europe verfolgt ausschließlich und unmittelbar gemeinnützige
und wissenschaftliche Zwecke im Sinne des Abschnitts
{\glqq}steuerbegünstigte Zwecke{\grqq} der Abgabenordnung. Die FSF Europe ist
selbstlos tätig und verfolgt keine eigenwirtschaftlichen Ziele. 
\A
Mittel der FSF Europe dürfen nur für die satzungsmäßigen Zwecke
verwandt werden. Es darf keine Person durch Ausgaben, die dem Zweck
der FSF Europe fremd sind oder durch unverhältnismäßige Vergütungen
begünstigt werden.

%}}}

%{{{ 3 Erwerb der Mitgliedschaft

\section{Erwerb der Mitgliedschaft}
\A
Mitglied der FSF Europe kann jede in- oder ausländische natürliche
oder juristische Person werden. Natürliche Personen müssen das
16. Lebensjahr vollendet haben. Minderjährige besitzen kein passives
Wahlrecht.  
\A
Voraussetzung für den Erwerb der Mitgliedschaft ist ein schriftlicher
Aufnahmeantrag an den Vorstand.  
\A
Die Mitgliederversammlung entscheidet über den Aufnahmeantrag mit drei
Vierteln sämtlicher Mitglieder, die natürliche Personen sind. Der
Vorstand kann dem Aufnahmeantrag vorübergehend stattgeben; der
Aufnahmeantrag muss dann bei der nächsten ordentlichen
Mitgliederversammlung bestätigt werden. Bei Ablehnung des Antrages
besteht keine Verpflichtung, dem Antragsteller die Gründe mitzuteilen.

%}}}

%{{{ 4 Beendigung der Mitgliedschaft

\section{Beendigung der Mitgliedschaft}
\A
Die Mitgliedschaft endet 
\renewcommand{\labelenumi}{\alph{enumi}.}
\begin{enumerate}
\item mit dem Tod des Mitglieds bei natürlichen Personen bzw. seiner
Liquidation bei juristischen Personen; 
\item durch Austritt aus dem Verein; 
\item durch Ausschluss aus dem Verein. 
\end{enumerate}
\renewcommand{\labelenumi}{\arabic{enumi}.}
\A
Der Austritt erfolgt durch schriftliche Erklärung gegenüber dem
Vorstand. Bei beschränkt Geschäftsfähigen ist die Austrittserklärung
auch von dem gesetzlichen Vertreter zu unterschreiben. Der Austritt
kann jederzeit erklärt werden.  
\A
Aus wichtigem Grund oder wenn das Vertrauensverhältnis zwischen den
Mitgliedern zerrüttet ist, kann ein Mitglied durch Beschluss des
Vorstandes aus dem Verein ausgeschlossen werden. Vor der
Beschlussfassung muss der Vorstand dem Mitglied Gelegenheit zur
schriftlichen Stellungnahme geben. Der Beschluss des Vorstandes ist
schriftlich zu begründen und dem Mitglied zuzusenden. Gegen den
Beschluss kann das Mitglied Berufung an die Mitgliederversammlung
einlegen. Die Berufung ist innerhalb von zwei Wochen nach Zugang des
Beschlusses beim Vorstand einzulegen. Der Vorstand hat binnen drei
Monaten nach fristgemäßer Einlegung der Berufung eine
Mitgliederversammlung einzuberufen, die den Beschluss des Vorstandes
mit einer Mehrheit von drei Vierteln der übrigen Mitglieder bestätigen
kann. Bis zur endgültigen Entscheidung über den Ausschluss bleibt das
Mitglied von allen Pflichten und Rechten suspendiert.

%}}}

%{{{ 5 Mitgliedsbeiträge

\section{Mitgliedsbeiträge}
Die Mitglieder leisten Beiträge durch ehrenamtliche Tätigkeit oder
durch die Ausübung eines Vereinsamtes.

%}}}

%{{{ 6 Gliederung des Vereins

\section{Gliederung des Vereins}
\A
Die Free Software Foundation Europe bildet eine europäische
Verbandsstruktur und gliedert sich in Landesvereine. Diese sind
Vereine mit eigener Rechtspersönlichkeit nach dem Vereinsrecht des
europäischen Staates für dessen Gebiet der Verein tätig ist. Alle
Mitglieder der Landesvereine sollen auch Mitglieder der FSF Europe
sein. 
\A
Die Landesvereine haben in ihren Satzungen die vom erweiterten
Vorstand zur Wahrung der Einheitlichkeit in der Free Software
Foundation Europe beschlossenen Mindesterfordernisse aufzunehmen, die
in einer Mustersatzung für die Landesvereine niedergelegt sind. Dies
gilt mit der Ausnahme solcher Erfordernisse, die nach dem Recht des
Staates, in dem der Landesverein gegründet wird, unzulässig sind. Die
Satzung ist in diesem Fall so auszugestalten, wie dies den Intentionen
der Mustersatzung am Besten entspricht. Die Satzung bedarf vor der
Aufnahme des Landesvereins in die FSF Europe der Genehmigung des
erweiterten Vorstands. 
\A
Die Finanzen der Landesvereine richtet sich nach der Finanzordnung
gem. \ref{finord} Abs. 2 dieser Satzung. 
\A
Die Landesvereine können im eigenen Namen, in Erledigung der örtlich
anfallenden Aufgaben, Verträge abschließen, sofern die Mittel zur
Erfüllung dieser Verträge vorhanden sind. Verpflichtungen für die FSF
Europe können sie in keinem Falle eingehen.
\A
Die Landesvereine können nur in Angelegenheiten, die sich im
Schwerpunkt auf das Gebiet des Staates beziehen, in dem sie tätig
werden, mit allen für ihr Gebiet zuständigen Behörden und
Organisationen in Verhandlungen treten. 
\A
Die Landesvereine sind berechtigt und verpflichtet zur Führung eines
eigenen Namens, aus dem die Zugehörigkeit zur Free Software Foundation
Europe zum Ausdruck kommt. Dies hat dadurch zu geschehen, dass dem
Namen {\glqq}Free Software Foundation Europe{\grqq} die Bezeichnung {\glqq}Chapter{\grqq}
mit der jeweiligen Staatsbezeichnung  in englischer Sprache anzuhängen
ist. Zusätzlich können sie den Namen in der Landessprache führen.  
\A
Der Vorstand ist berechtigt, einem Landesverein, der gegen diese
Satzung oder die eigene Satzung verstößt, das Recht zur Führung des
Namens {\glqq}FSF Europe{\grqq} zu entziehen. Gegen den Beschluss kann der
Mitgliedsverein Berufung an die Mitgliederversammlung einlegen. Die
Berufung ist innerhalb von zwei Wochen nach Zugang des Beschlusses
beim Vorstand einzulegen. Der Vorstand hat binnen drei Monaten nach
fristgemäßer Einlegung der Berufung eine Mitgliederversammlung
einzuberufen, die dem Beschluss des Vorstands mit einer Mehrheit von
drei Vierteln sämtlicher Mitglieder abhelfen kann.

%}}}

%{{{ 7 Organe des Vereins

\section{Organe des Vereins}
Die Free Software Foundation Europe bildet eine europäische
Verbandsstruktur. Organe der FSF Europe sind:  
\renewcommand{\labelenumi}{\alph{enumi})}
\begin{enumerate}
\item der Vorstand (Präsident);
\item der Vizepräsident;
\item der Verwaltungsleiter (Head of Office);
\item der erweiterte Vorstand;
\item die Mitgliederversammlung.
\end{enumerate}
\renewcommand{\labelenumi}{\arabic{enumi}.}

%}}}

%{{{ 8 Der Vorstand

\section{Der Vorstand}
Der Vorstand der FSF Europe besteht aus dem Präsidenten.

%}}}

%{{{ 9 Die Zuständigkeit des Vorstands

\section{Die Zuständigkeit des Vorstands}
\A
Der Vorstand ist für alle Angelegenheiten der FSF Europe zuständig,
soweit sie nicht durch Satzung einem anderen Organ der FSF Europe
übertragen sind. Er hat insbesondere folgende Aufgaben: 
\renewcommand{\labelenumi}{\alph{enumi})}
\begin{enumerate}
\item Einberufung und Vorbereitung der Mitgliederversammlung sowie
Aufstellung der Tagesordnung; 
\item Erstellung des Jahresberichts;
\item Beschlussfassung über die Aufnahme und den Ausschluss von
Mitgliedern; 
\item Unterstützung und Kontrolle der Landesvereine; 
\item Vertretung in gerichtlichen und außergerichtlichen
Angelegenheiten, insbesondere auch der Behördenverkehr; 
\item Beziehungen zur Presse.
\end{enumerate}
\renewcommand{\labelenumi}{\arabic{enumi}.}
\A\label{finord}
Der Vorstand führt die Finanzen nach der Finanzordnung, die die
Mitgliederversammlung mit einer Mehrheit von drei Viertel aller
Stimmen beschließt. Die Finanzordnung muss vorgeben,  
\begin{enumerate}
\item dass etwaige Gewinne nur für die satzungsmäßigen Zwecke verwandt
  werden dürfen.  
\item dass die Mitglieder keine Gewinnanteile und sonstige Zuwendungen
aus Mitteln der FSF Europe oder ihrer Landesvereine erhalten
dürfen. Dies gilt auch beim Ausscheiden von Mitgliedern.  
\item dass Verwaltungsausgaben, die dem Zweck der FSF Europe fremd
sind, nicht bewilligt werden dürfen. Dies gilt auch für
unverhältnismäßig hohe Vergütungen. 
\end{enumerate}
\A
Der Vorstand haftet gegenüber dem Verein nicht für leicht fahrlässiges
Verhalten.

%}}}

%{{{ 10 Wahl und Amtsdauer des Vorstandes

\section{Wahl und Amtsdauer des Vorstandes}

\A
Der Vorstand wird von der Mitgliederversammlung für die Dauer von zwei
Jahren, gerechnet von der Wahl an, gewählt. Er bleibt jedoch bis zur
Neuwahl des Vorstandes im Amt.  
\A
Wählbar sind nur Vereinsmitglieder. Scheidet der Vorstand vorzeitig
aus, vertritt der Vizepräsident den Verein bis zur Wahl eines neuen
Vorstandes. Dazu ist eine außerordentliche Mitgliederversammlung
innerhalb von drei Monaten durch den Vizepräsidenten einzuberufen.

%}}}

%{{{ 11 Der Vizepräsident

\section{Der Vizepräsident}
\A
Der Vizepräsident vertritt den Vorstand in folgenden Fällen:
\begin{enumerate}
\item Ausscheiden des Vorstands;
\item vorübergehende Unabkömmlichkeit des Vorstandes.
\end{enumerate}
\A
Der Vorstand ist unabkömmlich, wenn er dies dem Vizepräsidenten
schriftlich mitteilt. Der Vizepräsident führt die Geschäfte, solange
und in dem Umfang, wie ihm dies vom Vorstand schriftlich aufgetragen
wurde. Der Vorstand gilt als unabkömmlich, wenn er mehr als sieben
Tage nicht erreichbar ist oder wegen Krankheit sein Amt nicht ausüben
kann. 
\A
Der Vizepräsident wird von der Mitgliederversammlung für die Dauer von
zwei Jahren, gerechnet von der Wahl an, gewählt. Er bleibt jedoch bis
zur Neuwahl des Vizepräsidenten im Amt. Zum Vizepräsidenten können nur
Mitglieder der FSF Europe gewählt werden. Mit der Beendigung der
Mitgliedschaft im Verein endet auch das Amt des Vizepräsidenten.  
\A
Scheidet der Vizepräsident vorzeitig aus, so kann der Vorstand für die
restliche Amtsdauer des Ausgeschiedenen einen Nachfolger wählen.  
\A
Der Vizepräsident haftet gegenüber dem Verein nicht für leicht
fahrlässiges Verhalten.

%}}}

%{{{ 12 Der Verwaltungsleiter

\section{Der Verwaltungsleiter}
\A
Der Verwaltungsleiter führt das Büro der FSF Europe. Er vertritt den
Verein in dem damit zusammengehörenden Geschäftsbereich.  
\A
Der Verwaltungsleiter zertifiziert die Schlüssel zur Verschlüsselung
von digitalen Texten und Dokumenten. 
\A
Der Verwaltungsleiter wird von der Mitgliederversammlung für die Dauer
von zwei Jahren, gerechnet von der Wahl an, gewählt. Er bleibt jedoch
bis zur Neuwahl des Verwaltungsleiters im Amt. Zum Verwaltungsleiter
können nur Mitglieder der FSF Europe gewählt werden. Mit der
Beendigung der Mitgliedschaft im Verein endet auch das Amt des
Verwaltungsleiters.  
\A
Scheidet der Verwaltungsleiter vorzeitig aus, so kann der Vorstand für
die restliche Amtsdauer des Ausgeschiedenen einen Nachfolger wählen.  
\A
Der Verwaltungsleiter haftet gegenüber dem Verein nicht für leicht
fahrlässiges Verhalten.

%}}}

%{{{ 13 Der erweiterte Vorstand

\section{Der erweiterte Vorstand}
\A
Der erweiterte Vorstand besteht aus dem Präsidenten, dem
Vizepräsidenten und dem Verwaltungsleiter.  
\A
Der erweiterte Vorstand ist beschlussfähig, wenn mindestens zwei
Mitglieder anwesend sind, darunter der Präsident. Beschlüsse werden
mit Stimmenmehrheit gefasst.  
\A
Die Mitglieder des erweiterten Vorstandes haften gegenüber dem Verein
nicht für leicht fahrlässiges Verhalten.

%}}}

%{{{ 14 Die Zuständigkeit des erweiterten Vorstandes

\section{Die Zuständigkeit des erweiterten Vorstandes}
Der erweiterte Vorstand ist folgende Aufgaben zuständig:
\begin{enumerate}
\item die Genehmigung der Satzungen der Landesvereine; 
\item die Ausführung der {\glqq}Guidelines{\grqq}, die von der
Mitgliederversammlung beschlossen werden. 
\end{enumerate}

%}}}

%{{{ 15 Die Mitgliederversammlung

\section{Die Mitgliederversammlung}\label{sec:mv}
\A
In der Mitgliederversammlung hat jedes Mitglied, das eine natürliche
Person ist, eine Stimme. Zur Ausübung des Stimmrechts kann ein anderes
Mitglied durch schriftliche Mitteilung an den Vorstand bevollmächtigt
werden. Die Bevollmächtigung ist für jede Mitgliederversammlung
gesondert zu erteilen. Ein Mitglied kann nicht für mehr als ein
Drittel aller Mitglieder das Stimmrecht ausüben. 
\A
Die Mitgliederversammlung ist ausschließlich für folgende
Angelegenheiten zuständig: 
\begin{enumerate}
\item Erstellung von sog. {\glqq}Guidelines{\grqq}, die die Tätigkeit des
Vorstandes bzw. des erweiterten Vorstands leiten sollen; 
\item Entgegennahme des Jahresberichts des Vorstands;
\item Entlastung der Organe;
\item Wahl und Abberufung des Vorstands, des Vizepräsidenten und des
Verwaltungsleiters. 
\end{enumerate}

%}}}

%{{{ 16 Die Einberufung der Mitgliederversammlung

\section{Die Einberufung der Mitgliederversammlung}\label{sec:edmv}
\A
Mindestens einmal im Jahr, möglichst im ersten Quartal, soll die
ordentliche Mitgliederversammlung möglichst in einem der Staaten
stattfinden, in denen ein Landesverein existiert. Sie wird vom
Vorstand unter Einhaltung einer Frist von sechs Wochen schriftlich
unter Angabe der Tagesordnung einberufen. Die Frist beginnt mit dem
auf die Absendung des Einladungsschreibens folgenden Tag. Das
Einladungsschreiben gilt dem Mitglied als zugegangen, wenn es an die
letzte vom Mitglied dem Verein schriftlich bekannt gegebene Adresse
oder E-Mail-Adresse gerichtet ist. Mit schriftlicher Zustimmung von
drei Vierteln Mitgliedern kann die Einladungsfrist auf drei Wochen
verkürzt werden. 
\A
Die Tagesordnung setzt der Vorstand fest. Jedes Mitglied kann bis
spätestens einer Woche vor einer Mitgliederversammlung beim Vorstand
schriftlich eine Ergänzung der Tagesordnung beantragen. Der Vorstand
hat zu Beginn der Mitgliederversammlung die Ergänzung bekannt zu
geben. Über Anträge auf Ergänzung der Tagesordnung, die in
Mitgliederversammlung gestellt werden, beschließt die Versammlung.

%}}}

%{{{ 17 Die außerordentliche Mitgliederversammlung

\section{Die außerordentliche Mitgliederversammlung}
Der Vorstand kann jederzeit eine außerordentliche
Mitgliederversammlung einberufen. Diese muss einberufen werden, wenn
die Einberufung von mindestens einem Viertel aller Mitglieder
schriftlich unter Angabe des Zwecks und der Gründe vom Vorstand
verlangt wird. Für die außerordentliche Mitgliederversammlung
gelten \ref{sec:mv}, \ref{sec:edmv} und \ref{sec:bfmv} entsprechend.

%}}}

%{{{ 18 Beschlussfassung der Mitgliederversammlung

\section{Beschlussfassung der Mitgliederversammlung}\label{sec:bfmv}
\A
Die Mitgliederversammlung wird vom Vorstand, bei dessen Verhinderung
vom Vizepräsidenten und bei dessen Verhinderung vom Verwaltungsleiter
geleitet.  
\A
Die Art der Abstimmung bestimmt der Versammlungsleiter. Die Abstimmung
muss schriftlich durchgeführt werden, wenn ein Drittel der bei der
Abstimmung anwesenden stimmberechtigten Mitglieder dies beantragt. 
\A
Die Mitgliederversammlung ist nicht öffentlich.  
\A
Die Mitgliederversammlung ist beschlussfähig, wenn sie ordnungsgemäß
einberufen wurde und mindestens ein Drittel sämtlicher
Vereinsmitglieder anwesend ist oder durch anwesende Vereinsmitglieder
vertreten wird. Bei Beschlussunfähigkeit ist der Vorstand
verpflichtet, innerhalb von vier Wochen eine zweite
Mitgliederversammlung mit der gleichen Tagesordnung einzuberufen;
diese ist ohne Rücksicht auf die Zahl der erschienenen Mitglieder
beschlussfähig. Hierauf ist in der Einladung hinzuweisen.  
\A
Soweit nicht in der Satzung etwas anderes bestimmt ist, fasst die
Mitgliederversammlung Beschlüsse mit einfacher Mehrheit der
abgegebenen gültigen Stimmen; Stimmenenthaltungen bleiben daher außer
Betracht. Zur Änderung der Satzung ist jedoch eine Mehrheit von drei
Vierteln der (abgegebenen gültigen) Stimmen sämtlicher Mitglieder, zur
Auflösung der FSF Europe eine solche von vier Fünfteln der Stimmen
aller Mitglieder erforderlich. Eine Änderung des Zwecks der FSF Europe
kann nur mit Zustimmung aller Mitglieder beschlossen werden. Die
schriftliche Zustimmung der in der Mitgliederversammlung nicht
erschienenen Mitglieder kann nur innerhalb eines Monats gegenüber dem
Vorstand erklärt werden.  
\A
Bei Wahlen ist gewählt, wer mehr als die Hälfte der abgegebenen
gültigen Stimmen erhalten hat. Hat niemand mehr als die Hälfte der
abgegebenen gültigen Stimmen erhalten, so findet zwischen den beiden
Kandidaten, die die meisten Stimmen erhalten haben, eine Stichwahl
statt. Gewählt ist dann derjenige, der die meisten Stimmen erhalten
hat. Bei gleicher Stimmenzahl ist einmal eine neue Wahl erforderlich;
besteht die Stimmgleichheit fort, entscheidet das Los. 
\A
Über Beschlüsse der Mitgliederversammlung ist ein Protokoll
aufzunehmen, dass vom Protokollführer und dem Versammlungsleiter zu
unterzeichnen ist. Es soll folgende Feststellungen enthalten: Ort und
Zeit der Versammlung, die Person des Versammlungsleiters und des
Protokollführers, die Zahl der erschienenen Mitglieder, die
Tagesordnung, die einzelnen Abstimmungsergebnisse und die Art der
Abstimmung. Bei Satzungsänderungen soll der genaue Wortlaut angegeben
werden.

%}}}

%{{{ 19 Auflösung des Vereins

\section{Auflösung des Vereins}
\A
Die Auflösung der FSF Europe kann nur in einer Mitgliederversammlung
mit der in \ref{sec:bfmv} festgelegten Stimmenmehrheit beschlossen werden.
\A
Falls die Mitgliederversammlung nichts Anderes beschließt, ist der
Vorsitzende der vertretungsberechtigte Liquidator.  
\A
Das nach Beendigung der Liquidation vorhandene Vermögen fällt an die
\begin{alltt}
  Free Software Foundation
  59 Temple Place - Suite 330, 
  Boston, MA 02111-1307, USA, 
\end{alltt}
die es unmittelbar und ausschließlich für gemeinnützige Zwecke zu
verwenden hat.
\A
Die vorstehenden Bestimmungen gelten entsprechend, wenn die FSF Europe
aus einem anderen Grund aufgelöst wird oder seine Rechtsfähigkeit
verliert.

%}}}

%{{{ 20 Schriftform

\section{Schriftform}
Die Schriftform ist eingehalten, wenn eine der folgenden
Voraussetzungen erfüllt ist: 
\renewcommand{\labelenumi}{\alph{enumi})}
\begin{enumerate}
\item handschriftlich unterzeichnetes Papierdokument;
\item E-Mail, die mit einem dem Stand der Technik genügenden Schlüssel
verschlüsselt ist. Der Verein entscheidet, was als Stand der Technik
anzusehen ist und der Schlüssel muss von dem Verein zertifiziert
sein. 
\end{enumerate}
\renewcommand{\labelenumi}{\arabic{enumi}.}

%}}}

%{{{ 21 Gerichtsstand

\section{Gerichtsstand}
Gerichtsstand für alle aus dieser Satzung sich ergebenden Rechte und
Pflichten ist Hamburg.

%}}}

\end{document}

### 2001-07-17 about/Constitution.en.tex
Stuff for constitution
svn path=/trunk/; revision=1001

%% $Id: Constitution.en.tex,v 1.1 2001-07-17 13:16:16 taz Exp $
%% 
%% Constitution of the Free Software Foundation Europe
%%
%% Language: English
%%

%{{{ Header

\documentclass[a4wide,12pt]{article}
\usepackage{german,umlaut}
\usepackage{fancyheadings}
\usepackage{alltt}
\usepackage[german]{babel}

\selectlanguage{german}
\pagestyle{fancyplain}

\lhead[\fancyplain{}{\bfseries\thepage}]
        {\fancyplain{}{\bfseries\rightmark}}
\rhead[\fancyplain{}{\bfseries\leftmark}]
        {\fancyplain{}{\bfseries\thepage}}
\cfoot{}
\renewcommand{\sectionmark}[1]{\markright{\thesection\ #1}}
\renewcommand{\thesection}{\S~\arabic{section}}

%
% alternative ways of doing it...
%

% this makes the references work correctly:
%\newcommand{\A}{\subsection{}}
%\renewcommand{\subsectionmark}[1]{\markright{\thesection\ }}

% this looks better:
\newcounter{absatz}[section]
\newcommand{\A}{\par\vspace{1ex}
                \stepcounter{absatz}\noindent(\arabic{absatz})~~}

%}}}

\begin{document}

%{{{ 0 Preamble

\thispagestyle{empty}
\begin{center}
\Huge\bf Constitution
\end{center}

\section*{Preamble}

Upon entering the digital age, in which real and virtual space will
equally determine the social, cultural and scientific development of
mankind, the Free Software Foundation Europe has the long-term goal to
raise and work on the questions this will necessarily raise.\\

In this regard the direct function is the unselfish promotion of
Free Software as well as creating and propagating the awareness of the
related philosophical and social questions.\\

As its acknowledged sister organization, the FSF Europe will join
forces with the Free Software Foundation founded by Richard
M. Stallman in the United States of America. The latter, recognized
tax-exempt charitable organization in the USA, has been dedicating
itself since 1984 to the promotion and distribution of Free software
and in particular the GNU-System, a Unix-like operating system. This
system is mostly known by one of its variants, GNU/Linux, which since
1993 has been used successfully on many computers.\\

The term Free Software in the sense of the FSF Europe does not refer
to the price, but rather to the following four freedoms:
\begin{enumerate}
\item freedom: the freedom to use a program for any purpose
\item freedom: the freedom to study the program and adapt it to your
own needs. 
\item freedom: the freedom to make copies for others.
\item freedom: the freedom to improve a program and make these
improvements available to others, so that the whole community
benefits. 
\end{enumerate}

This definition of Free Software goes back to the idea of freely
exchanging knowledge and ideas that can traditionally be found in
scientific fields. Like thoughts, software is non-tangible and
duplicable without loss. Passing feeds an evolutionary process,
advancing thoughts and software.\\

Only Free Software preserves the possibility to comprehend and build
upon scientific results. For scientists, it is the only kind of
software which corresponds to the ideals of a free
science. Accordingly, the promotion of free software is also a
promotion of science.\\

The distribution of information and the forming of an opinion are done
increasingly by digital media, and the trend is to foster the use of
those means for a direct citizen participation to
democracy. Therefore, a central task of the FSF Europe is to train
proficient citizens in these media, thereby promoting democracy.\\

Digital space (``Cyberspace''), with software as its medium and its
language has an enormous potential for the promotion of all mental and
cultural aspects of mankind. By making it commonly available and
opening up the medium, Free Software grants equal chances and
protection of privacy.\\

Coining the awareness for the problems related to the digital age in
all parts of society is long-term goal and a core aspect of the work
of the FSF Europe.\\

Therefore the FSF Europe will seek to increase the use of Free
Software in schools and universities in order to parallelize the
education in real space matters with the creation of understanding and
awareness of problems in virtual space.\\

Free Software guarantees traceable results and decision-making
processes in science and public life as well as the individual rights
to free development of personality and liberty of opinion. It is the
job of the FSF Europe to carry Free Software into all areas that touch
public life or ``informational human rights'' of citizens.\\

%}}}

%{{{ 1 Name, seat, financial year

\section{Name, seat, financial year} 
\A 
The association bears the name ``Free Software Foundation Europe.''  It
is to be registered into the register of associations; after the
registration it leads the additive ``e.V..''
\A 
The association has its seat in Hamburg.  
\A
The financial year is the calendar year.

%}}}

%{{{ 2 Purpose, tasks, non-profit character

\section{Purpose, tasks, non-profit character}
\A 
Purpose of the FSF Europe is the promotion of the goals specified in
the preamble.
\A
The goals of the FSF Europe are namely to be achieved by:
\begin{enumerate}
\item the support of governmental and private organisations in all
aspects of the Free Software, 
\item the cooperation and coordination of the national associations
which pursue the same goals 
\item the support of projects working on Free Software,
\item the distribution of the philosophical ideals of Free Software  
\end{enumerate}
\A 
The FSF Europe pursues exclusively and directly publicly-spirited
and scientific purposes in the sense of the section ``tax-privileged
purposes'' of the tax code.  The FSF Europe is working selflessly and
does not pursue self-economic goals.

\A 
Means of the FSF Europe may be used for the statutory purposes only.
No person may be favored by expenses alien to the goals of the FSF
Europe or disproportionally high.

%}}}

%{{{ 3 Acquisition of membership

\section{Acquisition of membership}

\A 
Any national or foreign natural or legal person may become member of
the FSF Europe. Natural persons must be 16 years old.  Persons under
age do not have eligibility for election.
\A 
Condition for the acquisition of the membership is a written
application for membership to the executive committee.  
\A 
The general assembly of the members decides on the application
for membership with three quarters of all members, that are natural
people.  The executive committee can grant the application for
membership passing;  the application for membership must then be
confirmed by the next members assembly. In case of refusal of the
request no obligation exists to communicate the reasons to the
applicant.

%}}}

%{{{ 4 End of membership

\section{End of membership}
\A 
Membership ends 
\renewcommand{\labelenumi}{\alph{enumi}.}
\begin{enumerate}
\item with the death of the member with natural persons and/or its
liquidation in the case of legal persons;
\item by withdrawal from the association;
\item by exclusion from the association.  
\end{enumerate}
\renewcommand{\labelenumi}{\arabic{enumi}.}
\A
The withdrawal is made via written declaration vis-a-vis the executive
committee. The declaration withdrawal must be signed by the competent
legal representative. The withdrawal can be declared at any time.
\A
For important reasons or if the bond of trust between the members is
broken, a member can be excluded by decision of the executive
committee from the association. Before deciding on the matter, the
executive committee must give the member opportunity to state its
position in writing. The decision of the executive committee is to be
justified in writing and sent to the member. The member can appeal the
decision at the general assembly of the members.  The appeal must be
lodged within two weeks after communication of the decision at the
executive committee. The executive committee has to call in a general
assembly within three months of punctual insertion of the appointment,
which can support the decision of the executive committee with a
majority of three quarters of the remaining members.  Until the final
decision about the exclusion, the member remains suspended of all
obligations and all rights.

%}}}

%{{{ 5 Membership fees

\section{Membership fees}
The members make contributions by ways of honorary activity or through
holding a position in the association.

%}}}

%{{{ 6 Structure of the association

\section{Structure of the association}
\A 
The Free Software Foundation Europe forms a European federation
structure and is divided into national associations. Those are
associations with own juridical personality after the law on societies
of the European State for its area the association are active.  All
members of the national associations must also be members of the FSF
Europe.  
\A 
In order to preserve the uniformity, the national associations have to
fulfill minimum requirements determined by the extended executive
committee, which are written down in a constitution-template for
national associations. This applies with the exception of requirements
that are inadmissible according to the laws of the state in which the
national association is to be created. In this case the constitution
is to be modified so it reflects the intentions of the
constitution-template most closely. The constitution requires the
acceptance of the extended executive committee before the national
association becomes part of the FSF Europe.
\A
The finances of the national associations are determined by the
financial order in accordance with \S~9 sec. 2 of this
statute.
\A 
The national associations can conclude contracts in their own name for
the fulfillment of local tasks, if the means are present for the
fulfilment of these contracts.  They can in no case enter obligations
for the FSF Europe.
\A
The national associations can only enter negotiations with authorities
and organizations mainly active in their respective areas.
\A
The national associations are entitled and obliged to bear their own
name trough which the affiliation with the Free Software Foundation
Europe is expressed. This has to be done by adding the denomination
``Chapter'' with the name of the state in English to ``Free Software
Foundation Europe.'' Additionally they can bear the name in the
national language.
\A
The executive committee is entitled to extract the right for the
guidance of the name ``FSF Europe'' from a national association that
violates this constitution or its own constitution. The national
association can appeal the decision at the general assembly of the
members.  The appeal must be lodged within two weeks after
communication of the decision at the executive committee. The
executive committee has to call in a general assembly within three
months of punctual insertion of the appointment, which can remedy or
overrule the decision of the executive committee with a majority of
three quarters of the remaining members.  Until the final decision
about the exclusion, the member remains suspended of all obligations
and all rights.

%}}}

%{{{ 7 Bodies of the association

\section{Bodies of the association}
The Free software Foundation Europe forms a European federation
composed by the association of 7 organs/bodies. Organs of the FSF
Europe are:
\renewcommand{\labelenumi}{\alph{enumi})}
\begin{enumerate}
\item the executive committee (President);
\item the Vice-President;
\item the administrative director (Head of Office);
\item the extended executive committee;
\item the general assembly.
\end{enumerate}
\renewcommand{\labelenumi}{\arabic{enumi}.}

%}}}

%{{{ 8 The executive committee

\section{The executive committee}
The executive committee of the FSF Europe consists of the President.

%}}}

%{{{ 9 The competence of the executive committee

\section{The competence of the executive committee}
\A
The executive committee is responsibly for all affairs the FSF Europe,
as far as they are not transferred to another organ of the FSF Europe
by statute.  It has in particular the following tasks:
\begin{enumerate}
\item Convening and preparation of the general assembly of the members
as well as list the agenda;
\item Creation of the annual report;
\item Adoption of resolutions over the admission and the exclusion of
members;
\item Support and control of the national associations;
\item Agency in judicial and except-court affairs, in particular also
authority traffic; 
\item Relationships with the press.
\end{enumerate}
\A
The executive committee leads the finances after the finance plan,
which the general assembly with a majority of three quarters of all
voices decides on.  The finance plan must give,
\begin{enumerate}
\item that possible profits are used only for the statutory purposes
\item that no member may get no shares of the profits or other
allowances from the means of the FSF Europe or its local
associations. This also applies to separating members.
\item that administrative costs, which are alien to the purpose of the
FSF Europe may not be granted. Same applies to disproportionately high
payments.
\end{enumerate}
\A
The executive committee is not liable vis-a-vis the association
for slightly negligent behaviour.

%}}}

%{{{ 10 Election and term of office of the executive committee

\section{Election and term of office of the executive committee}
\A
The executive committee is elected by the general assembly of the
members for the duration of two years, from the election on. It
continues to be in office up to the new election of the executive
committee, however.
\A
Electable are only members of the FSF Europe. If the executive
committee separates prematurely, the vice-president represents the
association up to the selection of a new executive
committee. Additionally, an extraordinary general assembly of the
members is to be called in within three months by the vice-president.

%}}}

%{{{ 11 The Vice-President

\section{The Vice-President}
\A
The vice-president represents the executive comittee in the following
cases:
\begin{enumerate}
\item Separating of the executive committee;
\item Passing indispensability of the executive committee.
\end{enumerate}
\A 
The executive committee is indispensable, if he communicates this in
written form to the vice-president. The vice-president takes care of
all matters as long and to the extent they were transferred to him in
writing by the executive committee. The executive committee is
considered indispensable, if it is not attainable or cannot exercise
its office more than seven days because of illness.
\A
The vice-president is elected by the general assembly for the duration
of two years, from the election on.  He remains in office up to the
election of the new vice-president, however. Only members of the FSF
Europe can be elected for vice-president.  With the end of membership
in the association, the office of the vice-president also ends.
\A
If the vice-president is ruled out prematurely, the executive
committee can select a successor for the remaining term of office.
\A
The vice-president is not liable vis-a-vis the association for
slightly negligent behaviour.

%}}}

%{{{ 12 The Head of Office

\section{The Head of Office}
\A
The Head of Office leads the office of the FSF Europe. He represents
the association in the business area belonging to it.
\A
The Head of Office certifies the keys for encoding digital texts and
documents. 
\A
The Head of Office is elected by the general assembly for the duration
of two years, from election on. He remains in office up to the
election of the new Head of Office, however. Only members of the FSF
Europe can be selected for Head of Office. With the end of membership
in the association, the office of the administration director also
ends.
\A
If the administration director is ruled out prematurely, the executive
committee can select a successor for the remaining term of office.
\A
The administration director is not liable vis-a-vis the association
for slightly negligent behaviour.

%}}}

%{{{ 13 The extended executive committee

\section{The extended executive committee}
\A
The extended executive committee consists of the President, the
Vice-President and the Head of Office.  
\A
The extended committee is resolutionable, if at least two members are
present in person, among them the President. Decisions are passed by
single majority of votes.
\A
Members of the extended executive committee are not liable vis-a-vis
the association for slightly negligent behaviour.

%}}}

%{{{ 14 The competence of the extended executive committee

\section{The competence of the extended executive committee}
The extended executive committee is responsible for the following
tasks:
\begin{enumerate}
\item The approval of the statutes of the national associations;
\item The execution of the ``Guidelines,'' which are decided on by the
general assembly.
\end{enumerate}

%}}}

%{{{ 15 The general assembly

\section{The general assembly}\label{sec:ga}
\A 
In the general assembly, each member that is a natural person, has a
voice. For the practice of the right to vote another member can be
authorized by written message to the executive committee. The
authorisation is to be given for each general assembly separately.  A
member cannot practise the right to vote for more than one third of
all members.
\A
The general assembly has exclusive jurisdiction for the following
affairs:
\begin{enumerate}
\item Creation of so-called ``Guidelines,'' that are to lead the
activity of the executive committee and extended executive committee;
\item Receipt of the annual report of the executive committee;
\item Exoneration of the organs;
\item Choice and recall of the executive committee, the Vice-President
and of the Head of Office.
\end{enumerate}

%}}}

%{{{ 16 The convening of the general assembly

\section{The convening of the general assembly}\label{sec:cga}
\A 
At least once a year, if possible in the first quarter, is the orderly
general assembly to take place; if possible in one of the states, in
which a national association exists. It is called up in writing by the
executive committee under adherence to one period of six weeks under
indication of the agenda. The period begins with day following the
sending of the invitation letter. The invitation letter is to be
considered delivered to the member, if it is addressed to the last
address or e-mail address given by the member in writing. With written
agreement of three quarters of the members, the invitation period can
be shortened to three weeks.
\A
The agenda is derermined by the executive committee. Each member can
request a supplement of the agenda in writing to the executive
committee until at the latest one week before a general assembly. The
executive committee has to announce the supplement at the beginning of
the general assembly. About applications for supplement of the agenda,
which are made at the general assembly, are decided by the general
assembly.

%}}}

%{{{ 17 The extraordinary general assembly

\section{The extraordinary general assembly}
The executive committee can call up an extraordinary general assembly
at any time.  It must be called up, if at least a quarter of all
members require it in writing under indication of the purpose and
reasons to the executive committee. To the extraordinary meeting of
the members the \ref{sec:ga}, \ref{sec:cga} and \ref{sec:aorga} apply
accordingly.

%}}}

%{{{ 18 Adoption of resolutions of the general assembly

\section{Adoption of resolutions of the general assembly}\label{sec:aorga}
\A
The general assembly is led by the executive committee, at his
indespensability by the Vice-President and at the indispensability of
the latter by the Head of Office.
\A
The type of election is determined by the assembly director. The
election must be carried out in writing, if a third of the members
present in person with the election requests this.
\A
The general assembly is not public.
\A
The general assembly is resolutionable, if it was duly called up and
at least one third of all club members is present or represented by
club members present. In the case of decision inability, the executive
committee is obliged to call up a second general assembly with the
same agenda within four weeks; this general assembly will
resolutionable without consideration of the members present. This is
to be referred to in the invitation.
\A
Unless stated otherwise in the statute, the general assembly passes
decisions with simple majority of the delivered valid voices;
abstentions are therefore left out of the consideration. Changes of
the statute require a majority of three quarters of (the delivered
valid) voices, dissolution of the FSF Europe requires four fifths of
the voices of all members. Changing the purpose of the FSF Europe can
only be decided with unanimous agreement of all members. The written
agreement of the members not present in the general assembly can be
explained only within one month vis-a-vis the executive committee.
\A
In the case of elections it is selected who received more than half of
the delivered valid voices.  If nobody received more than half of the
delivered valid voices, a ballot takes place between the two
candidates, who received most of the voices.  Is selected then who
received most of the voices.  In the case of equal number of votes
once a new choice is necessary; if the mood equality continues to
exist, the lot decides.
\A
Over decisions of the general assembly, a protocol is to be led, that
is to be signed by the recording clerk and the assembly director.  It
is to contain the following ascertainments: Place and time of the
assembly, the person of the assembly director and the recording clerk,
the number of members present, the agenda, the individual election
results and the type of the election.  For amendments of the statute
the exact wording is to be given.

%}}}

%{{{ 19 Dissolution of the association

\section{Dissolution of the association}
\A
The dissolution of the FSF Europe can only be decided in a general
assembly by the majority in accordance to \ref{sec:aorga}.
\A
If the general assembly decides nothing else, the executive committee
is the liquidator entitled to act as substitute.
\A
After end of liquidation, existing funds go to the 
\begin{alltt}
  Free Software Foundation
  59 Temple Place - Suite 330, 
  Boston, MA 02111-1307, USA, 
\end{alltt}
which it has to use directly and exclusively for non-profit purposes.
\A 
The preceding regulations apply accordingly, if the FSF Europe is
dissolved for another reason or loses its legal capacity.

%}}}

%{{{ 20 Written form

\section{Written form}
Writing form is fulfilled if one of the following conditions is met:
\renewcommand{\labelenumi}{\alph{enumi})}
\begin{enumerate}
\item handwritten signed paper document;
\item E-mail signed with a key that is sufficiently state of the art.
The association decides what is to be regarded as state of the art and
the key must be certified by the association.
\end{enumerate}
\renewcommand{\labelenumi}{\arabic{enumi}.}

%}}}

%{{{ 21 Place of jurisdiction

\section{Place of jurisdiction}
Place of jurisdiction for all rights and duties resulting from this
statute is Hamburg, Germany.

%}}}

\end{document}

### 2001-10-03 drafts/IST/PartA.tex
first release
svn path=/trunk/; revision=1457

\documentclass[a4paper,11pt]{report}
\usepackage[latin1]{inputenc}
\begin{document}
\section*{Notice}
This file will be generated by Protool. Thus layout does not matter. It will
juste be cut\&paste.
\section*{A1 Administrative Overview}
\subsection*{Summary}
\begin{tabular}{|c|c|}
\hline
Proposal Acronym &  \\ \hline
Proposal Number &  \\ \hline
Thematic Priorities & IST-2001-4.3.3. \\ \hline
Call identifier & \\ \hline
Proposal Date & 16 10 2001 \\ \hline
Proposal full name & \\ \hline
\end{tabular}
\subsection*{Abstract (1000 characters)}
\textit{The proposal abstract should be a very short and precise presentation
of the main features of the proposal. Why is it proposed and what problem is it
solving? What are the objectives? How will the objectives be achieved? What
results are expected? This proposal abstract will be used together with the
proposal summary description in form A2 in the evaluation process and in
communications about the proposals to the interested parties (Commission
services and programme committees, etc.). Please use plain typed text, avoiding
formulae and other special characters. If the proposal is written in a language
other than English, please include in form A1 an English version of the
abstract.} \\
\\ \\
to be written ...
\begin{tabular}{|c|c|}
\hline
Duration in months & \\ \hline
Keyword & BESTP \\ \hline
\end{tabular}
\section*{A2 Proposal Summary}
\textit{This form, should be filled in by the co-ordinator only, for single
participant proposals by the principal contractor. You should not use more than
3,500 characters. The proposal summary should, at a glance, provide the reader
with a clear understanding of the proposal objectives and how the objectives
will be achieved, and their relevance in the context of the objectives of the
specific programme. This summary may be used as an alternative to the proposal
abstract, as the description of the proposal in the evaluation process and in
communications to the programme management committees and other interested
parties. It must therefore be short and precise. Please use plain typed text,
avoiding formulae and other special characters. If the proposal is written in a
language other than English, please include in form A2 an English version of
the proposal summary.}
\subsection*{Objectives (max 1000 characters)}
to be written
\subsection*{Description of the work (max 2000 characters)}
to be written
\subsection*{Milestones and expected result (max 500 characters)}

\end{document}

### 2001-10-03 drafts/IST/PartB.tex
first release
svn path=/trunk/; revision=1457

\documentclass[a4paper,11pt]{report}
\usepackage[latin1]{inputenc}
\usepackage{vmargin}

%% Marges
\setmarginsrb{2cm}{1cm}{2cm}{2cm}{1cm}{1cm}{1cm}{1cm}
%1 est la marge gauche
%2 est la marge en haut
%3 est la marge droite
%4 est la marge en bas
%5 fixe la hauteur de l'entête
%6 fixe la distance entre l'entête et le texte
%7 fixe la hauteur du pied de page
%8 fixe la distance entre le texte et le pied de page
\setlength{\marginparwidth}{0cm}
\setlength{\marginparsep}{0cm}
%\setcounter{tocdepth}{1}

%% XXX
\title{Proposal full title \\ Proposal Acronym}
\date{16/10/2001}
\begin{document}
\chapter*{Notice}
\begin{verbatim}
IST2001 - IV.3.3 Free software development: towards critical mass 
Objectives:

1.To foster in Europe a critical mass of development of free software released
  under GPL-compatible licenses. 
2.To make available European based support services for free software projects.

Focus:

1.Support services for free software developers at all stages of the project 
  lifecycle: project hosting certification, release, repositories, dissemination. 
2.Trials of free software development in the following fields: media technology 
  for personal users,co-operative information production and sharing, and usability. 
3.Socio-economic studies on technology assessment, economic assessment and derived 
  business models

Types of actions addressed: Trials and studies Accompanying Measures. 
Links with WP 2000: Further focused Action Lines IST 2000 - IV.3.4 and IV.3.5 
\end{verbatim}
\maketitle
\tableofcontents
\chapter*{B3. Objectives}
\textit{This section, which should not exceed two pages, describes the
scientific/technological objectives of the proposal. They should be achiebable
within the project, not through subsequent development, and should be stated in
a measurable and verifiable form. The Progess of the Take-Up actionwill be
measured against theses objectves in later reviews and assessments.\\
This section should clearly specify in addition to the technical objectioves -
also the quantified business objectives and how they are measured}
\\ \\
to be written
\chapter*{B4. Contribution to programme/Key action objectives}
\textit{This section, which should not exceed more than one page, describes how
the proposed Take-Up Action will contribute to the objectives of the programme
and Key action. This can be done by describing how the proposal meets the
objectives and focus of the Action line which it addresses}
\\ \\
to be written
\chapter*{B5. Innovation}
\textit{This secition, not exceeding two pages, describes the innovative
aspects of the proposed project. \\
Progress beyond the state of current practices within a user organisation
should be described. The individual action steps must be described in the
context of the real business case. A separate description of this underlying
baseline project/target application - in terms of technical and business
parameters - must be included. Choice and justfication for each of the new
practices including the supporting technologies (with comparison of not chosen
options) should be given (e.g. new methodologies, technologies, hardware and/or
software tools, subcontracted experts etc. -- whatever is applicable for the
experiment}
\\ \\
to be written.
\chapter*{B6. Project workplan}
\textit{This section consisely describes the work planned to achieve the
objectives of the proposed Take-Up Action. The recommended length, excluding
the forms specified below, is up to 10 pages. An introduction should explain
the structure of th workplan and how the workplan must be broken into
workpackages (WPs) which should follow the logical phases of the projects'
life, and include management of the project and assessment of progress and
results. \\ \\
There must be a separate WP each for 'measureement and evaluation of results}
\\ \\ not yet written
\end{document}

### 2001-10-03 drafts/IST/PartC.tex
PartC
svn path=/trunk/; revision=1461

\documentclass[a4paper,11pt]{report}
\usepackage[latin1]{inputenc}
\usepackage{vmargin}

%% Marges
\setmarginsrb{2cm}{1cm}{2cm}{2cm}{1cm}{1cm}{1cm}{1cm}
%1 est la marge gauche
%2 est la marge en haut
%3 est la marge droite
%4 est la marge en bas
%5 fixe la hauteur de l'entête
%6 fixe la distance entre l'entête et le texte
%7 fixe la hauteur du pied de page
%8 fixe la distance entre le texte et le pied de page
\setlength{\marginparwidth}{0cm}
\setlength{\marginparsep}{0cm}
%\setcounter{tocdepth}{1}

%% XXX
\title{Proposal full title \\ Proposal Acronym}
\date{16/10/2001}
\begin{document}
\chapter*{Notice}
\begin{verbatim}
IST2001 - IV.3.3 Free software development: towards critical mass 
Objectives:

1.To foster in Europe a critical mass of development of free software released
  under GPL-compatible licenses. 
2.To make available European based support services for free software projects.

Focus:

1.Support services for free software developers at all stages of the project 
  lifecycle: project hosting certification, release, repositories, dissemination. 
2.Trials of free software development in the following fields: media technology 
  for personal users,co-operative information production and sharing, and usability. 
3.Socio-economic studies on technology assessment, economic assessment and derived 
  business models

Types of actions addressed: Trials and studies Accompanying Measures. 
Links with WP 2000: Further focused Action Lines IST 2000 - IV.3.4 and IV.3.5

more info : http://www.cordis.lu/ist/bwp_en5.htm
\end{verbatim}
\maketitle
\tableofcontents
\chapter*{C3. Community added value and contribution to EC policies}
\textit{This section, which should not exceed two pages, should identify which
issue at the European level the proposal is addressing. It should also describe
how the experience will be transferred to other organisations across Europe. \\
This section must include a description of the European dimension of their
action, in terms of profile of the wider community which will benefit from the
result of their action. This will help define the European level issue
addressed by the proposal. \\
Participants should be prepared to be grouped under a Support Node or Cluster,
to ensure that those not directly involved can benefit from the results. In
this way, the Take-Up Actions will generate European added value, beyond the
limit achievable through individual projects.}
\\ \\
to be written
\chapter*{C4. Contribution to Community social objectives}
\textit{This section, not exceeding two pages, should describe how the proposed
subject wil contribute to meeting the social objectives of the Community, and
should focus on the contribution of the project to improving employment
prospects and the use and development of skills in Europe. \\
Where aplicable the descriptions should also cover how the proposed measure
contributes to improving quality of life and health and safety (including
working condition), and/or to preserving or enhancing the environment and
natural ressources. This might include the contribution of the proposed work to
meeting relevant regulatory requirements.}
\\ \\
to be written
\chapter*{B5. Management}
\textit[This section, not exceeding two pages, should describe how the Take-Up
Action will be managed, the decision making structures to be applied, the
communication flow within the consortium and the quality assurance measures
which will be implemented, and how legal and ethical obligations, for example
concerning encryption or security of personal data, will be met.
\\
The clustering of Assesment project will allow effective co-ordination, early
agreement - between users and suppliers - on globally accepted assesment
metrics and on competitive specifications, the timely delivery of relevant and
measurable results of assesment and their dissemination to other users
worldwide.\\
Actions management of an action is linked to the management of the underlying
business case (target application, baseline project. Relations and dependencies
including the risks involved should also be described. Beging linked under a
Support Node or Cluster will require co-operation with other actions.
\\ \\
to be written.
\chapter*{C6. Description of the consortium}
\textit{One page description of the consortium stating whom the participants
are, what their roles and functions in the consortium are, and how they
complement each other.\\
Where subcontractors are involved they should be clearly identified, and the
scope and nature of their work needs to be justified}
\\ \\ not yet written
\chapter*{C7. Description of the participants}
\textit{Short description of the participating organisations including (no more
than two pages per organisation):
The expertise and experience of the organisation,
Short CVs of the key persones to be involved indicating relevant exprerience,
expertise and involvement in other EC projects. (Each CV no more than 10
lines). \\
This section should include a brief description of the main business of the
proposer(s) (or consortium, if applicable), including their business motivation
for the proposed action, and the business relevance. \\
It should be clear from this section what the user's current practices are,
which improvements are introduced during the course of the action, and what the
required and expected business practice upon completion of the action is}.
\\ \\ to be written
\chapter{C8. Economic development and scientific and technological prospects}
\textit{This section, which should not exceed three pages, should describe
plans for the dissemination and - if relevant - the exploitation of the results
for the consortium as a whole and for the individual participants in contrete
terms. \\
This section must include an explanation of the potential for - internal and
external - replication, redeployment and/or reuse of results of the action. It
should a) describe in detail how the anticipated results will be used
internally within the participating organisations and what decisions may be
taken upon successful completion of the action, and b) identify the business
communities, sharing the problem to be solved by the action and describe a plan
how the results achieved, benefits obtained and lessons learnt (including key
competencies acquired, deficiencies removed, lasting changes) will be
disseminated/communicated and transferred to those communities. \\
Given the real benefits of an action may only become apparent several months
after completion of the action, proposers should indicate expected benefits in
the longer term (including measures that will be taken to assess them)}
\\ \\
to be written
\end{document}

### 2001-10-04 drafts/IST/PartA.tex
beginning of abstract
svn path=/trunk/; revision=1466

@@
 other than English, please include in form A1 an English version of the
-abstract.} \\
+abstract.} \\ \\ \\
+In almost two years, the biggest barriers for Free Software development has
+been raised by the hosting facilities. They provide most of the mandatory tools that free
+software developers need for everyday tasks. But although there are many hosting facilities 
+around the world, no software package is designed to setup and maintain a set of
+machines dedicated to this purpose. The NAME project will
+provide a generic solution that can be used and adapted by all
+of them. Furthermore, it will 
 \\ \\
@@
 language other than English, please include in form A2 an English version of
-the proposal summary.}
+the proposal summary.} \\
+
 \subsection*{Objectives (max 1000 characters)}

### 2001-10-04 drafts/IST/PartA.tex
fix
svn path=/trunk/; revision=1467

@@
 abstract.} \\ \\ \\
-In almost two years, the biggest barriers for Free Software development has
-been raised by the hosting facilities. They provide most of the mandatory tools that free
-software developers need for everyday tasks. But although there are many hosting facilities 
-around the world, no software package is designed to setup and maintain a set of
-machines dedicated to this purpose. The NAME project will
-provide a generic solution that can be used and adapted by all
-of them. Furthermore, it will 
+In almost two years, the biggest barriers for Free Software development have
+been raised by the hosting facilities which provide most of the mandatory tools
+that free software developers need for everyday tasks. But although there are
+many hosting facilities around the world, no software package is designed to
+setup and maintain a set of machines dedicated to this purpose. The NAME
+project will provide a generic solution that can be used and adapted by all of
+them. Furthermore it will focus on the ease of installation which will fix the
+biggest drawback of the current solutions.
 \\ \\ 
-to be written ...
+586 characters
 \begin{tabular}{|c|c|}

### 2001-10-04 drafts/IST/PartA.tex
upd
svn path=/trunk/; revision=1468

--- 2001-10-03 drafts/IST/PartA.tex
@@
 other than English, please include in form A1 an English version of the
-abstract.} \\
+abstract.} \\ \\ \\
+In almost two years, the biggest barriers for Free Software development have
+been raised by the hosting facilities which provide most of the mandatory tools
+that free software developers need for everyday tasks. But although there are
+many hosting facilities around the world, no software package is designed to
+setup and maintain a set of machines dedicated to this purpose. The NAME
+project will provide a generic solution that can be used and adapted by all of
+them. NAME will of course provide the same services as current facilities but
+will focus an the ease of configuration and installation that all current ones
+lack.
 \\ \\ 
-to be written ...
+628 characters...
 \begin{tabular}{|c|c|}
@@
 language other than English, please include in form A2 an English version of
-the proposal summary.}
+the proposal summary.} \\
 \subsection*{Objectives (max 1000 characters)}
-to be written
+\subsubsection{Context}
+Most of free or open source software is collaboratively developped using an
+hosting plateform. They are mostly web sites. The frontend integrates different
+functionnalities that are available to project developpers and administrators
+and linked them with each other. The backend is mainly implemented with
+scripts. \\
+Current hosting plaforms are implemented as a  web site that provides
+functionnalities to developpers. The most common ones include  users and groups
+of users management, project repository (usualy based on Concurent Versionning
+System), per project web pages and document sharing, per project mailing lists
+and web forums, bugtracking system, project management system (audit, task
+management), global site management and information feeding. There is currently
+four main web hosting facilities : SourceForge (http://sourceforge.net/),
+Savannah (http://savannah.gnu.org/), VHFFS (http://tuxfamily.org/), nd
+Picolibre (http://picolibre.eu.org/). In terms of number of users and projects,
+Savannah and SourceForge are the leaders.
+\subsubsection{Mission}
+Although hosting facilities are widely used in the freesoftware community the
+existing solutions have some drawbacks, that refrain their wide adoption as
+well in the Libre software community than in the software industry. Indeed
+they were poorly designed and had evolved quickly to provide functionnalities
+to developpers. The resulting platforms suffer from some annoying flaws :
+
+<ul> 
+<li> Also due to the number of services and their inter-relations, it is not
+	simple to set up a new site that provides hosting. This is a real
+	brake to a large number of sites using this technology.
+	There is no complete solutions to install easily any of the technologies
+	described above.
+<li> They are hard to distribute, they were mainly designed as a front end to
+	a group of dedicated machines that provide the given set of 
+	functionnalities, versionning system, web access, or data base 
+	access. Even if the access is remote all data is centralized,
+	as one site hosts a huge number of projects. This is a bad
+	practice in terms of data safety, data access and scalability.
+<li> They rely too much on managing procedures and human control. For example, a site
+	needs to be managed by peoplata for mrtg. These script are used on es.gnu.org and france.fsfeurope.org
+es that acknowledge project creation.
+<li> The most active effort to build such a sytem has recently become proprietary.
+Therefore the produced work is not available.
+</ul>
 \subsection*{Description of the work (max 2000 characters)}

### 2001-10-04 drafts/IST/PartA.tex
foo
svn path=/trunk/; revision=1469

@@
 existing solutions have some drawbacks, that refrain their wide adoption as
-well in the Libre software community than in the software industry. Indeed
-they were poorly designed and had evolved quickly to provide functionnalities
-to developpers. The resulting platforms suffer from some annoying flaws :
-
-<ul> 
-<li> Also due to the number of services and their inter-relations, it is not
-	simple to set up a new site that provides hosting. This is a real
-	brake to a large number of sites using this technology.
-	There is no complete solutions to install easily any of the technologies
-	described above.
-<li> They are hard to distribute, they were mainly designed as a front end to
-	a group of dedicated machines that provide the given set of 
-	functionnalities, versionning system, web access, or data base 
-	access. Even if the access is remote all data is centralized,
-	as one site hosts a huge number of projects. This is a bad
-	practice in terms of data safety, data access and scalability.
-<li> They rely too much on managing procedures and human control. For example, a site
-	needs to be managed by peoplata for mrtg. These script are used on es.gnu.org and france.fsfeurope.org
-es that acknowledge project creation.
-<li> The most active effort to build such a sytem has recently become proprietary.
-Therefore the produced work is not available.
-</ul>
-\subsection*{Description of the work (max 2000 characters)}
-to be written
+well in the Libre software community than in the software industry. \\
+Due to a quick design, they are hard to distribute. They mainly are a front end
+to a group of dedicated machines that provides a precise set of
+functionnalities. Even if the acces is remote, all data is centralized as one
+site physically hosts an huge number of projects. This is bad practice in terms
+of data safety, data access and scalability. \\
+Also due to the number of services and their inter-relations, it is not simple
+to set up a new site that provides hosting. This is a real brake to a large
+number of sites using this technology.  There is no complete solutions to
+install easily any of the technologies described above. \\
+They rely too much on managing procedures and human control. For example, a
+site needs to be managed by people that acknowledge project creation. \\
+At last the most active effort to build such a sytem has recently become
+proprietary.  Therefore the produced work is not available. 
+\subsection*{Description of the work (max 2000 characters)} to be written
 \subsection*{Milestones and expected result (max 500 characters)}

### 2001-10-04 drafts/IST/PartB.tex
URL
svn path=/trunk/; revision=1458
layout fix
svn path=/trunk/; revision=1462
typos
svn path=/trunk/; revision=1472

@@
 Links with WP 2000: Further focused Action Lines IST 2000 - IV.3.4 and IV.3.5
+
+more info : http://www.cordis.lu/ist/bwp_en5.htm
 \end{verbatim}
@@
 \textit{This section, which should not exceed two pages, describes the
-scientific/technological objectives of the proposal. They should be achiebable
+scientific/technological objectives of the proposal. They should be achievable
 within the project, not through subsequent development, and should be stated in
-a measurable and verifiable form. The Progess of the Take-Up actionwill be
-measured against theses objectves in later reviews and assessments.\\
-This section should clearly specify in addition to the technical objectioves -
+a measurable and verifiable form. The Progress of the Take-Up action will be
+measured against theses objectives in later reviews and assessments.\\
+This section should clearly specify in addition to the technical objectives -
 also the quantified business objectives and how they are measured}
@@
 to be written
-\chapter*{B4. Contribution to programme/Key action objectives}
+\chapter*{B4. Contribution to program/Key action objectives}
 \textit{This section, which should not exceed more than one page, describes how
-the proposed Take-Up Action will contribute to the objectives of the programme
+the proposed Take-Up Action will contribute to the objectives of the program
 and Key action. This can be done by describing how the proposal meets the
@@
 \chapter*{B5. Innovation}
-\textit{This secition, not exceeding two pages, describes the innovative
+\textit{This section, not exceeding two pages, describes the innovative
 aspects of the proposed project. \\
-Progress beyond the state of current practices within a user organisation
+Progress beyond the state of current practices within a user organization
 should be described. The individual action steps must be described in the
@@
 baseline project/target application - in terms of technical and business
-parameters - must be included. Choice and justfication for each of the new
+parameters - must be included. Choice and justification for each of the new
 practices including the supporting technologies (with comparison of not chosen
@@
 software tools, subcontracted experts etc. -- whatever is applicable for the
-experiment}
+experiment.}
 \\ \\
@@
 \chapter*{B6. Project workplan}
-\textit{This section consisely describes the work planned to achieve the
+\textit{This section concisely describes the work planned to achieve the
 objectives of the proposed Take-Up Action. The recommended length, excluding
 the forms specified below, is up to 10 pages. An introduction should explain
-the structure of th workplan and how the workplan must be broken into
+the structure of the workplan and how the workplan must be broken into
 workpackages (WPs) which should follow the logical phases of the projects'
@@
 results. \\ \\
-There must be a separate WP each for 'measureement and evaluation of results}
+There must be a separate WP each for 'measurement and evaluation of results}
 \\ \\ not yet written

### 2001-10-04 drafts/IST/PartA.tex
objectives
svn path=/trunk/; revision=1475

--- 2001-10-04 drafts/IST/PartA.tex
@@
 project will provide a generic solution that can be used and adapted by all of
-them. Furthermore it will focus on the ease of installation which will fix the
-biggest drawback of the current solutions.
+them. NAME will of course provide the same services as current facilities but
+will focus an the ease of configuration and installation that all current ones
+lack.
 \\ \\ 
-586 characters
+628 characters...
 \begin{tabular}{|c|c|}
@@
 the proposal summary.} \\
-
 \subsection*{Objectives (max 1000 characters)}
-to be written
-\subsection*{Description of the work (max 2000 characters)}
-to be written
+The NAME project aims at developping a complete hosting facility. It composed
+of two subprojects. \\
+The first one is the NAME project that will be a next generation hosting
+facility. It will provide a web frontend to a number of tools are now mandatory
+for software development (user/groups rights management, bug tracking, mailing
+lists and forums...). The main idea between its design is to make it
+transparently distributed. At each step of its development, we will focus on
+the ease of installation and configuration.
+The second one is the "CoopX" project that will provide a generic description of
+an hosting facility. With that tool we want to allow different kind of facility
+to communicate each other. It will help to develop tools to facilitate
+migrations between hosting facilities.
+\\ 800 chars need some check and some other stuff:w
+\subsection*{Description of the work (max 2000 characters)} to be written
 \subsection*{Milestones and expected result (max 500 characters)}

### 2001-10-04 drafts/IST/PartB.tex
objectives
svn path=/trunk/; revision=1475

@@
 \tableofcontents
+\chapter*{some text}
+%%%%%%%%%%%%%%%%%%%
+\subsubsection{Context}
+Most of free or open source software is collaboratively developped using an
+hosting plateform. They are mostly web sites. The frontend integrates different
+functionnalities that are available to project developpers and administrators
+and linked them with each other. The backend is mainly implemented with
+scripts. \\
+Current hosting plaforms are implemented as a  web site that provides
+functionnalities to developpers. The most common ones include  users and groups
+of users management, project repository (usualy based on Concurent Versionning
+System), per project web pages and document sharing, per project mailing lists
+and web forums, bugtracking system, project management system (audit, task
+management), global site management and information feeding. There is currently
+four main web hosting facilities : SourceForge (http://sourceforge.net/),
+Savannah (http://savannah.gnu.org/), VHFFS (http://tuxfamily.org/), nd
+Picolibre (http://picolibre.eu.org/). In terms of number of users and projects,
+Savannah and SourceForge are the leaders.
+\subsubsection{Mission}
+Although hosting facilities are widely used in the freesoftware community the
+existing solutions have some drawbacks, that refrain their wide adoption as
+well in the Libre software community than in the software industry. \\
+Due to a quick design, they are hard to distribute. They mainly are a front end
+to a group of dedicated machines that provides a precise set of
+functionnalities. Even if the acces is remote, all data is centralized as one
+site physically hosts an huge number of projects. This is bad practice in terms
+of data safety, data access and scalability. \\
+Also due to the number of services and their inter-relations, it is not simple
+to set up a new site that provides hosting. This is a real brake to a large
+number of sites using this technology.  There is no complete solutions to
+install easily any of the technologies described above. \\
+They rely too much on managing procedures and human control. For example, a
+site needs to be managed by people that acknowledge project creation. \\
+At last the most active effort to build such a sytem has recently become
+proprietary.  Therefore the produced work is not available. 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \chapter*{B3. Objectives}

### 2001-10-08 drafts/IST/PartA.tex
new objectives, and description of the work
svn path=/trunk/; revision=1483

--- 2001-10-04 drafts/IST/PartA.tex
@@
 In almost two years, the biggest barriers for Free Software development have
-been raised by the hosting facilities which provide most of the mandatory tools
-that free software developers need for everyday tasks. But although there are
+been raised by hosting facilities which provide most of the mandatory tools
+that free software developers need for everyday tasks such as Bug tracking,
+mailing lists, Project management... But although there are
 many hosting facilities around the world, no software package is designed to
@@
 project will provide a generic solution that can be used and adapted by all of
-them. NAME will of course provide the same services as current facilities but
+them.
+NAME will of course provide the same services as current facilities but
 will focus an the ease of configuration and installation that all current ones
@@
 \\ \\ 
-628 characters...
+680 characters...
 \begin{tabular}{|c|c|}
@@
 \subsection*{Objectives (max 1000 characters)}
-\subsubsection{Context}
-Most of free or open source software is collaboratively developped using an
-hosting plateform. They are mostly web sites. The frontend integrates different
-functionnalities that are available to project developpers and administrators
-and linked them with each other. The backend is mainly implemented with
-scripts. \\
-Current hosting plaforms are implemented as a  web site that provides
-functionnalities to developpers. The most common ones include  users and groups
-of users management, project repository (usualy based on Concurent Versionning
-System), per project web pages and document sharing, per project mailing lists
-and web forums, bugtracking system, project management system (audit, task
-management), global site management and information feeding. There is currently
-four main web hosting facilities : SourceForge (http://sourceforge.net/),
-Savannah (http://savannah.gnu.org/), VHFFS (http://tuxfamily.org/), nd
-Picolibre (http://picolibre.eu.org/). In terms of number of users and projects,
-Savannah and SourceForge are the leaders.
-\subsubsection{Mission}
-Although hosting facilities are widely used in the freesoftware community the
-existing solutions have some drawbacks, that refrain their wide adoption as
-well in the Libre software community than in the software industry. \\
-Due to a quick design, they are hard to distribute. They mainly are a front end
-to a group of dedicated machines that provides a precise set of
-functionnalities. Even if the acces is remote, all data is centralized as one
-site physically hosts an huge number of projects. This is bad practice in terms
-of data safety, data access and scalability. \\
-Also due to the number of services and their inter-relations, it is not simple
-to set up a new site that provides hosting. This is a real brake to a large
-number of sites using this technology.  There is no complete solutions to
-install easily any of the technologies described above. \\
-They rely too much on managing procedures and human control. For example, a
-site needs to be managed by people that acknowledge project creation. \\
-At last the most active effort to build such a sytem has recently become
-proprietary.  Therefore the produced work is not available. 
+The NAME project aims at developping a complete hosting facility. It composed
+of two subprojects. \\
+The first one is the NAME project that will be a next generation hosting
+facility. It will provide a web frontend to a number of tools are now mandatory
+for software development (user/groups rights management, bug tracking, mailing
+lists and forums...). The main idea between its design is to make it
+transparently distributed. At each step of its development, we will focus on
+the ease of installation and configuration.
+The second one is the "CoopX" project that will provide a generic description of
+an hosting facility. With that tool we want to allow different kind of facility
+to communicate each other. It will help to develop tools to facilitate
+migrations between hosting facilities.
+\\ 800 chars need some check and some other stuff:w
 \subsection*{Description of the work (max 2000 characters)} to be written
+%SAVANNAH
+The hosting facility is composed of two main parts, the frontend and the
+backend. The first one is the code base that interacts with the user and that
+will implement services that are not provided by external software. The backend
+will link the Database (filled by the frontend) to external software which
+provide services such as DNS, mailing lists or CVS. Concurrently the packaging
+of the solution will allow the validation and testing of each milestone.
+% 450
+%COOPX
 \subsection*{Milestones and expected result (max 500 characters)}
-
+%%% HELP HELP
 \end{document}

### 2001-10-08 drafts/IST/PartA.tex
update
svn path=/trunk/; revision=1484

@@
 \usepackage[latin1]{inputenc}
+\usepackage{graphics}
 \begin{document}
@@
 migrations between hosting facilities.
-\\ 800 chars need some check and some other stuff:w
+\\ 800 chars need some check and some other stuff
 \subsection*{Description of the work (max 2000 characters)} to be written
@@
 %%% HELP HELP
+\section*{A4 Cost Summary}
+%% will look at that later
+\rotatebox{90}{
+\begin{tabular}{c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|}
+Participant Role(CO/P) & Participant Number & Linked to Contractor No &
+Participant Short Name & Number of Person - Months & Personnel Costs &
+Durable Equipment & Consumables & Travel and subsitence & Computing &
+External assistance & Other Costs & Knowledge protection on Costs & 
+Overhead Costs & Total Costs & Requested Contribution from the Commision &
+Eligibility (Y/No)\\
+\end{tabular}
+}
 \end{document}

### 2001-10-08 drafts/IST/PartB.tex
update
svn path=/trunk/; revision=1484
Modifications in abstract, summary
svn path=/trunk/; revision=1485

--- 2001-10-04 drafts/IST/PartB.tex
@@
 \\ \\
-to be written
+
+The first objective is to build a code base with all tools that current hosting
+facilities have. The first audit we made shows there are a Bug tracking system,
+per project mailing lists and web forums, project management tools
+
 \chapter*{B4. Contribution to program/Key action objectives}

### 2001-10-08 drafts/IST/PartA.tex
removed cvs conflicts header
svn path=/trunk/; revision=1487

--- 2001-10-08 drafts/IST/PartA.tex
@@
 that free software developers need for everyday tasks such as Bug tracking,
-mailing lists, Project management... But although there are
-many hosting facilities around the world, no software package is designed to
-setup and maintain a set of machines dedicated to this purpose. The NAME
-project will provide a generic solution that can be used and adapted by all of
-them.
-NAME will of course provide the same services as current facilities but
-will focus an the ease of configuration and installation that all current ones
-lack.
+mailing lists, Project management... But although there are many hosting
+facilities around the world, no software package is designed to setup and
+maintain a set of machines dedicated to this purpose. \\ 
+The NAME project will provide a generic solution that can be used and adapted
+by all of them.  NAME will of course provide the same services as current
+facilities but will focus an the ease of configuration and installation that
+all current ones lack. \\
+NAME will also provide support for internationalization to help dissemination
+and adoption in European countries for non english speakers.\\ Finally NAME
+will introduce support for cooperation in hosting platforms, that will be a
+first step to the creation of networks of hosting facilities. 
 \\ \\ 
-680 characters...
+989 characters... (846 non white) \\
 \begin{tabular}{|c|c|}
@@
 \subsection*{Objectives (max 1000 characters)}
-The NAME project aims at developping a complete hosting facility. It composed
-of two subprojects. \\
-The first one is the NAME project that will be a next generation hosting
-facility. It will provide a web frontend to a number of tools are now mandatory
-for software development (user/groups rights management, bug tracking, mailing
-lists and forums...). The main idea between its design is to make it
-transparently distributed. At each step of its development, we will focus on
-the ease of installation and configuration.
-The second one is the "CoopX" project that will provide a generic description of
-an hosting facility. With that tool we want to allow different kind of facility
-to communicate each other. It will help to develop tools to facilitate
-migrations between hosting facilities.
-\\ 800 chars need some check and some other stuff:w
+The NAME project aims at developping a complete hosting facility. It is 
+composed of two parts. \\
+The first one will be a next generation easy to install and international
+hosting facility. It will provide a web frontend to a number of tools now 
+mandatory for software development (user/groups rights management, bug tracking, mailing lists and forums...). 
+The main idea between its design is to make it transparently distributed.
+At each step of its development, we will focus on the ease of installation and configuration.
+will also provide support for internationalization to help dissemination and adoption in European countries for non english speakers.\\
+The second one will provide a generic description of
+an hosting facility. With that tool we want to allow different kinds of facility
+to communicate with each other. It will help to develop tools to facilitate
+migrations between hosting facilities. This part is more oriented on data formats and tools to exchange informations between hosting facilities. It may
+be released as an individual package from the first part.\\
+\\ 1087/924 chars need some check and some other stuff
 \subsection*{Description of the work (max 2000 characters)} to be written
@@
 provide services such as DNS, mailing lists or CVS. Concurrently the packaging
-of the solution will allow the validation and testing of each milestone.
+of the solution will allow the validation and testing of each milestone.\\
+Our primary goal is to build a set of packages that is easy to
+install on a virgin machine or a machine already used for other
+purposes. 
+Although we have in mind that the set of packages produced could be used for
+multiple purposes, our focus will be on development hosting facilities. 
+created network (for example savannah or sourceforge) ;
+
 % 450

### 2001-10-09 drafts/IST/PartB.tex
objectives
svn path=/trunk/; revision=1496

--- 2001-10-04 drafts/IST/PartB.tex
@@
 \\ \\
-to be written
+\section*{Decentralized.}
+Any given machine running NAME can host a list of
+projects (with pointers to the machines hosting the
+projects), read-only copies of projects (with pointers
+to the machines hosting the read-write instance of the
+project), read-write projects.
+\\
+The read-only projects are mirrored from the machine
+hosting the read-write instance of the project or other
+machines hosting a read-only copy. The format used is
+based on XML and defined by CoopX (http://coopx.eu.org).
+\\
+The number of read-only and read-write projects hosted
+on a given machine is controlled by the maintainer of the
+machine, depending on the available resources. For instance,
+an average machine with lots of bandwidth could host at most
+one hundred read-write projects and ten thousand read-only
+projects.
+\\
+At a given time, only one machine hosts a read-write
+project. It is the responsibility of the project
+maintainers to ensure this.  Should a machine become
+unavailable for some reason, the administrators of the
+project can toggle the read-only flag of one of the
+read-only mirrors and start using it as a
+replacement. 
+\\
+Such a decentralized setup is not the most efficient
+replication system but it is definitely simple and
+requires very little effort.
+The first objective is to build a code base with all tools that current hosting
+facilities have. The first audit we made shows there are a Bug tracking system,
+per project mailing lists and web forums, project management tools
+
+\section*{Multi machine.}
+A single machine may not be powerful enough for someone
+willing to host a large number of projects. Planning a
+multi machine setup in advance is mostly impossible
+because many strategies can be implemented depending on
+the needs. For instance a site like SourceForge will not
+implement the same strategies than a site like Geocities
+because the usage pattern is not the same.
+\\
+A Debian package repository will be used to store the
+packages that implement various strategies. They can all
+coexist together. Let's say, for instance, that the
+package NAME-single implements a single machine setup
+and depends on exim, cvs and bind. There can also be a
+NAME-multi-cvs, NAME-multi-exim, NAME-multi-bind,
+each installed on a separate machine and that configure
+the corresponding service for a multi-machine setup.
+
+\section*{Packaging.}
+
+Every bit is packaged for Debian. Although it may be
+desirable to package it for other GNU/Linux distributions,
+this is not our goal. The cooperative nature of Debian and
+the package dependency management are essential. When the
+set of packages reach a stable state it will be possible
+to contribute them to Debian without a need to find an
+agreement with a company. The dependency management allows
+to consistently use a large set of packages.
+\\
+The packages will be produced in a source specific to the
+project. This source (currently
+http://france.fsfeurope.org/debian/) contains three types
+of packages: copy of existing Debian packages slightly modified
+for the needs of NAME, packages of unfinished software written
+for NAME, packages whose sole purpose is to bind a set of 
+packages together and configure them. 
+\\
+The modified copy of existing Debian packages are only stored
+while waiting for the Debian maintainer or the upstream of the
+software to integrate the change. It is never meant to be 
+permanent. Communication is always established with the Debian
+maintainer and the upstream before the modified copy is built.
+Should we fail to do that, NAME will become a concurrent Debian
+distribution and this is to be avoided at all cost.
+\\
+The software specific to NAME is packaged at a very
+early stage. Basically the package is done whenever the
+software performs at least one task. Properly packaging a
+software takes a lot of time and failing to do it from the
+very beginning is asking for troubles. It also allows
+developpers to use the software and contribute to it in an
+organized way. Finally, since NAME is a collection of
+many packages it allows to debug and architecture their
+interdepencies while writing code.
+
+\section*{Templates and internationalization. }
+
+User visible content is stored in tempates. The preferred
+format for HTML pages is XHTML. 
+\\
+Every string displayed is stored using gettext files and
+obeys the system locale.
+
+\section*{Upgradable.}
+
+Each release provides an upgrade procedure from the
+previous version. The upgrade may imply a planned
+down-time, NAME is commited to provide an upgrade
+facility that would work without stopping the service. The
+XML dump of the database will be used to upgrade the data
+using XSL files. The database structure itself is upgraded
+with a set of SQL orders. The associated services (smtp,
+cvs, etc.) are reset to their initial state (empty) and
+the backend scripts are run on the upgraded database to
+restore the desired state.
+
+\section*{Export/Import using CoopX.}
+
+CoopX provides a set of XML based schemas to represent all
+the data related to a project. This includes the description
+of a user (P3P), contact information (XMLvCard) and custom
+made schemas for the CVS tree and the bug tracker information
+for instance. 
+\\
+The goal of CoopX is not to promote a standard but to use
+them where possible. The bug tracker information, for
+instance, is codified using a DTD with some documentation
+to understand it.  There are no plans to submit it to W3C
+or other similar organizations.
+\\
+The CoopX format is used for upgrades and for communication
+between distinct NAME platforms.
+
 \chapter*{B4. Contribution to program/Key action objectives}

### 2001-10-09 drafts/IST/PartA.tex
Call identifier
svn path=/trunk/; revision=1497
added milestones where PM means project month and delivrables, and more descriptions on project
svn path=/trunk/; revision=1498

@@
 Thematic Priorities & IST-2001-4.3.3. \\ \hline
-Call identifier & \\ \hline
+Call identifier & IST-01-7-1B\\ \hline
 Proposal Date & 16 10 2001 \\ \hline
@@
 be released as an individual package from the first part.\\
-\\ 1087/924 chars need some check and some other stuff
+\\ 
+%1087/924 chars need some check and some other stuff
 \subsection*{Description of the work (max 2000 characters)} to be written
@@
 Although we have in mind that the set of packages produced could be used for
-multiple purposes, our focus will be on development hosting facilities. 
-created network (for example savannah or sourceforge) ;
+multiple purposes, our focus will be on development hosting facilities. \\
 
-% 450
-%COOPX
+To achieve this goal, we will 
+1 redesing existing hosting architectures such as Savannah and Picolibre 
+  so that they may be packaged, installed and taylored easily; 
+2 integrate multilanguage support to these platforms and implement support
+  versions for portuguese, german, spanish, english and french;
+3 define common data types and representation between the two types of
+  platform with possible interaction with other platform via the CoopX project;
+4 introduce functionalities to import/export informations from one site to
+  another, based on this common data representation;
+5 test and deploy over wide area network at least three sites in three out
+	of four participating countries.
+%1445/1224 chars need check and more description
 \subsection*{Milestones and expected result (max 500 characters)}
-%%% HELP HELP
+PM 3: specification of new hosting architecture
+PM 6: specification of common data formats
+PM 9: V1 of savannah with new design 1 language version
+PM12: V1 of picolibre with new design
+      V2 of savannah with support for at least 2 languages and import/export
+PM15: V2 of picolibre with multilingual support for at least 2 languages and
+      import/export functionalities
+PM18: deployment of redesigned platform in 3 countries, 
+      dissemination of the project results by a Debian package.
+%496/408
 \section*{A4 Cost Summary}

### 2001-10-09 drafts/IST/PartB.tex
TOC
svn path=/trunk/; revision=1499

--- 2001-10-08 drafts/IST/PartB.tex
@@
 %%%%%%%%%%%%%%%%%%%
-\subsubsection{Context}
-Most of free or open source software is collaboratively developped using an
-hosting plateform. They are mostly web sites. The frontend integrates different
-functionnalities that are available to project developpers and administrators
-and linked them with each other. The backend is mainly implemented with
-scripts. \\
-Current hosting plaforms are implemented as a  web site that provides
-functionnalities to developpers. The most common ones include  users and groups
-of users management, project repository (usualy based on Concurent Versionning
-System), per project web pages and document sharing, per project mailing lists
-and web forums, bugtracking system, project management system (audit, task
-management), global site management and information feeding. There is currently
-four main web hosting facilities : SourceForge (http://sourceforge.net/),
-Savannah (http://savannah.gnu.org/), VHFFS (http://tuxfamily.org/), nd
-Picolibre (http://picolibre.eu.org/). In terms of number of users and projects,
-Savannah and SourceForge are the leaders.
+
 \subsubsection{Mission}
@@
 \chapter*{B3. Objectives}
+\addcontentsline{toc}{chapter}{B3. Objectives}
 \textit{This section, which should not exceed two pages, describes the
@@
 \\ \\
-
+\section*{Decentralized.}
+Any given machine running NAME can host a list of
+projects (with pointers to the machines hosting the
+projects), read-only copies of projects (with pointers
+to the machines hosting the read-write instance of the
+project), read-write projects.
+\\
+The read-only projects are mirrored from the machine
+hosting the read-write instance of the project or other
+machines hosting a read-only copy. The format used is
+based on XML and defined by CoopX (http://coopx.eu.org).
+\\
+The number of read-only and read-write projects hosted
+on a given machine is controlled by the maintainer of the
+machine, depending on the available resources. For instance,
+an average machine with lots of bandwidth could host at most
+one hundred read-write projects and ten thousand read-only
+projects.
+\\
+At a given time, only one machine hosts a read-write
+project. It is the responsibility of the project
+maintainers to ensure this.  Should a machine become
+unavailable for some reason, the administrators of the
+project can toggle the read-only flag of one of the
+read-only mirrors and start using it as a
+replacement. 
+\\
+Such a decentralized setup is not the most efficient
+replication system but it is definitely simple and
+requires very little effort.
 The first objective is to build a code base with all tools that current hosting
@@
 per project mailing lists and web forums, project management tools
+
+\section*{Multi machine.}
+A single machine may not be powerful enough for someone
+willing to host a large number of projects. Planning a
+multi machine setup in advance is mostly impossible
+because many strategies can be implemented depending on
+the needs. For instance a site like SourceForge will not
+implement the same strategies than a site like Geocities
+because the usage pattern is not the same.
+\\
+A Debian package repository will be used to store the
+packages that implement various strategies. They can all
+coexist together. Let's say, for instance, that the
+package NAME-single implements a single machine setup
+and depends on exim, cvs and bind. There can also be a
+NAME-multi-cvs, NAME-multi-exim, NAME-multi-bind,
+each installed on a separate machine and that configure
+the corresponding service for a multi-machine setup.
+
+\section*{Packaging.}
+
+Every bit is packaged for Debian. Although it may be
+desirable to package it for other GNU/Linux distributions,
+this is not our goal. The cooperative nature of Debian and
+the package dependency management are essential. When the
+set of packages reach a stable state it will be possible
+to contribute them to Debian without a need to find an
+agreement with a company. The dependency management allows
+to consistently use a large set of packages.
+\\
+The packages will be produced in a source specific to the
+project. This source (currently
+http://france.fsfeurope.org/debian/) contains three types
+of packages: copy of existing Debian packages slightly modified
+for the needs of NAME, packages of unfinished software written
+for NAME, packages whose sole purpose is to bind a set of 
+packages together and configure them. 
+\\
+The modified copy of existing Debian packages are only stored
+while waiting for the Debian maintainer or the upstream of the
+software to integrate the change. It is never meant to be 
+permanent. Communication is always established with the Debian
+maintainer and the upstream before the modified copy is built.
+Should we fail to do that, NAME will become a concurrent Debian
+distribution and this is to be avoided at all cost.
+\\
+The software specific to NAME is packaged at a very
+early stage. Basically the package is done whenever the
+software performs at least one task. Properly packaging a
+software takes a lot of time and failing to do it from the
+very beginning is asking for troubles. It also allows
+developpers to use the software and contribute to it in an
+organized way. Finally, since NAME is a collection of
+many packages it allows to debug and architecture their
+interdepencies while writing code.
+
+\section*{Templates and internationalization. }
+
+User visible content is stored in tempates. The preferred
+format for HTML pages is XHTML. 
+\\
+Every string displayed is stored using gettext files and
+obeys the system locale.
+
+\section*{Upgradable.}
+
+Each release provides an upgrade procedure from the
+previous version. The upgrade may imply a planned
+down-time, NAME is commited to provide an upgrade
+facility that would work without stopping the service. The
+XML dump of the database will be used to upgrade the data
+using XSL files. The database structure itself is upgraded
+with a set of SQL orders. The associated services (smtp,
+cvs, etc.) are reset to their initial state (empty) and
+the backend scripts are run on the upgraded database to
+restore the desired state.
+
+\section*{Export/Import using CoopX.}
+
+CoopX provides a set of XML based schemas to represent all
+the data related to a project. This includes the description
+of a user (P3P), contact information (XMLvCard) and custom
+made schemas for the CVS tree and the bug tracker information
+for instance. 
+\\
+The goal of CoopX is not to promote a standard but to use
+them where possible. The bug tracker information, for
+instance, is codified using a DTD with some documentation
+to understand it.  There are no plans to submit it to W3C
+or other similar organizations.
+\\
+The CoopX format is used for upgrades and for communication
+between distinct NAME platforms.
 
@@
 \chapter*{B5. Innovation}
+\addcontentsline{toc}{chapter}{B5. Innovation}
 \textit{This section, not exceeding two pages, describes the innovative
@@
 \\ \\
-to be written.
+%%% just cut & paste for now
+\subsection*{Context}
+Most of free or open source software is collaboratively developped using an
+hosting plateform. They are mostly web sites. The frontend integrates different
+functionnalities that are available to project developpers and administrators
+and linked them with each other. The backend is mainly implemented with
+scripts. \\
+Current hosting plaforms are implemented as a  web site that provides
+functionnalities to developpers. The most common ones include  users and groups
+of users management, project repository (usualy based on Concurent Versionning
+System), per project web pages and document sharing, per project mailing lists
+and web forums, bugtracking system, project management system (audit, task
+management), global site management and information feeding. There is currently
+four main web hosting facilities : SourceForge (http://sourceforge.net/),
+Savannah (http://savannah.gnu.org/), VHFFS (http://tuxfamily.org/), nd
+Picolibre (http://picolibre.eu.org/). In terms of number of users and projects,
+Savannah and SourceForge are the leaders.
+\\
 \chapter*{B6. Project workplan}
+\addcontentsline{toc}{chapter}{B6. Project Workplan}
 \textit{This section concisely describes the work planned to achieve the

### 2001-10-09 drafts/IST/PartB.tex
foo
svn path=/trunk/; revision=1501

@@
 \tableofcontents
-\chapter*{some text}
-%%%%%%%%%%%%%%%%%%%
-
-\subsubsection{Mission}
-Although hosting facilities are widely used in the freesoftware community the
-existing solutions have some drawbacks, that refrain their wide adoption as
-well in the Libre software community than in the software industry. \\
-Due to a quick design, they are hard to distribute. They mainly are a front end
-to a group of dedicated machines that provides a precise set of
-functionnalities. Even if the acces is remote, all data is centralized as one
-site physically hosts an huge number of projects. This is bad practice in terms
-of data safety, data access and scalability. \\
-Also due to the number of services and their inter-relations, it is not simple
-to set up a new site that provides hosting. This is a real brake to a large
-number of sites using this technology.  There is no complete solutions to
-install easily any of the technologies described above. \\
-They rely too much on managing procedures and human control. For example, a
-site needs to be managed by people that acknowledge project creation. \\
-At last the most active effort to build such a sytem has recently become
-proprietary.  Therefore the produced work is not available. 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \chapter*{B3. Objectives}
@@
 Most of free or open source software is collaboratively developped using an
-hosting plateform. They are mostly web sites. The frontend integrates different
+hosting plateform. The frontend integrates different
 functionnalities that are available to project developpers and administrators
@@
 \\
+\subsubsection*{Innovation}
+Although hosting facilities are widely used in the freesoftware community the
+existing solutions have some drawbacks, that refrain their wide adoption as
+well in the Libre software community than in the software industry. \\
+Due to a quick design, they are hard to distribute. They mainly are a front end
+to a group of dedicated machines that provides a precise set of
+functionnalities. Even if the acces is remote, all data is centralized as one
+site physically hosts an huge number of projects. This is bad practice in terms
+of data safety, data access and scalability. \\
+Also due to the number of services and their inter-relations, it is not simple
+to set up a new site that provides hosting. This is a real brake to a large
+number of sites using this technology.  There is no complete solutions to
+install easily any of the technologies described above. \\
+They rely too much on managing procedures and human control. For example, a
+site needs to be managed by people that acknowledge project creation. \\
+At last the most active effort to build such a sytem has recently become
+proprietary.  Therefore the produced work is not available.
+
 \chapter*{B6. Project workplan}

### 2001-10-09 drafts/IST/PartC.tex
layout fix
svn path=/trunk/; revision=1462
toc
svn path=/trunk/; revision=1500
Riek CV
svn path=/trunk/; revision=1502

@@
 \chapter*{C3. Community added value and contribution to EC policies}
+\addcontentsline{toc}{chapter}{C3. Community added value and contribution to EC policies}
 \textit{This section, which should not exceed two pages, should identify which
@@
 \chapter*{C4. Contribution to Community social objectives}
+\addcontentsline{toc}{chapter}{C4. Contribution to Community social objectives}
+\textit{This section, which should not exceed two pages, should identify which
+issue at the European level the proposal is addressing. It should also describe
+how the experience will be transferred to other organisations across Europe. \\
+This section must include a description of the European dimension of their
+action, in terms of profile of the wider community which will benefit from the
+result of their action. This will help define the European level issue
+addressed by the proposal. \\
+Participants should be prepared to be grouped under a Support Node or Cluster,
+to ensure that those not directly involved can benefit from the results. In
+this way, the Take-Up Actions will generate European added value, beyond the
+limit achievable through individual projects.}
+\\ \\
+to be written
+\
 \textit{This section, not exceeding two pages, should describe how the proposed
@@
 to be written
-\chapter*{B5. Management}
-\textit[This section, not exceeding two pages, should describe how the Take-Up
+\chapter*{C5. Management}
+\addcontentsline{toc}{chapter}{C5. Management}
+\textit{This section, not exceeding two pages, should describe how the Take-Up
 Action will be managed, the decision making structures to be applied, the
@@
 including the risks involved should also be described. Beging linked under a
-Support Node or Cluster will require co-operation with other actions.
+Support Node or Cluster will require co-operation with other actions.}
 \\ \\
@@
 \chapter*{C6. Description of the consortium}
+\addcontentsline{toc}{chapter}{C6. Description of the consortium}
 \textit{One page description of the consortium stating whom the participants
@@
 \chapter*{C7. Description of the participants}
+\addcontentsline{toc}{chapter}{C7. Descriptions of the participants}
 \textit{Short description of the participating organisations including (no more
@@
 required and expected business practice upon completion of the action is}.
-\\ \\ to be written
-\chapter{C8. Economic development and scientific and technological prospects}
+\\ \\ 
+%% ALCOVE DE
+\section*{Daniel Riek - Alcove Deutschland CTO}
+Daniel Riek switched to GNU/Linux in late 1993 short after graduating
+from school. In early 1997 during his studies of computing science at the
+university of Bonn he founded the Linux service company ID-PRO together
+with some friends. During the next years he worked as a developer and
+engineer, later he served in the management area of ID-PRO. After ID-PRO
+received a major venture capital financing, Daniel became the chairman
+and speaker of the board of directors. There he was responsible for the
+technical strategy, representation, the community and public relationship.
+He also was, together with the companies CTO, responsible for the technical
+support and the software development activities of the company.
+Caused by the capital market crisis in late 2000 ID-PRO failed to acquire
+new funds, that would have been necessary due to the venture capital driven
+strategy. After a short rest Daniel joined Alcove to continue his work in
+the Free Software business. Daniel is a strong supporter of the Free
+Software Foundation Europe and is a member of the board of germany's Linux
+vendors association Linux-Verband LIVE e.V.
+\\
+Next contributors : Christian, Jaime.
+\chapter*{C8. Economic development and scientific and technological prospects}
+\addcontentsline{toc}{chapter}{C8. Economic development and scientific and
+technological prospecs}
 \textit{This section, which should not exceed three pages, should describe

### 2001-10-09 drafts/IST/PartB.tex
beginning of innovation
svn path=/trunk/; revision=1504

@@
 Most of free or open source software is collaboratively developped using an
-hosting plateform. The frontend integrates different
-functionnalities that are available to project developpers and administrators
-and linked them with each other. The backend is mainly implemented with
-scripts. \\
+hosting plateform. The frontend integrates different functionnalities that are
+available to project developpers and administrators and linked them with each
+other. The backend is mainly implemented with scripts. \\
+
 Current hosting plaforms are implemented as a  web site that provides
@@
 Picolibre (http://picolibre.eu.org/). In terms of number of users and projects,
-Savannah and SourceForge are the leaders.
-\\
-\subsubsection*{Innovation}
+Savannah and SourceForge are the leaders.  \\
+
+\subsubsection*{Limitations of current practice}
+
 Although hosting facilities are widely used in the freesoftware community the
 existing solutions have some drawbacks, that refrain their wide adoption as
-well in the Libre software community than in the software industry. \\
+well in the Libre software community than in the proprietary software industry.
+\\
+
 Due to a quick design, they are hard to distribute. They mainly are a front end
 to a group of dedicated machines that provides a precise set of
-functionnalities. Even if the acces is remote, all data is centralized as one
+functionnalities. Even if the access is remote, all data is centralized as one
 site physically hosts an huge number of projects. This is bad practice in terms
 of data safety, data access and scalability.  \\
+
 Also due to the number of services and their inter-relations, it is not simple
@@
 install easily any of the technologies described above.  \\
+
 They rely too much on managing procedures and human control. For example, a
 site needs to be managed by people that acknowledge project creation. \\
-At last the most active effort to build such a sytem has recently become
-proprietary.  Therefore the produced work is not available.
 
+Internationalization lacks on most platforms, the only one that provides it had
+recently turned proprietary. Thus this effort is not available.
 \chapter*{B6. Project workplan}

### 2001-10-09 drafts/IST/PartB.tex
adapted objectives
svn path=/trunk/; revision=1505

@@
 The packages will be produced in a source specific to the
-project. This source (currently
-http://france.fsfeurope.org/debian/) contains three types
+project. This source will contain three types
 of packages: copy of existing Debian packages slightly modified
@@
 \\
-The modified copy of existing Debian packages are only stored
-while waiting for the Debian maintainer or the upstream of the
-software to integrate the change. It is never meant to be 
-permanent. Communication is always established with the Debian
-maintainer and the upstream before the modified copy is built.
-Should we fail to do that, NAME will become a concurrent Debian
-distribution and this is to be avoided at all cost.
-\\
-The software specific to NAME is packaged at a very
-early stage. Basically the package is done whenever the
+The software specific to NAME will be packaged at a very
+early stage. Basically the package will be done whenever the
 software performs at least one task. Properly packaging a

### 2001-10-09 drafts/IST/PartB.tex
B4 draft
svn path=/trunk/; revision=1506

@@
 \\ \\
-to be written
+
+%% DRAFT, be nice GM
+The NAME project completely follows the \" Take up Action\" second objective.
+Indeed, it is designed to allow anybody even with limited ressources to install
+a development for free software. Therefore, even smallest projects will have a
+complete platform which will use the generic interfaces available on bigger
+sites. Furthermore, free software related organizations will have a complete
+packaged solution to provide all the needed tools for development. Since the
+project is completely distributed, it will have little architecture
+requirements and then will be easy to integrate into an existing network.
+\\ \\
+Furthermore, the availabiliy of such software that makes high quality 
+free software development a reachable reality for everyone had already
+encouraged free software development. Platforms like Savannah or SourceForge
+have allowed a huge number of people to launch their project and to simplify
+its diffusion. We are confident that the availability of an platform that is
+trivial to install and configure will encourage the creation of lots of hosting
+facilities in Europe and then encourages Free software development.
 \chapter*{B5. Innovation}
@@
 
-\subsubsection*{Limitations of current practice}
+\subsection*{Limitations of current practice}
 
@@
 recently turned proprietary. Thus this effort is not available.
+
 \chapter*{B6. Project workplan}

### 2001-10-09 drafts/IST/PartB.tex
toc update
svn path=/trunk/; revision=1507
small fix
svn path=/trunk/; revision=1508

@@
 \chapter*{B4. Contribution to program/Key action objectives}
+\addcontentsline{toc}{chapter}{B4. Contribution to Key action objectives}
 \textit{This section, which should not exceed more than one page, describes how
@@
 
-%% DRAFT, be nice GM
+%% DRAFT, be nice :-) . GM
 The NAME project completely follows the \" Take up Action\" second objective.
@@
 \\ \\
-Furthermore, the availabiliy of such software that makes high quality 
-free software development a reachable reality for everyone had already
-encouraged free software development. Platforms like Savannah or SourceForge
-have allowed a huge number of people to launch their project and to simplify
-its diffusion. We are confident that the availability of an platform that is
-trivial to install and configure will encourage the creation of lots of hosting
-facilities in Europe and then encourages Free software development.
+
+Furthermore, the availabiliy of such software that makes high quality free
+software development a reachable reality for everyone had already encouraged
+free software development. Platforms like Savannah or SourceForge have allowed
+a huge number of people to launch their project and to simplify its diffusion.
+We are confident that the availability of a free and powerful platform in most
+European language will encourage the creation of lots of hosting facilities in
+Europe and then encourages Free software development.
+
 \chapter*{B5. Innovation}

### 2001-10-10 drafts/IST/PartB.tex
trivial fixes
svn path=/trunk/; revision=1523

@@
 %% DRAFT, be nice :-) . GM
-The NAME project completely follows the \" Take up Action\" second objective.
+The NAME project completely follows the '' Take up Action `` second objective.
 Indeed, it is designed to allow anybody even with limited ressources to install
@@
 four main web hosting facilities : SourceForge (http://sourceforge.net/),
-Savannah (http://savannah.gnu.org/), VHFFS (http://tuxfamily.org/), nd
+Savannah (http://savannah.gnu.org/), VHFFS (http://tuxfamily.org/), and
 Picolibre (http://picolibre.eu.org/). In terms of number of users and projects,

### 2001-10-10 drafts/IST/PartC.tex
foo
svn path=/trunk/; revision=1503
First writings of PartC.tex
svn path=/trunk/; revision=1524

@@
 limit achievable through individual projects.}
-\\ \\
-to be written
+\\ 
+\section{European dimension of the problem}
+A big part of free or open software development is done in Europe. As an 
+example see http://www.debian.org/devel/developers.loc, that gives an overview
+of the Debian developers community. There are as many developpers in the 
+western part of europe than there is in the rest of the world.
+This community, as others in Europe, uses
+hosting platforms in Europe as well as in the rest of the world, mostly of
+which are in the United States.
+
+For example, the Savannah hosting architecture is deployed on one machine
+located in Boston Massachusset, this machine is vital to the GNU developpers
+community and it is replicated using rsync techniques in Europe, although 
+the replicating machines have the same informations as the original machine
+Europeen developpers are obliged to use the original machine. This situtation
+is a threat to the data safety, accessibility and security. Let's imagine that
+the building where the computer that is hosting savannah burn, how long 
+will it take to set up a new plant usable in the same way that the current
+machine is. Nobody knows, even if the data are replicated, the procedures to
+switch projects from one machine to another do not exist. Nobody is able to 
+guess if it is possible or how long will it take.
+
+Having many hosting facilities divide the global risk brought by a centralized 
+approach. Having this hosting facilities in area close to the developpers ones
+is a good practice in terms of network usage.
+
+Having hosting facilities designed to support multiligual usage will allow a 
+greater number of developpers to start developping using their national 
+language, lowering the barrier that prevents them from sharing their code.
+
+\section{European added value of the consortium}
+The multinational structure of the consortium will help the consortium being
+concerned with multilingual support in the development and contribute to
+start the platforms with support for four european languages.
+
+The links between the consortium and the GNU project, many members of the 
+consortium are also members of the GNU project,
+will foster the dissemination of the project results and help the project
+in having a real based installed and used by existing projects.
+
+To ensure the NAME project dissemination it will be packaged so that 
+it may be included in existing
+distribution non commercial or commercial.
+
+The presence of an industrial partner that will propose services such as
+installing, tayloring, and dimensionning hosting platforms for 
+industrial purpose, will favour installation of such platforms in industrial
+environment to help entreprises in adopting this kind of technologie with a 
+reduced access cost. The easy installation features of the product will also
+be relevent for industrial usage.
+
+\section{Contribution to EU policies}
+The choice of licensing the results of the project in GPL is the guaranty 
+that the results will remain accessible to anyone in the future.
+
+Contributing to adoption of good practice in software development at low
+cost, will help improve global software production in europe as well as in
+the rest of the world.
+
+Having a hosting platform specific for high education will
+contribute to help teaching the best practices for software development in
+every university at no extra licensing cost.
+
+Building a network of platforms will help EU in increasing the number of 
+developpers that will act for free or open software development.
+
+The tools used in development platforms may also be used for other kinds of 
+groupware activities such as cooperative writing. Providing such tools 
+easily instalable and taylorable may help EU to create new networks of 
+contributors on different topics such as education (cooperative writing of
+courses), helping EU in highering the global level of education by creating
+less cost effectives documents.
+
 \chapter*{C4. Contribution to Community social objectives}
 \addcontentsline{toc}{chapter}{C4. Contribution to Community social objectives}
-\textit{This section, which should not exceed two pages, should identify which
-issue at the European level the proposal is addressing. It should also describe
-how the experience will be transferred to other organisations across Europe. \\
-This section must include a description of the European dimension of their
-action, in terms of profile of the wider community which will benefit from the
-result of their action. This will help define the European level issue
-addressed by the proposal. \\
-Participants should be prepared to be grouped under a Support Node or Cluster,
-to ensure that those not directly involved can benefit from the results. In
-this way, the Take-Up Actions will generate European added value, beyond the
-limit achievable through individual projects.}
-\\ \\
-to be written
-\
 \textit{This section, not exceeding two pages, should describe how the proposed
@@
 \\ \\
-to be written
+The NAME project will help in creating more free/open software in EU by easing 
+the access to a software hosting project platform.
+
+The software created using the hosting platforms, resulting from the NAME 
+project, will help reduce the software costs for individuals as well as for
+small companies and large industrial groups.
+
+The NAME project will also strengthen existing software by allowing more 
+developpers and testers to access hosting platforms and bug reports facilities.
+
+By allowing people to work collectively on projects using their native 
+speaking language,
+the NAME project will help the documentation translation and generaly the 
+software documentation, that will facilitate a larger adoption of free/open
+software in the mass of computer users that don't speak english.
+
 \chapter*{C5. Management}
@@
 \section*{Daniel Riek - Alcove Deutschland CTO}
+%% Need to be improved and must be shorter
 Daniel Riek switched to GNU/Linux in late 1993 short after graduating

### 2001-10-10 drafts/IST/PartC.tex
english and formatting fixes
svn path=/trunk/; revision=1526

--- 2001-10-09 drafts/IST/PartC.tex
@@
 limit achievable through individual projects.}
-\\ \\
-to be written
+\\ 
+\section{European dimension of the problem}
+A big part of free or open software development is done in Europe. As an 
+example see http://www.debian.org/devel/developers.loc, that gives an overview
+of the Debian developers community. There are as more developpers in the 
+western part of europe than in the rest of the world.
+This community, as others in Europe, uses
+hosting platforms in Europe as well as in the rest of the world, mostly of
+which are in the United States.
+\\
+For example, the Savannah hosting architecture is deployed on one machine
+located in Boston Massachusset, this machine is vital to the GNU developpers
+community and it is replicated using rsync techniques in Europe, although 
+the replicating machines have the same informations as the original machine
+Europeen developpers are obliged to use the original machine. This situtation
+is a threat to the data safety, accessibility and security. Let's imagine that
+the building where the computer that is hosting savannah burn, how long 
+will it take to set up a new plant usable in the same way that the current
+machine is. Nobody knows, even if the data are replicated, the procedures to
+switch projects from one machine to another do not exist. Nobody is able to 
+guess if it is possible or how long will it take.
+\\
+Having many hosting facilities divide the global risk brought by a centralized 
+approach. Having this hosting facilities in area close to the developpers ones
+is a good practice in terms of network usage.
+\\
+Having hosting facilities designed to support multiligual usage will allow a 
+greater number of developpers to start developping using their national 
+language, lowering the barrier that prevents them from sharing their code.
+\\
+\section{European added value of the consortium}
+The multinational structure of the consortium will help the consortium being
+concerned with multilingual support in the development and contribute to
+start the platforms with support for four european languages.
+\\
+The links between the consortium and the GNU project, many members of the 
+consortium are also members of the GNU project,
+will foster the dissemination of the project results and help the project
+in having a real based installed and used by existing projects.
+\\
+To ensure the NAME project dissemination it will be packaged so that 
+it may be included in existing
+distribution non commercial or commercial.
+\\
+The presence of an industrial partner that will propose services such as
+installing, tayloring, and dimensionning hosting platforms for 
+industrial purpose, will favour installation of such platforms in industrial
+environment to help entreprises in adopting this kind of technologie with a 
+reduced access cost. The easy installation features of the product will also
+be relevent for industrial usage.
+\\
+\section{Contribution to EU policies}
+The choice of licensing the results of the project under the GNU GPL warrants
+that the results will remain available to anyone in the future.
+\\
+Contributing to adoption of good practice in software development at low
+cost, will help improve global software production in europe as well as in
+the rest of the world.
+\\
+Having a hosting platform specific for high education will
+contribute to help teaching the best practices for software development in
+every university at no extra licensing cost.
+\\
+Building a network of platforms will help EU in increasing the number of 
+developpers that will act for free or open software development. 
+\\
+The tools used in development platforms may also be used for other kinds of 
+groupware activities such as cooperative writing. Providing such tools 
+easily instalable and taylorable may help EU to create new networks of 
+contributors on different topics such as education (cooperative writing of
+courses), helping EU in highering the global level of education by creating
+less cost effectives documents.
+
 \chapter*{C4. Contribution to Community social objectives}
 \addcontentsline{toc}{chapter}{C4. Contribution to Community social objectives}
-\textit{This section, which should not exceed two pages, should identify which
-issue at the European level the proposal is addressing. It should also describe
-how the experience will be transferred to other organisations across Europe. \\
-This section must include a description of the European dimension of their
-action, in terms of profile of the wider community which will benefit from the
-result of their action. This will help define the European level issue
-addressed by the proposal. \\
-Participants should be prepared to be grouped under a Support Node or Cluster,
-to ensure that those not directly involved can benefit from the results. In
-this way, the Take-Up Actions will generate European added value, beyond the
-limit achievable through individual projects.}
-\\ \\
-to be written
-\
 \textit{This section, not exceeding two pages, should describe how the proposed
@@
 \\ \\
-to be written
+The NAME project will help in creating more free/open software in EU by easing 
+the access to a software hosting project platform.
+\\
+The software created using the hosting platforms, resulting from the NAME 
+project, will help reduce the software costs for individuals as well as for
+small companies and large industrial groups. 
+\\
+The NAME project will also strengthen existing software by allowing more 
+developpers and testers to access hosting platforms and bug reports facilities.
+\\
+By allowing people to work collectively on projects using their native 
+speaking language,
+the NAME project will help the documentation translation and generaly the 
+software documentation, that will facilitate a larger adoption of free/open
+software in the mass of computer users that do not speak english.
+\\
+
 \chapter*{C5. Management}

### 2001-10-11 drafts/IST/PartC.tex
add small part in C7.
svn path=/trunk/; revision=1530

@@
 vendors association Linux-Verband LIVE e.V.
+\section*{GET - groupe PeCoVall}
+GET (I'll have the official presentation on friday).\\ 
+The PeCoVall group is composed of 5 professors and researchers 
+from the 3 institutions ENST, ENST Bretagne and INT, that provide high education
+in telecommunication.
+
+This group has been formed last year to set up a project under GPL on an hosting 
+environment specialy designed for high education. 
+The members of the group are much concerned with free/open software and during the 
+last three years, they organised five conferences related to free/open software 
+in their institutions. One member is also a Debian maintener.
+
+The telecommunication background of the group, and the need of some common 
+data between the three sites led them to start a reflexion on cooperative
+exchange in hosting platforms.
+
 \\
-Next contributors : Christian, Jaime.
+Next contributors :  Jaime.
 \chapter*{C8. Economic development and scientific and technological prospects}

### 2001-10-15 drafts/IST/PartC.tex
Added GET description/pecovall description and C. Bac description in PartC.
svn path=/trunk/; revision=1547

--- 2001-10-10 drafts/IST/PartC.tex
@@
 vendors association Linux-Verband LIVE e.V.
-\\
-Next contributors : Christian, Jaime.
+\section*{GET}
+Made up of the six major Graduate Schools of France in the field of Information
+Technology, the Groupe des Ecoles de Telecommunications (GET) represents more 
+than 600 researchers in 60 research labs, 600 PhD, 2500 engineering and 
+management students, 300 master's students. 
+The GET is a major higher education group in France and one of the top level 
+university in the IT field in Europe.
+
+The vocation of the Groupe des Ecoles des Télécommunications is to provide 
+graduate programs in engineering and management tailored to meet these 
+challenges in IT field.
+
+As well as demanding academic program the six schools require their students 
+to spend time in industry or companies. This hands on experience is to give 
+the students first hand knowledge of corporate life and to help them to 
+better approach their first employment.
+
+The strategy of the Groupe des Ecoles des Télécommunications to create 
+partnerships with companies world-wide is intended to help students become 
+an active part of the global market.
+
+The expanding international world of business requires professionals with 
+finely honed personal communications and cross cultural management skills. 
+As a result, the six graduate schools have traditionally been committed 
+to an effective program in foreign languages and social sciences.
+
+Students are required to go abroad to carry out either an internship or an 
+academic program.
+
+The GET works closely with the corporate world through its activities in 
+research, continuing education teaching and internship programs and also 
+welcomes the membership of corporate management for its steering committees. 
+Businesses implicated in this way provide valuable feedback so that it may 
+best serve the evolving needs of the business community.
+
+\subsection*{PeCoVall group}
+The PeCoVall group is composed of five professors and researchers 
+from the three institutions ENST, ENST Bretagne and INT.
+
+This group has been formed last year to set up a project under GPL on an 
+hosting environment specialy designed for high education. 
+The members of the group are much concerned with free/open software and 
+during the last three years, they organised five conferences related to 
+free/open software in their institutions. One member is also a Debian 
+maintener.
+
+The telecommunication background of the group, and the need of some common 
+data between the three sites led them to start a reflexion on cooperative
+exchange in hosting platforms.  \\
+\subsection*{Dr. Christian Bac: GET coordinator}
+Dr. Bac obtained his Bsc in Computer Science from Bordeaux I university.
+He then started to work for France Telecommunication Research and 
+Development.
+Concurently, he obtained his MSc in Computer Science fron Paris
+VI University in 1984. 
+He began to work on UNIX kernel in 1984 and on the Chorus distributed 
+system in 1986.
+
+He moved to the "Institut National des Telecommunications"
+in 1989, as an Assistant Professor
+where he teaches UNIX at various level (user, system calls,
+administration, internals), programming languages C, C++, JAVA and
+distributed computing (distributed applications, distributed
+operating systems).
+
+His research on distributed systems and operating systems were collected
+in a Ph.D. document and he recieved a Ph.D in Computer Science from 
+the University of Paris XI-Orsay in 1995.
+
+He switched his research theme to free software two years ago and sat up a
+group in INT focussing on helping people work with GNU/Linux and other related
+free software. With researchers from the three biggest schools of GET, they
+started a research/development project on hosting platforms focussing on 
+high education needs.
+
+Next contributors :  Jaime.
 \chapter*{C8. Economic development and scientific and technological prospects}

### 2001-10-15 drafts/IST/PartA.tex
ALERT
svn path=/trunk/; revision=1509
Switched from NAME to SICHA.
svn path=/trunk/; revision=1549

@@
 \hline
-Proposal Acronym &  \\ \hline
-Proposal Number &  \\ \hline
+Proposal Acronym &  SICHA \\ \hline
+Proposal Number &  ??????? \\ \hline
 Thematic Priorities & IST-2001-4.3.3. \\ \hline
@@
 Proposal Date & 16 10 2001 \\ \hline
-Proposal full name & \\ \hline
+Proposal full name & Simple International Cooperative Hosting Architecture \\ \hline
 \end{tabular}
@@
 maintain a set of machines dedicated to this purpose. \\ 
-The NAME project will provide a generic solution that can be used and adapted
-by all of them.  NAME will of course provide the same services as current
+The SICHA project will provide a generic solution that can be used and adapted
+by all of them.  SICHA will of course provide the same services as current
 facilities but will focus an the ease of configuration and installation that
 all current ones lack. \\
-NAME will also provide support for internationalization to help dissemination
-and adoption in European countries for non english speakers.\\ Finally NAME
+SICHA will also provide support for internationalization to help dissemination
+and adoption in European countries for non english speakers.\\ Finally SICHA
 will introduce support for cooperation in hosting platforms, that will be a
@@
 \hline
-Duration in months & \\ \hline
+Duration in months & ?????? \\ \hline
 Keyword & BESTP \\ \hline
@@
 \subsection*{Objectives (max 1000 characters)}
-The NAME project aims at developping a complete hosting facility. It is 
+The SICHA project aims at developping a complete hosting facility. It is 
 composed of two parts. \\
@@
 %% will look at that later
+\textbf{This must be discussed very soon}
 \rotatebox{90}{

### 2001-10-15 drafts/IST/PartB.tex
Switched from NAME to SICHA.
svn path=/trunk/; revision=1549

@@
 \section*{Decentralized.}
-Any given machine running NAME can host a list of
+Any given machine running SICHA can host a list of
 projects (with pointers to the machines hosting the
@@
 coexist together. Let's say, for instance, that the
-package NAME-single implements a single machine setup
+package SICHA-single implements a single machine setup
 and depends on exim, cvs and bind. There can also be a
-NAME-multi-cvs, NAME-multi-exim, NAME-multi-bind,
+SICHA-multi-cvs, SICHA-multi-exim, SICHA-multi-bind,
 each installed on a separate machine and that configure
@@
 of packages: copy of existing Debian packages slightly modified
-for the needs of NAME, packages of unfinished software written
-for NAME, packages whose sole purpose is to bind a set of 
+for the needs of SICHA, packages of unfinished software written
+for SICHA, packages whose sole purpose is to bind a set of 
 packages together and configure them. 
 \\
-The software specific to NAME will be packaged at a very
+The software specific to SICHA will be packaged at a very
 early stage. Basically the package will be done whenever the
@@
 developpers to use the software and contribute to it in an
-organized way. Finally, since NAME is a collection of
+organized way. Finally, since SICHA is a collection of
 many packages it allows to debug and architecture their
@@
 previous version. The upgrade may imply a planned
-down-time, NAME is commited to provide an upgrade
+down-time, SICHA is commited to provide an upgrade
 facility that would work without stopping the service. The
@@
 The CoopX format is used for upgrades and for communication
-between distinct NAME platforms.
+between distinct SICHA platforms.
 
@@
 %% DRAFT, be nice :-) . GM
-The NAME project completely follows the '' Take up Action `` second objective.
+The SICHA project completely follows the '' Take up Action `` second objective.
 Indeed, it is designed to allow anybody even with limited ressources to install

### 2001-10-15 drafts/IST/PartC.tex
Switched from NAME to SICHA.
svn path=/trunk/; revision=1549

@@
 \\
-To ensure the NAME project dissemination it will be packaged so that 
+To ensure the SICHA project dissemination it will be packaged so that 
 it may be included in existing
@@
 \\ \\
-The NAME project will help in creating more free/open software in EU by easing 
+The SICHA project will help in creating more free/open software in EU by easing 
 the access to a software hosting project platform.
 \\
-The software created using the hosting platforms, resulting from the NAME 
+The software created using the hosting platforms, resulting from the SICHA 
 project, will help reduce the software costs for individuals as well as for
@@
 \\
-The NAME project will also strengthen existing software by allowing more 
+The SICHA project will also strengthen existing software by allowing more 
 developpers and testers to access hosting platforms and bug reports facilities.
@@
 speaking language,
-the NAME project will help the documentation translation and generaly the 
+the SICHA project will help the documentation translation and generaly the 
 software documentation, that will facilitate a larger adoption of free/open

### 2001-10-15 drafts/IST/PartA.tex
Some minor changes and grammar check.
svn path=/trunk/; revision=1550

@@
 This file will be generated by Protool. Thus layout does not matter. It will
-juste be cut\&paste.
+just be cut\&paste.
 \section*{A1 Administrative Overview}
@@
 abstract.} \\ \\ \\
-In almost two years, the biggest barriers for Free Software development have
-been raised by hosting facilities which provide most of the mandatory tools
-that free software developers need for everyday tasks such as Bug tracking,
-mailing lists, Project management... But although there are many hosting
+In the last two years, the biggest barriers for Free Software development have
+been raised by hosting facilities which provide most of the essential tools
+that free software developers need for everyday tasks such as bug tracking,
+mailing lists, project management... But even though there are many hosting
 facilities around the world, no software package is designed to setup and
@@
 by all of them. SICHA will of course provide the same services as current
-facilities but will focus an the ease of configuration and installation that
-all current ones lack. \\
+facilities ---and more--- but will focus on the ease of configuration and
+installation that all current ones lack. \\
 SICHA will also provide support for internationalization to help dissemination
-and adoption in European countries for non english speakers.\\ Finally SICHA
-will introduce support for cooperation in hosting platforms, that will be a
-first step to the creation of networks of hosting facilities. 
+and adoption for non-English speakers, in European countries.\\
+Finally, SICHA will introduce support for cooperation in hosting platforms;
+that will be a first step to the creation of networks of hosting facilities. 
 \\ \\ 
-989 characters... (846 non white) \\
+998 characters... (846 non white) \\
 \begin{tabular}{|c|c|}
@@
 composed of two parts. \\
-The first one will be a next generation easy to install and international
-hosting facility. It will provide a web frontend to a number of tools now 
-mandatory for software development (user/groups rights management, bug tracking, mailing lists and forums...). 
-The main idea between its design is to make it transparently distributed.
-At each step of its development, we will focus on the ease of installation and configuration.
-will also provide support for internationalization to help dissemination and adoption in European countries for non english speakers.\\
-The second one will provide a generic description of
-an hosting facility. With that tool we want to allow different kinds of facility
-to communicate with each other. It will help to develop tools to facilitate
-migrations between hosting facilities. This part is more oriented on data formats and tools to exchange informations between hosting facilities. It may
-be released as an individual package from the first part.\\
+The first one will be a next generation, easy to install, international
+hosting facility. It will provide a web front-end to a number of tools now 
+essential for software development (user/groups rights management, bug
+tracking, mailing lists and forums...). 
+The main idea guiding its design is to make it transparently distributed.
+At each step of its development, we will focus on the ease of installation and
+configuration. We will also provide support for internationalization, to help
+its dissemination and adoption in European countries for non-English speakers.
+\\
+The second part of the project will provide a specification for a generic
+description of a hosting facility. With that tool we want to allow different
+kinds of facilities to communicate with each other. It will help develop tools
+to facilitate migrations among hosting facilities. This part is more oriented
+towards data formats and tools to exchange information among hosting
+servers. The two parts of the project may be released as separate packages,
+but its joint development will benefit both.\\
 \\ 
 %1087/924 chars need some check and some other stuff
-\subsection*{Description of the work (max 2000 characters)} to be written
+\subsection*{Description of the work (max 2000 characters)}
 %SAVANNAH
-The hosting facility is composed of two main parts, the frontend and the
-backend. The first one is the code base that interacts with the user and that
-will implement services that are not provided by external software. The backend
-will link the Database (filled by the frontend) to external software which
+The hosting facility is composed of two main parts, the front-end and the
+back-end. The first one is the code base that interacts with the user and that
+will implement services that are not provided by external software. The
+back-end will link the database (filled by the front-end) to external
+software which
 provide services such as DNS, mailing lists or CVS. Concurrently the packaging
@@
 Our primary goal is to build a set of packages that is easy to
-install on a virgin machine or a machine already used for other
+install on a brand new machine or a machine already used for other
 purposes. 
 Although we have in mind that the set of packages produced could be used for
-multiple purposes, our focus will be on development hosting facilities. \\
+various purposes, our focus will be on development hosting facilities. \\
 
-To achieve this goal, we will 
-1 redesing existing hosting architectures such as Savannah and Picolibre 
+To achieve the project's goal, we will: 
+1 Re-desing existing hosting architectures such as Savannah and Picolibre 
   so that they may be packaged, installed and taylored easily; 
-2 integrate multilanguage support to these platforms and implement support
-  versions for portuguese, german, spanish, english and french;
-3 define common data types and representation between the two types of
-  platform with possible interaction with other platform via the CoopX project;
-4 introduce functionalities to import/export informations from one site to
+2 Integrate multi-language support to these platforms and implement support
+  versions for Portuguese, German, Spanish, English and French;
+3 Define common data types and representation between the two types of
+  platform (Savannah and Picolibre) and open the possibility of interacting
+  with other platforms via the CoopX project;
+4 Introduce utilities to import/export information from one site to
   another, based on this common data representation;
-5 test and deploy over wide area network at least three sites in three out
-	of four participating countries.
+5 Test and deploy the system over a wide area network; at least at three
+  different sites from the group of four participant countries.
 %1445/1224 chars need check and more description
@@
 PM 6: specification of common data formats
-PM 9: V1 of savannah with new design 1 language version
-PM12: V1 of picolibre with new design
-      V2 of savannah with support for at least 2 languages and import/export
-PM15: V2 of picolibre with multilingual support for at least 2 languages and
+PM 9: V1 of Savannah with new design 1 language version
+PM12: V1 of Picolibre with new design
+      V2 of Savannah with support for at least 2 languages and import/export
+PM15: V2 of Picolibre with multilingual support for at least 2 languages and
       import/export functionalities
 PM18: deployment of redesigned platform in 3 countries, 
-      dissemination of the project results by a Debian package.
+      dissemination of the project results through a Debian package.
 %496/408

### 2001-11-08 fr/donations/demande_autorisation_emission_de_recus.fr.tex
add
svn path=/trunk/; revision=1657

\documentclass[a4paper]{letter}
\usepackage[french]{babel}
\usepackage[latin1]{inputenc}

\begin{document}

\address{Centre Des Impôts Vendôme\\
	Fiscalité Numéro 3\\
	6 rue Ste Hyacinthe\\
	75001 Paris}

\signature{Loïc Dachary\\
	Trésorier}

\begin{letter}{FSF France\\
	M. Dachary - Trésorier\\
	8 rue de Valois\\
	75001 Paris
	}

\opening{Madame, Monsieur,}

Afin de compléter les formalités nécessaires pour rendre possible la
déduction fiscale par les donateurs de l'association \textit{Free
Software Foundation Europe Chapter France (FSF France)} basée au 8 rue
de Valois, 75001 Paris dont je suis le trésorier, je vous adresse
cette demande d'Autorisation d'Émission de reçus.

Vous trouverez ci joint un exemplaire des statuts de l'association
ainsi qu'un exemplaire du reçu que nous pensons utiliser, il est basé
sur le CERFA issu de l'article 200-5 du Code Général des Impôts.

\closing{Merci de prêter la meilleure attention à cette lettre,}
\end{letter}
\end{document}

### 2002-04-28 fr/drafts/fsfgoodies.tex
fsfgoodies.tex
svn path=/trunk/; revision=2136

\documentclass[a4paper,10pt]{article}
\usepackage[french]{babel}
\usepackage[T1]{fontenc}
\usepackage[latin1]{inputenc}

\date{translated \today}
\title{cahier des charges fsf goodies}
\author{Stéphane Salès}

\begin{document}
	\maketitle
	\tableofcontents

\section{existant}

ce qui est fait n'est plus à faire

\subsection{détail de l'existant}

qu'est ce qu'il y a déjà été fait et par qui ?

\section{Besoins}

- personnes participant au projet

- statut des personnes

- paiement en ligne

- obtention des goodies

- stockage des goodies

- éxpédition des goodies

\subsection{personnes}

- nom

- prénom

- adresse(au moins la ville)

- adresse email valide

\subsection{statut des personnes}

Ceci essentielement pour la partie paiement en  ligne .

Qui à pouvoir de décision ?

Qui peut signer un document contractuel ?

\subsection{paiement en ligne}

 contact(s) avec le Crédit Mutuel ?

\subsection{obtention des goodies}

 contacts nécessaire pour l'obtention des goodies ?

\subsection{stockage}

 volume à stocker ?

\subsection{éxpédition}

- reception des commandes ?

- confirmation ?

- papeterie ?

\section{moyens}

liste de diffusion

\subsection{liste de diffusion}

mettre en place une liste de diffusion permettant aux personnes participant au projet de communiquer plus facilement .
Il m'est possibe de l'héberger sur ma machine perso, ainsi que d'autres "services" qui pourraient nous etre utile par la suite, tel qu'un ftp par exemple.

\end{document}

### 2002-04-28 fr/drafts/fsfgoodies.tex
Comments
svn path=/trunk/; revision=2140

@@
 Qui à pouvoir de décision ?
+   Un des trois http://france.fsfeurope.org/about/signatures.html
 
 Qui peut signer un document contractuel ?
+   Un des trois http://france.fsfeurope.org/about/signatures.html
 
@@
  contact(s) avec le Crédit Mutuel ?
+   Je n'ai plus les coordonnées (c'est nul :-( ) Il faut 
+   appeler le crédit mutuel, demander un commercial qui
+   s'occupe de CyberMut, demander les conditions d'adhésion,
+   les papiers nécessaires.
 
@@
  contacts nécessaire pour l'obtention des goodies ?
+ Brian Youmans (3diff@gnu.org) pour FSF, 
+ Peter Gerwinski (peter@gerwinski.de) pour FSF Europe
+ Frédéric Couchet (mad@april.org) pour APRIL (qui a déjà
+  des goodies)
 
@@
 
- volume à stocker ?
+ volume à stocker ?  Stock 8 rue de Valois chez Lolix, Loïc Dachary
+ (loic@gnu.org) a la clé. Acheter une étager pour stocker les
+ cartons. Permettra aussi de stocker les cartons pour APRIL qui a
+ tendance a stocker des trucs au même endroit.
 
@@
 - reception des commandes ?
+  via le web
 
@@
 
+- affranchissement
+
 \section{moyens}
@@
 liste de diffusion
-
-\subsection{liste de diffusion}
-
-mettre en place une liste de diffusion permettant aux personnes participant au projet de communiquer plus facilement .
-Il m'est possibe de l'héberger sur ma machine perso, ainsi que d'autres "services" qui pourraient nous etre utile par la suite, tel qu'un ftp par exemple.
+  fsfe-france@gnu.org tant que pas oppérationnel, fsfe-france-goodies@gnu.org après (suggestion)
 

### 2003-02-01 tools/build.pl
Sync. Automatically updated pages.
svn path=/trunk/; revision=3103

@@ -221,8 +221,10 @@ while (my ($file, $langs) = each %bases) {
             }
 	}
 
-        next if ( (stat("$opts{o}/$dir/$file.$lang.html"))[9] >
-                  (stat($source))[9] && $opts{u} && ! -f "$opts{i}/$file.xsl" );
+        if ( (stat("$opts{o}/$dir/$file.$lang.html"))[9] >
+             (stat($source))[9] && $opts{u} && ! -f "$opts{i}/$file.xsl" ) {
+           next;
+        }
 
         #
         # Here begins automated magic for those pages which we need to
@@ -374,16 +376,16 @@ while (my ($file, $langs) = each %bases) {
         # so that they point to the correct language.
         #
         foreach ($results->documentElement->getElementsByTagName("a")) {
-          if ($_->getAttribute("href") !~ /^http/) {
-            my $href = $_->getAttribute("href");
+          my $href = $_->getAttribute("href");
+          if ($href !~ /^http/) {
             if ($href !~ /\.html$/) {
               if (-d $opts{i}."/$href") {
                 $href =~ s/\/?$/\/index.$lang.html/;
-              } else {
+              } elsif ($href =~ /\/\w+$/) {
                 $href .= ".$lang.html";
               }
             } else {
-              $href =~ s/([^\.][a-z][a-z])\.html/$1.$lang.html/;
+              $href =~ s/([^\.][a-z0-9-][a-z0-9-])\.html/$1.$lang.html/;
             }
             $_->setAttribute("href", $href);
           }
@@ -397,13 +399,29 @@ while (my ($file, $langs) = each %bases) {
   print STDERR "\n" unless $opts{q};
 }
 
+print STDERR "Fixing index links\n" unless $opts{q};
+
+while (my ($path, undef) = each %countries) {   
+  my @dirs = File::Find::Rule->directory()
+                             ->in($opts{o}."/$path");
+  foreach (@dirs) {
+    my $base = basename($_);
+    while (my ($lang, undef) = each %languages) {
+      if (-f "$_/$base.$lang.html" &&
+          ! -f "$_/index.$lang.html") {
+        link("$_/$base.$lang.html", "$_/index.$lang.html");
+      }
+    }
+  }
+}
+
 #
 # For all files that are not XHTML source files, we copy them verbatim to
 # the final location, for each focus. These should be links instead to
[…]

### 2003-02-01 projects/eucd/documents/argumentaire.tex
Moved projects to their right places. Added thankgnus.
svn path=/trunk/; revision=3104

% This file was converted from HTML to LaTeX with
% Tomasz Wegrzanowski's <maniek@beer.com> gnuhtml2latex program
% Version : 0.1
%\documentstyle{article}

\documentclass[a4paper,11pt,french]{article}

\usepackage{german,umlaut}
\usepackage{fancyheadings}
\usepackage{alltt}
\usepackage[german]{babel}
\setlength{\parskip}{1.2ex}
\setlength{\parindent}{0cm}
\setlength{\oddsidemargin}{-0.5cm}
\setlength{\evensidemargin}{-0.5cm}
\setlength{\marginparwidth}{0cm}
\setlength{\marginparsep}{0cm}

\setlength{\topmargin}{-1cm}
\setlength{\textheight}{25cm}
\setlength{\textwidth}{17cm}

\pagestyle{fancyplain}
\lhead[\fancyplain{}{\bfseries\thepage}]
        {\fancyplain{}{\bfseries\rightmark}}
\rhead[\fancyplain{}{\bfseries\leftmark}]
        {\fancyplain{}{\bfseries\thepage}}
\cfoot{}
\renewcommand{\sectionmark}[1]{\markright{\thesection\ #1}}
\renewcommand{\thesection}{\S~\arabic{section}}

\begin{document}

      \section*{Les dommages économiques et sociaux de l'article 6 de l'EUCD}

      \subsection*{1. Préambule}

	{\it 
	  \verb|"|Le livre, comme livre, appartient à l'auteur, mais comme
	  pensée, il appartient - le mot n'est pas trop vaste - au
	  genre humain. Toutes les intelligences y ont droit. Si l'un
	  des deux droits, le droit de l'écrivain et le droit de
	  l'esprit humain, devait être sacrifié, ce serait, certes, le
	  droit de l'écrivain, car l'intérêt public est notre
	  préoccupation unique, et tous, je le déclare, doivent passer
	  avant nous.\verb|"|
	}
	Victor Hugo, 1878 - Discours d'ouverture du congrès littéraire international.

      \par {\bf [Ce document ne remet pas le droit d'auteur en question]}

      \par 
	Les individus, organisations et entreprises qui soutiennent
	l'initiative EUCD.INFO (http://eucd.info/), qui est à
	l'origine de ce document, croient fermement que les droits
	moraux et les intérêts économiques des auteurs doivent être
	préservés dans l'environnement numérique. Il n'est pas dans
	leurs intentions de légitimer des pratiques illicites ni de
	léser les intérêts économiques des auteurs. Bien au
	contraire. D'un point de vue légal, ils pensent qu'il est
	socialement dommageable de criminaliser les pratiques honnêtes
	et légitimes de toute une population sous pretexte de punir
	une minorité de contrevenants. D'un point de vue économique, ils
	défendent fermement une saine concurrence et une rémunération
	équitable des auteurs mais sont hostiles aux monopoles, aux
	abus de position dominante et aux ententes illicites.

      \par {\bf [Justifié par un traité de 1996, l'EUCD met à mal l'intérêt général]}

      \par 
	Le traité OMPI[9]
	  sur le droit d'auteur (1996) a permis le DMCA[4]
	  aux Etats-Unis (1998) et la directive européenne du 22 mai
	  2001 (ou EUCD pour European Union Copyright Directive)[8]. Dans
	  l'EUCD comme dans le DMCA, l'article 11 du traité OMPI a été
	  incorrectement interprété: le législateur a qualifié de
	  contrefaçon tout acte susceptible de neutraliser des mesures
	  techniques de protection[6],
	  remplaçant ainsi partiellement la loi par la technique[15]. Les
	  très nombreuses exceptions incluses dans l'EUCD ne corrigent
	  pas cette erreur fondamentale car ses racines sont sociales
	  et économiques. Victor Hugo contribuait à fonder le droit
	  d'auteur sur cet \verb|"|intérêt général\verb|"| que l'on retrouve dans
	  les considérants de l'EUCD (considérants 3 et 14) mais qui
	  disparaît en raison de l'article 6 relatif à la protection
	  des mesures techniques[7].

      \par {\bf [L'EUCD pose de grands problèmes juridiques mais cette synthèse
	porte sur les problèmes économiques et sociaux]}

      \par 
	Cela fait plus d'un an que des juristes renommés s'escriment
	sur l'article 6 et tentent, en vain,
	de trouver un moyen de le transposer[3]. Ce
	document n'a pas la prétention d'apporter des solutions à ces problèmes
	qui tiennent de la quadrature du cercle pour certains[2][11] et contredisent les objectifs communautaires pour
	d'autres[6]. Il
	s'agit seulement ici de montrer, synthétiquement, les
	nombreuses conséquences négatives, tant sur le plan social
	qu'économique, qu'aurait l'EUCD dans l'hypothèse d'une
	transposition. Nous bénéficions pour ce faire de l'expérience
	américaine après quatre ans d'entrée en vigueur du DMCA
	(1998)[12]. Cela
	permet d'ancrer un argumentaire sur des faits, des expériences
	concrètes et des procédures judiciaires, certaines faisant
	appel à la Cour Suprême des Etats-Unis.

      \par {\bf [Le législateur risque d'ajouter des problèmes lors de la
	transposition]}

      \par 
	Le DMCA et l'EUCD ont une différence qui joue en faveur de
	l'EUCD. Le DMCA pourrait rendre illégale la divulguation
	d'informations relatives aux mesures techniques de protection mais l'EUCD
	ne va pas si loin et permet explicitement, par exemple, la
	diffusion d'alertes de sécurité avertissant d'une faille dans
	un système par laquelle un virus pourrait s'immiscer[5]. Or,
	lors de la transposition de l'EUCD en droit interne, il n'est
	pas interdit au législateur national de modifier le droit
	d'auteur pour le rendre plus contraignant.  On trouve ainsi
	dans le document de travail du CSPLA de Décembre 2002[1]
	visant à transposer l'EUCD en droit français, à l'article 14,
	des termes qui interdisent explicitement la divulgation
	d'informations relatives aux mesures techniques de protection. Lorsque la
	transposition proposée est plus sévère que la directive, il
	revient aux citoyens de rappeler le législateur national à de
	meilleures dispositions.

      \par {\bf [Un contexte économique et social radicalement différent rend
	l'EUCD obsolète]}

      \par 
	On peut accorder des circonstances atténuantes aux rédacteurs
	de 1996; l'internet naissant à peine il était possible
	de céder à la tentation de rédiger l'article 11 du
	traité OMPI[9]. Mais
	le législateur d'aujourd'hui vit dans un monde radicalement
	différent et ne peut l'ignorer. Quand le respectable institut
	Forrester publie en août 2002 une étude titrée \verb|"|Downloads Save
	The Music Business\verb|"|[10],
	il est indispensable de remettre en question la pertinence de
	dispositions législatives dont les fondements économiques sont
	d'une autre époque.

      \par {\bf [L'EUCD n'est qu'une étape d'une démarche dont l'orientation
	doit être corrigée]}

      \par 
	L'EUCD fait partie d'une démarche d'ensemble et sera suivie
	par une directive concernant la lutte contre la contrefaçon,
	la gestion de droits numériques.  Les Etats-Unis ont pris un
	peu d'avance sur ces sujets et nous permettent donc d'en
	observer les effets[12].
	Aujourd'hui les modifications du droit d'auteur qui sont
	proposées dans les pays d'Europe servent les puissants[13] au détriment
	du grand public et des auteurs[14]. Il
	est nécessaire d'agir dès maintenant pour enrayer cette
	logique et ne pas léser les générations futures.

      \par Références:

      \begin{itemize}
	\item [1] FR: Décembre 2002. Document de travail CSPLA 5-12-02 (article 14) \\
http://www.planetelibre.net/article/cspla/doc.txt
	\item [2] FR: Janvier 2002. Propriétés intellectuelles. p52-57. Gilles Vercken, Recherche clarté désespérément: à propos de l'article 6.4 de la directive du 22 mai 2001. {\it Je souhaite citer les opinions d'éminents juristes et professeurs de droit à propos de l'article 6.4 de la directive. \verb|"|C'est l'une des questions des plus épineuses\verb|"|, \verb|"|le résultat n'a pas le mérite de la clarté\verb|"| - et, à propos du paragraphe 2 sur la copie privée, \verb|"|c'est l'ensemble du paragraphe qui échappe à la compréhension\verb|"| nous dit Séverine Dussolier, chercheuse au CRID; \verb|"|un texte dont l'application s'avérera très délicate\verb|"|, d'après le Professeur Christophe Caron. \verb|"|Un texte très, voire trop complexe\verb|"|, affirment le Professeur Alain Strowel et Séverine Dusollier. \verb|"|Le texte laisse perplexe\verb|"| écrit le Professeur Pierre Sirinelli et il ajoute : \verb|"|Les Etats seront sans doute embarassés au moment de transposer le texte communautaire\verb|"|.}
	\item [3] FR: 2002. L'articulation des exceptions au droit d'auteur et des mesures techniques de protection face à l'harmonisation \\
http://www.droit-ntic.com/pdf/dir\_da.pdf
	\item [4] US: Décembre 1998. Digital Millennium Copyrigth Act (DMCA) \\
http://www.loc.gov/copyright/legislation/dmca.pdf
	\item [5] US: Septembre 2001. Anticircumvention Rules: Threat to Science. \\
http://www.law.upenn.edu/law619/f2001/week09/samuelson\_dmca.pdf
	\item [6] EU: Octobre 2000. Why the Copyright Directive is Unimportant, and Possibly Invalid \\
http://www.ivir.nl/publications/hugenholtz/opinion-EIPR.html
	\item [7] EU: Juin 1998. EBLIDA's five minutes Statement on the proposed Directive on the harmonisation of certain aspects of copyright and related rights in the Information Society. \\
http://www.eblida.org/topics/position/legaffa.htm
	\item [8] EU: Mai 2001. Directive 2001/29/CE du 22 mai 2001 \\
http://eucd.info/directive-2001-29-ce.pdf
	\item [9] INT: Décembre 1996. Traité de l'OMPI sur le droit d'auteur \\
http://www.wipo.int/clea/docs/fr/wo/wo033fr.htm
	\item [10] INT: Août 2002. Downloads Save The Music Business. \\
http://www.forrester.com/ER/Research/Report/Summary/0,1338,14854,FF.html
	\item [11] BE: Mars 2000. Réponse à une question du Sénateur Monfils: \verb|"|La relation entre les exceptions (article 5) et la protection des mesures techniques de protection auxquelles ont recours les ayants droit en vue de protéger leurs oeuvres, reste la difficulté majeure des discussions\verb|"|
	\item [12] US: 2003. EFF Whitepaper: Unintended Consequences, Three Years under the DMCA \\
http://www.eff.org/IP/DMCA/20030102\_dmca\_unintended\_consequences.html.html
	\item [13] INT: 2002. BMG Company Statement on Copy Control \\
http://www.bmg-copycontrol.info/
        \item [14] INT: 2001. Information rights and intellectual freedom, Julie E. Cohen (page 8, 9 et 11). \\
http://www.law.georgetown.edu/faculty/jec/intellfreedom.pdf
        \item [15] INT: 2001. Information rights and intellectual freedom, Julie E. Cohen (page 52). \\
http://www.law.georgetown.edu/faculty/jec/intellfreedom.pdf
      \end{itemize}

      \subsection*{Quels sont les effets négatifs de l'EUCD ?}

      \begin{description}
	\item[{\bf 2. Menace la copie privée}]
	  \par 
	    La copie privée est une pratique licite dans la plupart
	    des pays d'Europe. L'EUCD interdit la distribution de
	    logiciels de contournement de mesures techniques  de protection et
	    n'oblige pas les auteurs de mesures techniques de protection à en
	    permettre le contournement à des fins de copie privée. Par
	    conséquent l'EUCD a pour effet de rendre très onéreux ou
	    pratiquement impossible la copie privée d'une oeuvre
	    protégée par une mesure technique.

	  \par 
	    La copie privée étant de facto interdite par la protection
	    des mesures techniques, la rémunération pour copie privée,
	    lorsqu'elle existe en droit national, n'a plus lieu
	    d'être.

	  \par 
	    En France, l'exception de copie privée a été insérée dans
	    la loi en 1957. La rémunération pour copie privée a été
	    insérée en 1985. L'exception de copie privée ne concerne
	    pas les logiciels.

          \par 
	    En Belgique, l'exception de copie privée a été insérée
	    dans la loi en 19??. En 1996, un arreté royal est déposé
	    pour la rémunération pour copie privée.

	  Références:
	  \begin{itemize}
	    \item {\bf Le CSPLA affirme son souci de préserver la copie privée lors de la transposition de l'EUCD.}FR: 27 Juin 2002. Compte rendu CSPLA (page 9). \\
http://www.culture.fr/culture/cspla/co270602.pdf
	    \item {\bf L'EUCD ne permet pas de faire coexister la copie privée et les mesures techniques de protection.}EU: Octobre 2000. Why the Copyright Directive is Unimportant, and Possibly Invalid (paragraphe {\it I have read and reread this text several times ...}) \\
http://www.ivir.nl/publications/hugenholtz/opinion-EIPR.html
	    \item {\bf L'EUCD ne permet pas de faire coexister la copie privée et les mesures techniques de protection.} EU: 2002. L'articulation des exceptions au droit d'auteur et des mesures techniques de protection face à l'harmonisation (page 53 et suivantes). \\
http://www.droit-ntic.com/pdf/dir\_da.pdf
	    \item {\bf Les dommages sociaux de la réduction du fair use.} INT: 2001. Information rights and intellectual freedom, Julie E. Cohen (page 8, 9 et 11). \\
http://www.law.georgetown.edu/faculty/jec/intellfreedom.pdf
	    \item {\bf Les enjeux de la copie privée.} INT: 2000. The Digital Dilemma (chapitre 4, The challenge of private use and fair use with digital information). \\
http://www.nap.edu/html/digital\_dilemma/ch4.html
	  \end{itemize}

	\item[{\bf 3. Porte atteinte au droit de lire, au droit d'usage}]
	  \par 
	    Une personne loue un film sur DVD. Elle utilise une copie
	    du logiciel DeCSS (qui permet de décrypter le film stocké
	    sur le DVD) et le regarde sur son ordinateur. Or l'EUCD
	    interdit la distribution du logiciel DeCSS, par conséquent
	    cette personne est en situation de contrefaçon. Donc seuls
	    les logiciels autorisés par l'auteur de la mesure
	    technique de protection (CSS dans ce cas) permettent de lire les
	    contenus protégés. Si l'utilisateur ne dispose pas de ces
	    logiciels autorisés, pour une raison ou pour une autre, il
	    lui est de fait interdit de lire l'oeuvre pour laquelle il
	    s'est déjà acquitté des droits. Il en va de même, par
	    exemple, pour les eBook.

	  Références:
	  \begin{itemize}
	    \item {\bf L'EUCD veut interdire la distribution de certains logiciels.} EU: 2002. L'articulation des exceptions au droit d'auteur et des mesures techniques de protection face à l'harmonisation (page 9). \\
http://www.droit-ntic.com/pdf/dir\_da.pdf
	    \item {\bf Le DMCA interdit DeCSS.} INT: Décembre 2001. The law \& economics of reverse engineering (page 61). \\
http://www.sims.berkeley.edu/\verb|~|pam/papers/l\&e\%20reveng5.pdf
	    \item {\bf Le DMCA interdit la distribution de logiciels permettant de lire un eBook.} US: Décembre 2002. U.S. v. ElcomSoft \& Sklyarov. \\
http://www.eff.org/IP/DMCA/US\_v\_Elcomsoft/us\_v\_elcomsoft\_faq.html\#IllegalInUSA 
	    \item {\bf Un eBook qui ne peut être lu à haute voix.}EU: Février 2001.  Don't Read Aloud This Version of Alice in Wonderland. \\
http://www.thestandard.com/article/display/0,1151,22377,00.html
	    \item {\bf Le droit de lire.} INT: Février 1997 mis à jour en 2002. The Right to Read, Communications of the ACM (Volume 40, Number 2). \\
http://www.gnu.org/philosophy/right-to-read.html
	    \item {\bf Information à propos de CSS et DeCSS.} INT: DeCSS Central. \\
http://www.lemuria.org/DeCSS/decss.html
	  \end{itemize}

	\item[{\bf 4. Nie le principe des bibliothèques, de l'accès à la
	    culture pour tous et du domaine public}]
	  \par 
	    Les bibliothèques et les archives font un pont entre les
	    générations. Les oeuvres numériques sont éternelles car la
	    sauvegarde d'une oeuvre numérique d'un support à un autre
	    est quasi instantanée et ne pose pas les problèmes
	    pratiques liés aux oeuvres stockées sur support papier. Les
	    mesures techniques qui protègent des oeuvres numériques
	    sont donc une menace d'envergure, même sans l'EUCD. Elles
	    sont conçues pour empêcher des personnes d'accéder à
	    l'oeuvre, pour limiter cet accès à certaines personnes
	    dans certaines circonstances. C'est un principe totalement
	    opposé à celui des bibliothèques.

	  \par 
	    Au Moyen-Age, l'accès à la littérature imposait
	    l'apprentissage d'un code particulier, le latin. La
	    protection légale des mesures techniques ajoute un code de
	    même nature dont la clé est l'argent. Cette condition
	    réduit donc l'accès à la culture pour tous.

	  \par 
	    Dans le cas d'oeuvres tombées dans le domaine public et
	    archivées, l'EUCD interdit de fait aux bibliothèques de
	    contourner les mesures techniques qui les protègent.  En
	    effet, bien qu'elles en aient la permission en théorie,
	    comme la distribution de logiciels de contournement est
	    prohibée, les bibliothèques se trouvent dans
	    l'impossibilité pratique de jouir de cette permission. Par
	    ailleurs, alors que le droit d'auteur est limité dans le
	    temps, les mesures techniques de protection n'ont pas, au
	    titre de l'EUCD, obligation de se dévérouiller dans le
	    même délai. Voici un exemple parmi d'autres: une
	    bibliothèque fait l'acquisition d'un livre sous forme
	    numérique et assorti d'une mesure technique de
	    protection. Vingt ans plus tard, le livre tombe dans le
	    domaine public et le logiciel permettant d'y accéder
	    n'existe plus, la société qui l'éditait ayant déposé le
	    bilan depuis dix ans. La bibliothèque se trouve dans
	    l'impossibilité pratique de copier le livre sous une forme
	    non protégée par la mesure technique afin de le mettre à
	    disposition du public. Dans le meilleur des cas, la
	    bibliothèque devra chercher et faire l'acquisition d'un
	    autre exemplaire, non protégé par une mesure technique,
	    doublant ainsi ses frais. Dans le cas d'un livre seulement
	    distribué sous forme numérique et protégé par une mesure
	    technique devenue obsolète, la bibliothèque perd la 
	    possibilité d'accéder à ce livre.

	  Références: 
	  \begin{itemize}
	    \item {\bf La technique doit rester neutre quant à l'application du droit d'auteur et de ses exceptions} FR: Juillet 2002. Communiqué commun concernant la position de la France sur la transposition de la directive européenne du 22 mai 2001 sur le droit d'auteur. \\
http://addnb.org/fr/docs/Communiqu\%E9\%20du\%2016\%20juillet\%202002.htm
            \item {\bf La transposition l'EUCD doit préserver le libre accès à la culture pour tous.} BE: Mai 2000. Le libre accès l'information numérique, ABD-BVD, \\
http://www.abd-bvd.be/misc/lib/ds-fr.html
	    \item {\bf La transposition l'EUCD doit préserver le libre accès à la culture pour tous.} FR: Janvier 2002. La diffusion numérique du patrimoine, dimension de la politique culturelle (page 11) \\
http://www.culture.fr/culture/actualites/rapports/ory-lavollee/ory-lavollee.pdf
	    \item {\bf Critique des effets du DMCA sur l'accès à la culture pour tous.} US: Août 2000. Comments of the Library Associations. \\
http://www.copyright.gov/reports/studies/dmca/comments/Init018.pdf
	    \item {\bf Critique des effets de l'EUCD sur l'accès à la culture pour tous.} EU: Juin 1998. EBLIDA's five minutes Statement on the proposed Directive on the harmonisation of certain aspects of copyright and related rights in the Information Society. \\
http://www.eblida.org/topics/position/legaffa.htm
	    \item {\bf Le droit de lire.} INT: Février 1997 mis à jour en 2002. The Right to Read, Communications of the ACM (Volume 40, Number 2), \\
http://www.gnu.org/philosophy/right-to-read.html
	  \end{itemize}

	\item[{\bf 5. Viole la vie privée}]
	  \par 
	    L'EUCD entend accorder une protection légale à toutes les
	    mesures techniques de protection \verb|"|efficaces\verb|"|.  Si
	    l'efficacité d'une mesure technique repose sur la collecte
	    de données personnelles, une personne divulgant des
	    informations fausses pourrait être accusée de contourner
	    une mesure technique de protection. Or, l'efficacité d'une
	    mesure technique de protection sur les réseaux dépend de
	    la collecte de données personnelles. Il en découle que,
	    sur les réseaux, les mesures techniques de protection sont
	    soit inefficaces, soit contreviennent aux dispositions
	    réglementaires concernant la protection de la vie privée.

	  Références:  
	  \begin{itemize}
	    \item {\bf Une mesure technique de protection efficace implique nécessairement une violation de la vie privée.} FR: Octobre 2001. Rapport particulier de Monsieur Leonardo Chiariglione (Telecom Italia Lab, Italie) portant sur la gestion et la protection des oeuvres et de la propriété intellectuelle. Etat des travaux et des réflexions. (Point C. La protection de la vie privée des utilisateurs). \\
http://www.culture.fr/culture/cspla/rapchariglione.htm
	    \item {\bf Les dommages sociaux de l'appropriation des données personnelles.} INT: 2001. Information rights and intellectual freedom, Julie E. Cohen (page 7). \\
http://www.law.georgetown.edu/faculty/jec/intellfreedom.pdf
	    \item {\bf Le droit de lire.} INT: Février 1997 mis à jour en 2002. The Right to Read, Communications of the ACM (Volume 40, Number 2), \\
http://www.gnu.org/philosophy/right-to-read.html
	  \end{itemize}

	\item[{\bf 6. Force les ventes liées}]
	  \par 
	    Avec l'EUCD, en plus de l'obtention de droits sur une
	    oeuvre numérique protégée par une mesure technique, la
	    personne qui souhaite en jouir doit faire l'acquisition
	    d'un logiciel autorisé par l'auteur de la mesure technique
	    qui protège l'oeuvre.  Par exemple, une personne loue un
	    film sur DVD. Pour regarder ce film, l'EUCD l'oblige à
	    utiliser les logiciels qui ont été autorisés par l'auteur
	    de la mesure technique qui protège le film (CSS) à
	    l'exclusion de ceux qui seraient basés, par exemple, sur
	    DeCSS qui est un logiciel qui contourne la mesure
	    technique de protection.

	  Références: 
	  \begin{itemize}
	    \item {\bf La protection des mesures techniques crée un marché captif.} INT: Décembre 2001. The law \& economics of reverse engineering (page 63). \\
http://www.sims.berkeley.edu/\verb|~|pam/papers/l\&e\%20reveng5.pdf
	    \item {\bf Information à propos de CSS et DeCSS.} INT: DeCSS Central. \\
http://www.lemuria.org/DeCSS/decss.html
	  \end{itemize}

	\item[{\bf 7. Met en danger l'économie}]

	  \par 
	    Une économie compétitive dépend de la possibilité pour le
	    consommateur ou pour l'entreprise de remplacer un produit
	    par un autre. Pour des biens de consommation simples tels
	    qu'un bol ou une chaise, c'est à l'évidence possible. Pour
	    des biens numériques pour lesquels il existe une mesure
	    technique de protection, cela peut s'avérer
	    impossible. L'article 6 de l'EUCD empêche de pratiquer le
	    reverse engineering (extraction de savoir faire par
	    l'observation d'un objet fabriqué par l'homme) qui est
	    nécessaire pour créer et diffuser des produits
	    concurrents, car le reverse engineering implique un
	    contournement de mesure technique de protection. Par
	    conséquent, les auteurs de mesures techniques de
	    protection peuvent se servir de l'EUCD pour empêcher la
	    création de produits compatibles ou susceptibles de se
	    substituer aux leurs.

	  Références: 
	  \begin{itemize}
	    \item {\bf De l'impossibilité de faire du reverse engineering sans contourner une mesure technique de protection.} INT: Décembre 2001. The law \& economics of reverse engineering (page 50). \\
http://www.sims.berkeley.edu/\verb|~|pam/papers/l\&e\%20reveng5.pdf
	  \end{itemize}

	  \begin{description}
	    \item[8. Permet des monopoles sur les formats de fichier]
	      \par Il suffit à l'auteur d'un format de fichier d'inclure
		dans celui-ci une mesure technique de protection pour
		détenir un monopole de fait sur l'exploitation de ce
		format. Le logiciel DeCSS permet de decrypter les
		films stockés sur les DVD. S'il est interdit de le
		distribuer, alors l'auteur du cryptage CSS dispose
		d'un monopole sur les logiciels permettant de
		décrypter CSS. Autoriser la distribution de DeCSS
		revient à autoriser la distribution d'un logiciel
		permettant le contournement d'une mesure technique de
		protection.

	      Références: 
	      \begin{itemize}
		\item {\bf Le DMCA interdit la distribution de logiciels permettant de lire un eBook.} US: Décembre 2002. U.S. v. ElcomSoft \& Sklyarov. \\
http://www.eff.org/IP/DMCA/US\_v\_Elcomsoft/us\_v\_elcomsoft\_faq.html\#IllegalInUSA 
	      \end{itemize}

	    \item[9. Encourage les abus de position dominante]
	      \par 
		L'ajout d'une mesure technique de protection dans un logiciel
		étant suffisant pour obtenir un monopole de fait,
		la concurrence ne régule pas les prix en fonction
		de l'offre et de la demande. Le fournisseur du
		logiciel aura donc naturellement tendance à le
		proposer à un prix supérieur à sa valeur théorique
		en situation de concurrence.

	      Références: 
	      \begin{itemize}
		\item {\bf La protection des mesures techniques fait monter les prix.} INT: Décembre 2001. The law \& economics of reverse engineering (page 45, 46 et 56) \\
http://www.sims.berkeley.edu/\verb|~|pam/papers/l\&e\%20reveng5.pdf
	      \end{itemize}

	    \item[10. Encourage les ententes illicites]
	      \par 
		Les titulaires de droits et les auteurs de mesures
		techniques sont contraints de passer entre eux des
		accords. Si l'ensemble de ces acteurs ont un monopole,
		il peut s'agir d'entente illicite. Un acteur
		indépendant qui voudrait s'affranchir de ces accords
		tout en continuant à utiliser la même mesure
		technique serait en situation de contrefaçon.

	      Références: 
	      \begin{itemize}
		\item {\bf Le DMCA est utilisé pour forcer une politique de prix.} INT: Décembre 2001. The law \& economics of reverse engineering (page 62) \\
http://www.sims.berkeley.edu/\verb|~|pam/papers/l\&e\%20reveng5.pdf
	      \end{itemize}

	    \item[11. Empêche une saine concurrence]
	      \par 
		L'ajout d'une mesure technique de protection sur une oeuvre est
		utilisé pour empêcher l'apparition de concurrents. Un
		concurrent potentiel ne peut distribuer ou concevoir
		un produit de substitution compatible car pour ce
		faire il doit nécessairement contourner la mesure
		technique de protection.

	      Références: 
	      \begin{itemize}
		\item {\bf Le DMCA utilisé pour éliminer un concurrent.} US: 19 Février 2002. Blizzard v. bnetd \\
http://www.eff.org/IP/Emulation/Blizzard\_v\_bnetd/20020219\_blizzard\_bnetd\_letter.html
		\item {\bf Empêcher le reverse engineering permet un monopole sur les idées.} INT: Décembre 2001. The law \& economics of reverse engineering (page 32, 33). \\
http://www.sims.berkeley.edu/\verb|~|pam/papers/l\&e\%20reveng5.pdf
		\item {\bf Le DMCA nuit à la concurrence sur les contenus.} INT: Décembre 2001. The law \& economics of reverse engineering (page 56, 61). \\
http://www.sims.berkeley.edu/\verb|~|pam/papers/l\&e\%20reveng5.pdf
	      \end{itemize}

	    \item[12. Menace l'interopérabilité]
	      \par 
		Prenons un serveur de jeux vidéo sur internet. Les
		utilisateurs se connectent au serveur avec un logiciel
		client afin de jouer ensemble. Le serveur vérifie le
		numéro de série du logiciel client lorsqu'il se
		connecte: c'est une mesure technique de protection de
		l'ensemble logiciel et données graphiques du
		client. La société Blizzard exploite le serveur et
		publie le client. Un serveur concurrent est créé,
		compatible (interopérable) avec le client publié par
		Blizzard. Ce serveur, exploité et publié par bnetd.org
		ne contient pas la mesure technique de protection. Il
		permet donc un contournement de la mesure technique de
		protection constituée par le couple client/serveur de
		Blizzard. Par exemple, un contrefacteur ayant fait une
		copie illicite du client fourni par Blizzard peut
		interagir avec le serveur bnetd.org car celui-ci ne
		sait pas vérifier le numéro de série du client. C'est
		sur cette base que Blizzard poursuit actuellement
		bnetd.org.

	      Références: 
	      \begin{itemize}
		\item {\bf Le DMCA utilisé pour sanctionner un serveur compatible.} US: 19 Février 2002. Blizzard v. bnetd \\
http://www.eff.org/IP/Emulation/Blizzard\_v\_bnetd/20020219\_blizzard\_bnetd\_letter.html
		\item {\bf Empêcher le reverse engineering permet un monopole sur les idées.} INT: Décembre 2001. The law \& economics of reverse engineering (page 32, 33). \\
http://www.sims.berkeley.edu/\verb|~|pam/papers/l\&e\%20reveng5.pdf
	      \end{itemize}

	    \item[13. Supprime les bénéfices des usages non autorisés]

	      \par 
		Les majors de l'industrie de la musique disent que
		leurs difficultés économiques sont dues aux
		téléchargements non autorisés (via le peer to peer
		notamment) et ont fortement influencé la rédaction du
		traité OMPI et de l'EUCD.  Cependant les faits leurs
		donnent tort et leurs difficultés sont dues à une
		conjoncture difficile. Aux États-Unis les ventes ont
		déclinées de 15\% sur les deux dernières années dont
		2,5\% seulement sont imputables à des réduction
		d'habitudes d'achat de personnes utilisant des
		services d'échange de musique non soumis à des mesures
		techniques de protection. Les majors ne proposant
		aucun service concurrent qui réponde aux attentes
		exprimées par ces consommateurs, la réduction de 2,5\%
		pourrait donc être une simple sanction de ce manque.

	      \par 
		Dans tous les domaines de la création, les usages non
		autorisés mais licites que sont le fair use ou la
		copie privée sont générateurs de richesse économique.
		En exposant plus fréquemment les personnes aux oeuvres elles
		en font un consommateur potentiel.

	      Références: 
	      \begin{itemize}
		\item {\bf Les téléchargements sauvent l'industrie de la musique.} INT: Août 2002. Downloads Save The Music Business. \\
http://www.forrester.com/ER/Research/Report/Summary/0,1338,14854,FF.html
		\item {\bf Le piratage crée une taxe progressive.} US: Novembre 2002. Piracy is Progressive Taxation, and Other Thoughts on the Evolution of Online Distribution. \\
http://www.openp2p.com/pub/a/p2p/2002/12/11/piracy.html
		\item {\bf Les dommages sociaux de la réduction du fair use.} INT: 2001. Information rights and intellectual freedom, Julie E. Cohen (page 8, 9 et 11). \\
http://www.law.georgetown.edu/faculty/jec/intellfreedom.pdf
	      \end{itemize}

	  \end{description}

	\item[{\bf 14. Porte atteinte au droit de divulgation des logiciels}]
	  \par 
	    Un auteur a le droit de divulguer son oeuvre. Dans le cas
	    du logiciel, ce droit moral inaliénable ne souffre pas
	    d'exception. Or, l'EUCD rend illicite la divulgation de
	    certains logiciels et crée ainsi un domaine nouveau, celui
	    des logiciels hors la loi. Il est difficile, voire
	    impossible, de déterminer quels logiciels en font partie
	    en raison de l'imprécision des termes \verb|"|mesure technique\verb|"|,
	    \verb|"|efficace\verb|"| et \verb|"|contournement\verb|"|. Ce flou légal joue en
	    faveur des puissants et au détriment du grand public, des
	    auteurs et des petites entreprises. Un motif aussi ténu
	    que l'absence de mesure technique de protection permettant d'engager des
	    poursuites (cas Blizzard vs bnetd), les plus riches
	    peuvent intimider les moins riches par une menace de
	    procès.

	  Références: 
	  \begin{itemize}
	    \item {\bf L'EUCD veut interdire la distribution de certains logiciels.} EU: 2002. L'articulation des exceptions au droit d'auteur et des mesures techniques de protection face à l'harmonisation (page 9). \\
http://www.droit-ntic.com/pdf/dir\_da.pdf
	    \item {\bf Le DMCA interdit la distribution de logiciels permettant de lire un eBook.} US: Décembre 2002. U.S. v. ElcomSoft \& Sklyarov. \\
http://www.eff.org/IP/DMCA/US\_v\_Elcomsoft/us\_v\_elcomsoft\_faq.html\#IllegalInUSA 
	    \item {\bf Le DMCA utilisé pour intimidation sur la base de l'absence de mesures techniques de protection.} US: 19 Février 2002. Blizzard v. bnetd \\
http://www.eff.org/IP/Emulation/Blizzard\_v\_bnetd/20020219\_blizzard\_bnetd\_letter.html
	    \item {\bf Le droit de lire.} INT: Février 1997 mis à jour en 2002. The Right to Read, Communications of the ACM (Volume 40, Number 2), \\
http://www.gnu.org/philosophy/right-to-read.html
	  \end{itemize}

	\item[{\bf 15. Contrarie l'harmonisation légale}]
	  \par 
	    Un objectif majeur des directives européennes est
	    d'harmoniser les législations de tous les pays. Or, trois
	    points laissent présager que les transpositions en droit
	    national dans les pays de la communauté européenne seront
	    hétérogènes:

	  \begin{itemize}
	    \item 
	      Les nombreuses exceptions optionnelles (20);

	    \item 
	      La grande complexité de l'article 6.4;

	    \item 
	      L'imprécision ou le caractère tautologique des
	      définitions de certains termes fondamentaux: \verb|"|mesure
	      technique\verb|"|, \verb|"|efficace\verb|"| et \verb|"|contournement\verb|"|.

	  \end{itemize}

	  Références:

	  \begin{itemize}
	    \item {\bf La complexité de l'article 6 de l'EUCD.} FR: Janvier 2002. Propriétés intellectuelles. p52-57. Gilles Vercken, Recherche clarté désespérément: à propos de l'article 6.4 de la directive du 22 mai 2001. {\it Je souhaite citer les opinions d'éminents juristes et professeurs de droit à propos de l'article 6.4 de la directive. \verb|"|C'est l'une des questions des plus épineuses\verb|"|, \verb|"|le résultat n'a pas le mérite de la clarté\verb|"| - et, à propos du paragraphe 2 sur la copie privée, \verb|"|c'est l'ensemble du paragraphe qui échappe à la compréhension\verb|"| nous dit Séverine Dussolier, chercheuse au CRID; \verb|"|un texte dont l'application s'avérera très délicate\verb|"|, d'après le Professeur Christophe Caron. \verb|"|Un texte très, voire trop complexe\verb|"|, affirment le Professeur Alain Strowel et Séverine Dusollier. \verb|"|Le texte laisse perplexe\verb|"| écrit le Professeur Pierre Sirinelli et il ajoute : \verb|"|Les Etats seront sans doute embarassés au moment de transposer le texte communautaire\verb|"|.}
	    \item {\bf L'EUCD n'atteint pas l'objectif d'harmonisation.} EU: 2002. L'articulation des exceptions au droit d'auteur et des mesures techniques de protection face à l'harmonisation (page 19 et 63). \\
http://www.droit-ntic.com/pdf/dir\_da.pdf
	    \item {\bf L'EUCD n'atteint pas l'objectif d'harmonisation.} EU: Octobre 2000. Why the Copyright Directive is Unimportant, and Possibly Invalid (paragraphe {\it What makes the Directive a total failure, in terms of harmonisation, ...}). \\
http://www.ivir.nl/publications/hugenholtz/opinion-EIPR.html
	  \end{itemize}

      \end{description}

  \end{document}

### 2003-02-01 projects/eucd/documents/introduction.fr.tex
Moved projects to their right places. Added thankgnus.
svn path=/trunk/; revision=3104

%% 
%% Language: French
%%

%{{{ Header

\documentclass[a4wide,12pt]{scrartcl}
\usepackage{german,umlaut}
\usepackage{fancyheadings}
\usepackage{alltt}
\usepackage[german]{babel}

\selectlanguage{german}
\pagestyle{fancyplain}

\lhead[\fancyplain{}{\bfseries\thepage}]
        {\fancyplain{}{\bfseries\rightmark}}
\rhead[\fancyplain{}{\bfseries\leftmark}]
        {\fancyplain{}{\bfseries\thepage}}
\cfoot{}
\renewcommand{\sectionmark}[1]{\markright{\thesection\ #1}}
\renewcommand{\thesection}{\S~\arabic{section}}

%
% alternative ways of doing it...
%

% this makes the references work correctly:
%\newcommand{\A}{\subsection{}}
%\renewcommand{\subsectionmark}[1]{\markright{\thesection\ }}

% this looks better:
\newcounter{absatz}[section]
\newcommand{\A}{\par\vspace{1ex}
                \stepcounter{absatz}\noindent(\arabic{absatz})~~}

%}}}

\begin{document}

%{{{ 0 Préambule

\thispagestyle{empty}
%\begin{center}
%\Huge\bf STATUTS
%\end{center}

\section*{Menace sur la copie privée: désir de monopole ?}

En 1878, lors de son discours d'ouverture du congrès littéraire
international, Victor Hugo contribuait à fonder le droit d'auteur avec
ces mots: "Le livre, comme livre, appartient à l'auteur, mais comme
pensée, il appartient - le mot n'est pas trop vaste - au genre
humain. Toutes les intelligences y ont droit. Si l'un des deux droits,
le droit de l'écrivain et le droit de l'esprit humain, devait être
sacrifié, ce serait, certes, le droit de l'écrivain, car l'intérêt
public est notre préoccupation unique, et tous, je le déclare, doivent
passer avant nous". Ses vues ne plaisaient guère au cercle des
libraires qui s'opposait, en particulier, à son idée du domaine public
en prétendant qu'elle nuirait au commerce et donc à toute la
profession littéraire. Et nous voici, deux siècles plus tard,
confrontés à une volonté de changement du droit d'auteur dont on nous
dit qu'elle permet de réduire le manque à gagner des commerçants. Le
cercle des libraires a certes disparu mais les majors du disque et les
éditeurs tiennent son rôle. Le temps n'a pas atténué l'influence des
commerçants ni, fort heureusement, la force du discours Hugolien.\\

Les "mesures techniques de protection" sont la pomme de la discorde.
Celle utilisée sur les DVD est largement répandue: elle permet de
cacher le contenu du film en le mélangeant selon une méthode appellée
CSS. Chaque logiciel permettant de visionner un film doit auparavant
remettre son contenu dans l'ordre. Si le logiciel ne connait pas CSS,
il ne peut donc projeter le film. Bien entendu, l'inventeur de CSS a
gardé le secret et ne l'a divulgué qu'à des personnes choisies. Or il
s'est trouvé qu'un programmeur avait acquis un DVD et ne disposait
d'aucun logiciel pour le voir. A force de travail, il a réussi à
deviner par lui-même le secret de CSS et écrit un petit logiciel,
appelé DeCSS, permettant de remettre en ordre son film. Il l'a ensuite
intégré à un logiciel d'affichage et a enfin profité de son
acquisition. Depuis lors, et comme DeCSS est un Logiciel Libre
disponible à tous, chacun peut voir le contenu du DVD qu'il a acheté
ou loué sans obligation d'acquérir un logiciel approuvé par
l'inventeur de CSS.\\

Comme on l'imagine, l'exemple se décline pour toute forme de création
immatérielle, qu'il s'agisse de textes, de sons ou d'images. Il ne se
trouve d'ailleurs personne pour prétendre concevoir une mesure
technique de protection infaillible: dans le meilleur des cas elle
résiste quelque mois à la créativité des hommes. C'est ce constat qui
a mis un terme, vers la fin des années 80, à la course aux mesures
techniques de protection. Dans la décade qui a suivi, elles ont
progressivement disparu. L'internet s'est développé avec le succès que
l'on sait, propulsant sur les réseaux une grande richesse de créations
immatérielles dont l'immense majorité n'est protégée par aucune mesure
technique. Les auteurs de l'immatériel, participants activement à
cette évolution, ne sont pas pour autant privés de moyens légaux pour
protéger leurs droits et les défendent chaque jour grâce au droit
d'auteur que nous connaissons.\\

Le nouveau cercle des libraires, c'est-à-dire quelques puissantes
multinationales, ont vu dans les mutations techniques l'opportunité de
modifier les lois afin d'acquérir un contrôle nouveau sur chaque
citoyen. Constituées en lobby, elles ont reclamé et obtenu en 1996 un
traité international (Traité OMPI sur le droit d'auteur) imposant aux
États signataires de prendre des mesures légales appropriées pour
punir le contournement de mesures techniques de protection. Quelle
idée surprenante, quand on y pense: si j'ai obtenu le droit de jouir
d'une oeuvre et que je dois pour cela contourner une mesure technique,
comme l'a fait l'auteur de DeCSS, quel tort ai-je causé à l'auteur ?
Si je suis un contrefacteur ayant la ferme intention de jouir de
l'oeuvre sans la permission de l'auteur, je suis déjà un délinquant et
ajouter à mon délit celui de contournement de la mesure technique ne
change rien à l'affaire. Mais, on l'a dit plus haut, l'objectif
recherché n'est ni la punition des contrefacteurs ni la défense des
intérêts légitimes de l'auteur. Il s'agit d'instituer un contrôle.\\

L'inventeur d'une mesure technique de protection prétend détenir un
monopole sur son utilisation. Il entend ainsi imposer à tous des
conditions contractuelles, dont il détermine les termes à sa
convenance, et qui contrôlent le fonctionnement de tout dispositif qui
met en oeuvre la mesure technique de protection. Par exemple, si un
créateur de logiciel permettant de lire un DVD demande la permission à
l'inventeur de CSS, il l'obtiendra à condition que son logiciel se
comporte d'une certaine façon. Le logiciel devra, entre autre,
interdire à l'utilisateur de passer outre la séquence d'introduction
contenant diverses publicités. L'inventeur de CSS a attaqué en justice
le créateur de DeCSS car il lui devenait impossible d'imposer ses
conditions contractuelles, il perdait le contrôle.\\

On cherche en vain à comprendre en quoi les auteurs, les titulaires de
droits ou les sociétés d'auteurs sont concernés par de telles
manoeuvres. Après tout, l'auteur d'un film ou d'une mélodie qui
choisit de distribuer son oeuvre sur un support numérique n'a aucune
influence sur les logiciels qui vont servir à son lecteur. Un
industriel français doit pouvoir décider en toute indépendance de
créer un logiciel permettant de jouir d'une oeuvre numérique. Un
auteur cherchant à imposer des conditions à cet industriel se verrait
immédiatement accusé de fausser les règles de la concurrence. A
l'inverse, un consommateur qui tenterait de forcer l'auteur à fournir,
par exemple, le moyen de faire des copies privées, se verrait opposer
que cela dépend exclusivement des logiciels, sur lesquels l'auteur n'a
pas et ne peut pas avoir prise.  La mesure technique de protection ne
permet donc aucun contrôle par l'auteur de l'utilisation de son
oeuvre, ce contrôle est tout entier dans les relations contractuelles
entre l'inventeur de la mesure technique de protection et les
créateurs de logiciels.\\

La question se pose donc au législateur dans ces termes : instituer un
monopole sur les mesures techniques de protection permet-il de mieux
protéger les auteurs dans l'environnement numérique ? On opposera à
cela que les textes européens (directive du 22 mai 2001 relative au
droit d'auteur et aux droits voisins dans la société de l'information
ou EUCD) ne parlent à aucun moment de monopole sur les mesures
techniques de protection. Mais s'il n'y a pas monopole, il s'en suit
par exemple que DeCSS peut être légalement créé et distribué comme
concurrent des logiciels autorisés par l'inventeur de CSS et selon des
conditions contractuelles différentes. Dans ce cas la mesure technique
de protection devient donc inutile. Si, au contraire, le législateur
entend effectivement accorder un monopole à l'inventeur de la mesure
technique de protection, il lui permet d'imposer à tous les conditions
contractuelles de son choix. L'auteur ne dispose d'aucun moyen légal
pour s'assurer que ces conditions contractuelles défendent ses
intérêts. La loi glisse vers la technique et oblige l'auteur à
négocier les modalités d'usage de son oeuvre avec l'inventeur de la
mesure technique de protection.\\

L'immense confusion qui règne depuis des années sur l'EUCD et la
grande complexité de sa rédaction n'ont pas d'autre source qu'un refus
d'énoncer clairement le problème comme nous venons de le faire. On ne
peut pretexter que les mutations de l'environnement numérique en sont
la cause, elles sont seulement l'excuse qui permet de cacher un désir
de contrôle. Ceux qui réclament un nouveau monopole savent que pour
l'inscrire dans la loi il faudrait en mesurer soigneusement les
conséquences pour toute la société. Il faudrait mettre en balance et
équilibrer les bénéfices et les privations qu'il entraine pour
chacun. Quel impact aurait-il sur les auteurs et les titulaires de
droits ? Sur la libre concurrence ?  Sur les droits des consommateurs
?  Sur la protection de la vie privée ?  Sur la mission des
bibliothèques ? Or, tout montre que le déséquilibre est grand et que
les bénéfices tirés d'un tel monopole par une minorité se ferait au
détriment de tous. \\

Loïc Dachary, fondateur de http://EUCD.INFO/, 15 Janvier 2003

%}}}

\end{document}

### 2003-02-01 tools/build.pl
Changed sociological to community. Fixed so that index.html files are generated.
svn path=/trunk/; revision=3106

@@ -411,6 +411,13 @@ while (my ($path, undef) = each %countries) {
           ! -f "$_/index.$lang.html") {
         link("$_/$base.$lang.html", "$_/index.$lang.html");
       }
+      if (! -f "$_/index.html") {
+        if (-f "$_/index.en.html") {
+          link("$_/index.en.html", "$_/index.html");
+        } elsif (-f "$_/$base.en.html") {
+          link("$_/$base.en.html", "$_/index.html");
+        }
+      }
     }
   }
 }

### 2003-02-01 tools/validate.pl
Added validation tool :-)
svn path=/trunk/; revision=3113

@@ -1,0 +1,13 @@
+#! /usr/bin/perl
+#
+# This is a simple wrapper to XML::LibXML. It will validate (more or less)
+# an XML file. Use this before committing something to CVS to make sure that
+# the file is at least parseable as XML.
+#
+use XML::LibXML;
+
+my $parser = XML::LibXML->new;
+
+foreach (@ARGV) {
+  my $doc = $parser->parse_file($_);
+}

### 2003-02-01 tools/build.pl
Added switch to support "non-output" run.
svn path=/trunk/; revision=3115

@@ -60,12 +60,13 @@ our %languages = (
 # Parse the command line options. We need two; where to put the finished
 # pages and what to use as base for the input.
 #
-getopts('o:i:duq', \%opts);
+getopts('o:i:duqn', \%opts);
 unless ($opts{o} && $opts{i}) {
-  print STDERR "Usage: $0 [-q] [-u] [-d] -o <output directory> -i <input directory>\n";
+  print STDERR "Usage: $0 [-q] [-u] [-d] [-n] -o <output directory> -i <input directory>\n";
   print STDERR "  -q   Quiet\n";
   print STDERR "  -u   Update only\n";
   print STDERR "  -d   Print some debug information\n";
+  print STDERR "  -n   Don't write any files\n";
   exit 1;
 }
 
@@ -393,7 +394,8 @@ while (my ($file, $langs) = each %bases) {
 
 	print "Writing: $opts{o}/$dir/$file.$lang.html\n" if $opts{d};
 
-	$stylesheet->output_file($results, "$opts{o}/$dir/$file.$lang.html");
+	$stylesheet->output_file($results, "$opts{o}/$dir/$file.$lang.html")
+		unless $opts{n};
     }
   }    
   print STDERR "\n" unless $opts{q};
@@ -409,13 +411,16 @@ while (my ($path, undef) = each %countries) {
     while (my ($lang, undef) = each %languages) {
       if (-f "$_/$base.$lang.html" &&
           ! -f "$_/index.$lang.html") {
-        link("$_/$base.$lang.html", "$_/index.$lang.html");
+        link("$_/$base.$lang.html", "$_/index.$lang.html")
+		unless $opts{n};
       }
       if (! -f "$_/index.html") {
         if (-f "$_/index.en.html") {
-          link("$_/index.en.html", "$_/index.html");
+          link("$_/index.en.html", "$_/index.html")
+		unless $opts{n};
         } elsif (-f "$_/$base.en.html") {
-          link("$_/$base.en.html", "$_/index.html");
+          link("$_/$base.en.html", "$_/index.html")
+		unless $opts{n};
         }
       }
     }
@@ -430,7 +435,9 @@ while (my ($path, undef) = each %countries) {
 print STDERR "Copying misc files\n" unless $opts{q};
 foreach (grep { !/\.xsl$/ } grep(!/\.xhtml$/, @files)) {
[…]

### 2003-02-01 tools/build.sh
*** empty log message ***
svn path=/trunk/; revision=3116

@@ -1,12 +1,18 @@
 #! /bin/bash
-TMP=/home/fsfe/tmp
-DEST=/home/www/
+#
+# This is a simple wrapper that will build the web pages in two steps.
+# First, it will call build.pl with options that will not produce any
+# output files. If this run is successful, it will make a second run
+# and create the files.
+#
+SOURCE=/home/www/fsfe/
+DEST=/home/www/html/
 
-tools/build.pl -o $TMP -i .
+tools/build.pl -n -o $DEST -i $SOURCE
+
 if test $? -ne 0; then
    echo "Build not complete. Aborting."
    exit 1
 fi
-rm -r $DEST/*
-mv $TMP/* $DEST/
 
+tools/build.pl -o $DEST -i $SOURCE

### 2003-02-01 tools/build.pl
Fixed reset bug.
svn path=/trunk/; revision=3117

@@ -84,11 +84,11 @@ my @dirs = File::Find::Rule->directory()
 
 while (my ($path, undef) = each %countries) {
   print STDERR "Reseting path for $path\n" unless $opts{q};
-  rmtree($opts{o}.'/'.$path) unless $opts{u};
+  rmtree($opts{o}.'/'.$path) unless ($opts{u} || $opts{n});
   my @paths = map { $opts{o}."/$path/".$_ } grep(!/^\.\.?$/, @dirs);
   foreach (@paths) {
     print "Creating $_\n" if $opts{d};
-    mkpath($_);
+    mkpath($_) unless $opts{n};
   }
 }
 

### 2003-02-01 tools/build.sh
Fix
svn path=/trunk/; revision=3118

@@ -5,8 +5,8 @@
 # output files. If this run is successful, it will make a second run
 # and create the files.
 #
-SOURCE=/home/www/fsfe/
-DEST=/home/www/html/
+SOURCE=/home/www/fsfe
+DEST=/home/www/html
 
 tools/build.pl -n -o $DEST -i $SOURCE
 

### 2003-02-01 tools/build.pl
Removed Swedish section.
svn path=/trunk/; revision=3119

@@ -36,7 +36,6 @@ use File::Copy;
 #
 our %countries = (
   global => 'en',
-  se => 'sv',
   de => 'de',
   fr => 'fr' );
 

### 2003-02-01 tools/build.sh
There's something wrong with the world today.
svn path=/trunk/; revision=3120

@@ -8,11 +8,12 @@
 SOURCE=/home/www/fsfe
 DEST=/home/www/html
 
-tools/build.pl -n -o $DEST -i $SOURCE
+cd $SOURCE
+tools/build.pl -n -o $DEST -i .
 
 if test $? -ne 0; then
    echo "Build not complete. Aborting."
    exit 1
 fi
 
-tools/build.pl -o $DEST -i $SOURCE
+tools/build.pl -o $DEST -i .

### 2003-02-01 tools/build.pl
Bugfix.
svn path=/trunk/; revision=3121

@@ -60,14 +60,19 @@ our %languages = (
 # pages and what to use as base for the input.
 #
 getopts('o:i:duqn', \%opts);
-unless ($opts{o} && $opts{i}) {
-  print STDERR "Usage: $0 [-q] [-u] [-d] [-n] -o <output directory> -i <input directory>\n";
+unless ($opts{o}) {
+  print STDERR "Usage: $0 [-q] [-u] [-d] [-n] -o <output directory>\n";
   print STDERR "  -q   Quiet\n";
   print STDERR "  -u   Update only\n";
   print STDERR "  -d   Print some debug information\n";
   print STDERR "  -n   Don't write any files\n";
   exit 1;
 }
+
+# It might be nice to be able to specify this, but it will break things as
+# they are now. This is on the TODO list :-)
+
+$opts{i} = ".";
 
 $| = 1;
 

### 2003-02-01 tools/build.pl
Fixed broken links.
svn path=/trunk/; revision=3123

@@ -380,8 +380,14 @@ while (my ($file, $langs) = each %bases) {
         # In post-processing, we replace links pointing back to ourselves
         # so that they point to the correct language.
         #
-        foreach ($results->documentElement->getElementsByTagName("a")) {
+        my @hrefs = $results->documentElement->getElementsByTagName("a");
+        push(@hrefs, $results->documentElement->getElementsByTagName("A"));
+
+        foreach (@hrefs) {
           my $href = $_->getAttribute("href");
+          if ($href =~ /^http:\/\/www.fsfeurope.org/) {
+            $href =~ s/http:\/\/www.fsfeurope.org//;
+          }
           if ($href !~ /^http/) {
             if ($href !~ /\.html$/) {
               if (-d $opts{i}."/$href") {

### 2003-02-01 tools/build.sh
Made sure that tools/build.sh calls symlinks.
svn path=/trunk/; revision=3124

@@ -17,3 +17,6 @@ if test $? -ne 0; then
 fi
 
 tools/build.pl -o $DEST -i .
+
+cd $DEST
+symlinks

### 2003-02-01 tools/build.sh
Fixed script.
svn path=/trunk/; revision=3129

@@ -1,22 +1,25 @@
 #! /bin/bash
 #
 # This is a simple wrapper that will build the web pages in two steps.
-# First, it will call build.pl with options that will not produce any
-# output files. If this run is successful, it will make a second run
-# and create the files.
+# First, it will call build.pl with options to send the output to a
+# temporary directory. After that is done, it will move the contents of
+# that directory over the normal root.
 #
 SOURCE=/home/www/fsfe
 DEST=/home/www/html
+TMP=/home/www/tmp
 
 cd $SOURCE
-tools/build.pl -n -o $DEST -i .
+tools/build.pl -q -o $TMP -i .
 
 if test $? -ne 0; then
    echo "Build not complete. Aborting."
    exit 1
 fi
-
-tools/build.pl -o $DEST -i .
-
-cd $DEST
+cd $TMP
 symlinks
+
+mv $DEST ${DEST}.old
+mv $TMP $DEST
+rm -rf ${DEST}.old
+

### 2003-02-01 tools/build.pl
Removed unnecesary code.
svn path=/trunk/; revision=3133

@@ -380,10 +380,7 @@ while (my ($file, $langs) = each %bases) {
         # In post-processing, we replace links pointing back to ourselves
         # so that they point to the correct language.
         #
-        my @hrefs = $results->documentElement->getElementsByTagName("a");
-        push(@hrefs, $results->documentElement->getElementsByTagName("A"));
-
-        foreach (@hrefs) {
+        foreach ($results->documentElement->getElementsByTagName("a")) {
           my $href = $_->getAttribute("href");
           if ($href =~ /^http:\/\/www.fsfeurope.org/) {
             $href =~ s/http:\/\/www.fsfeurope.org//;

### 2003-02-01 tools/build.sh
Fixed build script so it calls correct symlinks.
svn path=/trunk/; revision=3147

@@ -17,7 +17,7 @@ if test $? -ne 0; then
    exit 1
 fi
 cd $TMP
-symlinks
+/usr/local/bin/symlinks
 
 mv $DEST ${DEST}.old
 mv $TMP $DEST

### 2003-02-03 tools/build.pl
Added debug statement.
svn path=/trunk/; revision=3163

@@ -268,6 +268,8 @@ while (my ($file, $langs) = each %bases) {
                       $files{$1} = $2;
                     } elsif ($2 eq $lang) {
                       $files{$1} = $2;
+                    } elsif (!exists $files{$1}) {
+                      $files{$1} = $2;
                     }
                  }
               }
@@ -283,6 +285,7 @@ while (my ($file, $langs) = each %bases) {
           my $auto_data = $sourcedoc->createElement("set");
 
           while (my ($base, $l) = each %files) {
+              print STDERR "Loading $base.$l.xml" if $opts{d};
               my $source_data = $parser->parse_file("$base.$l.xml");
               foreach ($source_data->documentElement->childNodes) {
                  my $c = $_->cloneNode(1);

### 2003-02-03 tools/build.pl
Left old file lying around.
svn path=/trunk/; revision=3164

@@ -285,7 +285,7 @@ while (my ($file, $langs) = each %bases) {
           my $auto_data = $sourcedoc->createElement("set");
 
           while (my ($base, $l) = each %files) {
-              print STDERR "Loading $base.$l.xml" if $opts{d};
+              print STDERR "Loading $base.$l.xml\n" if $opts{d};
               my $source_data = $parser->parse_file("$base.$l.xml");
               foreach ($source_data->documentElement->childNodes) {
                  my $c = $_->cloneNode(1);

### 2003-02-03 tools/build.sh
Fixed temporary directory.
svn path=/trunk/; revision=3166

@@ -7,7 +7,7 @@
 #
 SOURCE=/home/www/fsfe
 DEST=/home/www/html
-TMP=/home/www/tmp
+TMP=/home/www/tmp.$$
 
 cd $SOURCE
 tools/build.pl -q -o $TMP -i .

### 2007-01-31 projects/fla/FLA.ccasa.en.tex
updated FLA to reflect new licensing terms, added new versions of FLA for people to use
svn path=/trunk/; revision=7687

%% $Id: FLA.ccasa.en.tex,v 1.1 2007-01-31 16:55:36 shanecoughlan Exp $
%% 
%% Fiduciary License Agreement (Copyright Assignment)
%% of the Free Software Foundation Europe paying respect to the two
%% different Copyright traditions  
%%
%% Language: English
%%

%{{{ Header

\newcommand{\fiduciary}{FSFE}
\newcommand{\fiduciaryaddress}{\begin{alltt}
Free Software Foundation Europe
Talstrasse 110
40217 Dusseldorf
Germany
\end{alltt}}
\documentclass[a4wide,12pt]{article}
\usepackage{german,umlaut}
\usepackage{fancyheadings}
\usepackage{alltt}
\usepackage{epsfig}
\usepackage[english]{babel}

\selectlanguage{english}
\pagestyle{fancyplain}

\lhead[\fancyplain{}{\bfseries\thepage}]
        {\fancyplain{}{\bfseries\rightmark}}
\rhead[\fancyplain{}{\bfseries\leftmark}]
        {\fancyplain{}{\bfseries\thepage}}
\cfoot{}
\renewcommand{\sectionmark}[1]{\markright{\thesection\ #1}}
\renewcommand{\thesection}{\S~\arabic{section}}

%
% alternative ways of doing it...
%

% this makes the references work correctly:
%\newcommand{\A}{\subsection{}}
%\renewcommand{\subsectionmark}[1]{\markright{\thesection\ }}

% this looks better:
\newcounter{absatz}[section]
\newcommand{\A}{\par\vspace{1ex}
                \stepcounter{absatz}\noindent(\arabic{absatz})~~}

% make the page a little wider
%\addtolength{\textwidth}{3cm}
%\addtolength{\headwidth}{3cm}
%\addtolength{\hoffset}{-1.5cm}
%\addtolength{\textheight}{2cm}
\addtolength{\textwidth}{3cm}
\addtolength{\headwidth}{3cm}
\addtolength{\hoffset}{-1.5cm}
\addtolength{\textheight}{4cm}

%}}}

\begin{document}

%{{{ Head of front page

\thispagestyle{empty}

\begin{center}
{\LARGE\bf Fiduciary Licence Agreement}\\
(Version 1.2)
\end{center}

\begin{center}
\sc\small Copyright (C) 2002-2007 FSFE, e.V.,\\
\sc\small Talstrasse 110,
40217 Dusseldorf,
Germany\\
\sc\small This licence is released under the terms of\\
the Creative Commons Attribution/Share-alike licence version 2.5.
\end{center}

%}}}

\begin{center}
{\Large\bf Preamble}
\end{center}

Access to software determines participation in a digital society. To
secure equal participation in the information age, the Free Software
Foundation Europe (FSFE) pursues and is dedicated to the furthering of
Free Software, defined by the freedoms to use, study, modify and
copy. Independent of the issue of commercial exploitation, it is
proprietary, freedom-diminishing licensing that works against the
interests of people and society at large, which is therefore rejected
by FSFE.\\

The purpose of this agreement is to ensure the lasting protection of
Free Software by making FSFE the fiduciary of the author's interests.
It empowers FSFE -- and its sister organisations -- to uphold the
interests of Free Software authors and protect them in court, if
necessary.\\

FSFE is given the right to relicense the software as necessary for the
long-term legal maintainability and protection of the software. The
agreement also grants the author an unlimited amount of non-exclusive
licences by FSFE, which allow using and distributing the program in
other projects and under other licences. \\

The contracting parties sign the following agreement in full
consciousness that by the grant of exclusive licence to the Free
Software Foundation Europe e.V. and by the administration of these
rights the FSFE becomes trustee of the author's interests for
the benefit of Free Software.\\

\pagebreak
\begin{center}
{\Large\bf Agreement}
\end{center}

Between 
{\bf (please cross out unsuitable alternatives)}\\
\begin{itemize}
\item the author
\begin{alltt}
    _______________________________________________
    _______________________________________________ (Occupation, D.o. Birth)
    _______________________________________________ (Street)
    _______________________________________________ (ZIP, City, Country)
    _______________________________________________ (Pseudonym, Email)
\end{alltt}
\item the owner of the exclusive licence,
\begin{alltt}
    _______________________________________________ (Company or name)
    _______________________________________________ (HQ or addr.)
    _______________________________________________ (managing director)
\end{alltt}
\begin{itemize}
\item acquired by virtue of a contract date as of
\item contracting party:
\item acquired as employer in the context of a work and service
  relationship
\end{itemize}
\end{itemize}
\begin{flushright}
- hereinafter referred to as "Beneficiary" --
\end{flushright}
and the
\begin{alltt}
    Free Software Foundation Europe, e.V.,
    Talstrasse 110
    40217 Dsseldorf
    Germany
\end{alltt}
\begin{flushright}
- hereinafter referred to as "FSFE" --
\end{flushright}
the following agreement is entered into:\\

\section{Grant}
\A Subject to the provision of {\S}~2, Beneficiary assigns to FSFE the
Copyright in computer programs and other copyrightable material
world-wide, or in countries where such an assignment is not
possible,\footnote{Countries where assignments of the copyright in a
work are impossible include, but are not limited to, Germany, Austria,
Slovenia and Hungary.} grants an exclusive licence, including, inter
alia:
\begin{enumerate}
\item the right to reproduce in original or modified form;
\item the right to redistribute in original or modified form;
\item the right of making available in data networks, in particular
  via the Internet, as well as by providing downloads, in original or
  modified form;
\item the right to authorize third parties to make derivative works of
  the Software, or to work on and commit changes or perform this
  conduct themselves.
\end{enumerate}
\A Beneficiary's moral or personal rights remain unaffected by this
Agreement.\\ 
\A In some countries, the law may provide that the employer is deemed
to be the owner of the rights on materials developed by an employee in
the course of his or her employment, unless the parties have agreed
otherwise. The Beneficiary is aware of these provisions, and therefore
warrants, represents and guarantees that the Subject Matter is free of
any of his or her employer's exclusive exploitation rights.

\section{Subject Matter}

The rights and licences granted in {\S}~1 are subject to all
``Software'' and ``Documentation.'' For the purpose of this Agreement,
``Software'' shall mean all computer programs, copyrightable sections
of computer programs or modifications of computer programs that have
been developed or programmed by Beneficiary and that are specified in
this Agreement below or that are listed in Exhibit A attached to this
Agreement and dated and signed by the contracting parties. Likewise,
``Documentation'' shall refer to all manuals and documentation written
by Beneficiary alongside and usually distributed with the ``Software''
and are similarly specified below or listed in Exhibit A:
\begin{alltt}
    __________________________________________________________________
    __________________________________________________________________
    __________________________________________________________________
\end{alltt}
Except in countries where such an assignment is not
possible,\footnote{Countries where assignments of the copyright in a
future work are impossible include, but are not limited to, France.}
the rights [and licences] granted under this agreement by Beneficiary
shall also include future developments, future corrections of errors
or faults and other future modifications and derivative works of the
software that Beneficiary obtains copyright ownership.  Excluded from
this provision are modifications that are not derived from the subject
matter and that have to be regarded as independent and original
software.

\section{FSFE's Rights and \\Re-Transfer of Non-Exclusive Licence}
\A FSFE shall exercise the granted rights and licences in its
own name. Furthermore, FSFE shall be authorized to enjoin third
parties from using the software and forbid any unlawful or copyright
infringing use of the Software, and shall be entitled to enforce all
its rights in its own name in and out of court. FSFE shall also
be authorized to permit third parties to exercise FSFE's rights
in and out of court.\\

\A FSFE grants to Beneficiary a non-exclusive, worldwide, perpetual
and unrestricted licence in the Software. This right's [and licence's]
scope shall encompass and include all the rights [and licences]
specified in {\S}~1. Furthermore, FSFE grants to Beneficiary
additional non-exclusive, transferable license to use, reproduce,
redistribute and make available the Software as needed for releases of
the Software under other licences.  This re-transfer shall not limit
the scope of FSFE's exclusive licence in the Software and FSFE's
rights pursuant to {\S}~1.\\

\A FSFE shall only exercise the granted rights and licences in
accordance with the principles of Free Software as defined by 
the Free Software Foundations. FSFE guarantees to use the rights 
and licences transferred in strict accordance with the
regulations imposed by Free Software licences, including, but not
limited to, the GNU General Public Licence (GPL) or the GNU Lesser
General Public Licence (LGPL) respectively. In the event FSFE violates
the principles of Free Software, all granted rights and licences shall
automatically return to the Beneficiary and the licences granted
hereunder shall be terminated and expire.\\

\A The transfer of the rights and licences specified in {\S}~1 shall
be unrestricted in territory and thus shall apply world-wide and be
temporally unlimited.

\section{Miscellaneous}
\A Regarding the succession of rights in this contractual
relationship, German law shall apply, unless this Agreement imposes
deviating regulations. In case of the Beneficiary's death, the
assignment of exclusive rights shall continue with the heirs. In case
of more than one heir, all heirs have to exercise their rights through
a common authorized person.\\

\A Place of jurisdiction for all legal conflicts arising out of or in
connection with this Agreement is Munich, Germany.\\

\enlargethispage{2cm}
\vspace{3cm}
\begin{alltt}
__________________, ______________    __________________, ______________

__________________________________    __________________________________
           (Beneficiary)                         (FSFE)
\end{alltt}

\end{document}

### 2012-07-20 cgi-bin/stacs-register-capacity.pl
changing more old links
svn path=/trunk/; revision=23888

@@ -43,7 +43,7 @@ close MAIL;
 # -----------------------------------------------------------------------------
 
 print "Content-type: text/html\n\n";
-open TEMPLATE, "/home/www/html/global/projects/stacs/tmpl-london." . $query->param("language") . ".html";
+open TEMPLATE, "/home/www/html/global/campaigns/stacs/tmpl-london." . $query->param("language") . ".html";
 while (<TEMPLATE>) {
   s/:REFERENCE:/$reference/g;
   print;

### 2012-07-20 cgi-bin/stacs-register-workshop.pl

@@ -43,7 +43,7 @@ close MAIL;
 # -----------------------------------------------------------------------------
 
 print "Content-type: text/html\n\n";
-open TEMPLATE, "/home/www/html/global/projects/stacs/tmpl-belgrade." . $query->param("language") . ".html";
+open TEMPLATE, "/home/www/html/global/campaigns/stacs/tmpl-belgrade." . $query->param("language") . ".html";
 while (<TEMPLATE>) {
   s/:REFERENCE:/$reference/g;
   print;

### 2012-07-20 cgi-bin/ticket.pl

@@ -29,21 +29,21 @@ my $reference = "elnticket.$date." . substr($time, -5);
 
 print "Content-type: text/html\n\n";
 
-open TEMPLATE, "/home/www/html/global/projects/ftf/tickets/tmpl-buy.$language.html";
+open TEMPLATE, "/home/www/html/global/activities/ftf/tickets/tmpl-buy.$language.html";
 while (<TEMPLATE>) {
   if (/:FORM:/) {
     my $passphrase = "Only4TestingPurposes";
     my $shastring = 
-        "ACCEPTURL=http://www.fsfe.org/projects/ftf/tickets/thankyou.$language.html$passphrase" .
+        "ACCEPTURL=http://www.fsfe.org/activities/ftf/tickets/thankyou.$language.html$passphrase" .
         "AMOUNT=$amount100$passphrase" .
-        "CANCELURL=http://www.fsfe.org/projects/ftf/tickets/cancel.$language.html$passphrase" .
+        "CANCELURL=http://www.fsfe.org/activities/ftf/tickets/cancel.$language.html$passphrase" .
         "COM=$text$passphrase" .
         "CURRENCY=EUR$passphrase" .
         "LANGUAGE=$language$passphrase" .
         "ORDERID=$reference$passphrase" .
         "PMLISTTYPE=2$passphrase" .
         "PSPID=40F00871$passphrase" .
-        "TP=http://www.fsfe.org/projects/ftf/tickets/tmpl-concardis.$language.html$passphrase";
+        "TP=http://www.fsfe.org/activities/ftf/tickets/tmpl-concardis.$language.html$passphrase";
     my $shasum = uc(sha1_hex($shastring));
     print "    <form name=\"donate\" action=\"https://secure.payengine.de/ncol/prod/orderstandard.asp\" method=\"post\">\n";
     print "      <!-- payment parameters -->\n";
@@ -54,11 +54,11 @@ while (<TEMPLATE>) {
     print "      <input type=\"hidden\" name=\"currency\"     value=\"EUR\"/>\n";
     print "      <input type=\"hidden\" name=\"language\"     value=\"$language\"/>\n";
     print "      <!-- interface template -->\n";
-    print "      <input type=\"hidden\" name=\"TP\"           value=\"http://www.fsfe.org/projects/ftf/tickets/tmpl-concardis.$language.html\"/>\n";
[…]
